### 2024-01-07 written/paper.tex
writing

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12,
]{article}
\usepackage{lmodern}
\usepackage[table,xcdraw]{xcolor}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Empirical Option Returns and the risk-free Rate in Crypto Asset Markets},
  pdfauthor={Julian Winkel; Wolfgang K. Härdle},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin = 2.5cm]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{float}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{subfig}
\usepackage{caption}
\usepackage{eucal}
\floatplacement{figure}{H}

\title{Pricing Kernels and Risk Premia implied in Bitcoin Options}
\author{Julian Winkel\footnote{International Research Training Group 1792;
  Humboldt-Universität zu Berlin,
  \href{mailto:julian.winkel@hu-berlin.de}{\nolinkurl{julian.winkel@hu-berlin.de}}} \and Wolfgang K. Härdle\footnote{BRC Blockchain Research Center,
  Humboldt-Universität zu Berlin, Berlin, Germany; Sim Kee Boon
  Institute, Singapore Management University, Singapore; WISE Wang Yanan
  Institute for Studies in Economics, Xiamen University, Xiamen, China;
  Dept. Information Management and Finance, National Chiao Tung
  University, Hsinchu, Taiwan, ROC; Dept. Mathematics and Physics,
  Charles University, Prague, Czech Republic, Grants -- DFG IRTG 1792
  and the Czech Science Foundation's grant no. 19-28231X / CAS XDA
  23020303 gratefully acknowledged,
  \href{mailto:haerdle@hu-berlin.de}{\nolinkurl{haerdle@hu-berlin.de}}}}
\date{7 January 2024}

\begin{document}
\maketitle

\newcommand{\quantnet}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,}
\newcommand{\quantinar}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/quantinar_logo.png}}\,}

\definecolor{bittersweet}{rgb}{1.0, 0.44, 0.37} 
\definecolor{ao}{rgb}{0.0, 0.5, 0.0}
\definecolor{bananamania}{rgb}{0.98, 0.91, 0.71}
\definecolor{bluegray}{rgb}{0.4, 0.6, 0.8}

\pagenumbering{roman}

\begin{abstract}

\singlespacing 
abstract

 \medskip
 \textbf{Keywords:} Bitcoin, Deribit, pricing kernel, risk aversion, speculation, hedging  \\
 \indent \textbf{JEL Classification:} C14, C50, G10  \\
 \bigskip

\end{abstract}

\pagenumbering{arabic}

\newpage

\section{Introduction}

Another, so far rarely used pathway is the assessment of the risk-return characteristics of Bitcoin through the lens of the options written on it, which allow new model-free insights into the existence and degree of a leverage effect, the existence and size of risk premia, resulting market inefficiencies (measurable in economic loss / money lost) and to explore the option (pun intended)/possibility of curing named inefficiencies under transaction cost and liquidity constraints.

This reasoning follows the insights of Shumway and Bakshi, who separately evaluated option returns in equity markets by means of delta-neutral option positions in the early 2000s.

Bakshi and Shumway both found that expected call returns exceed those of the underlying and that expected put returns underperform the underlying. Both also found that the effect increases (decreases) with the strike. 
Bakshi found that zero-beta, at-the-money straddle positions produce average losses of approximately three percent per week and suggests that the reason may be a systematic volatility premium being priced in options.

Following Shumway, the risk-return characteristics of options can be separated in two elements, a leverage effect and a curvature effect due to the sensitivity of option payoffs to higher moments of the underlying. 
First, the holder (buyer) of an option is subject to a leverage effect since exposure to the underlying can be obtained through the payment of a premium, which is small compared to the price of the underlying. 
If there is a leverage effect, then (long) call options should outperform their underlying and (long) put options should underperform their underlying. Furthermore, the effect should increase (decrease) with the strike price for calls (puts).
This reasoning allows a simple test of existence and size of the leverage effect and thus conclusions about the level of speculation associated to Bitcoin valuation.

(The second component, related to the sensitivity of options towards higher moments, implies that options are redundant assets...)

If options were redundant assets, then apart of the leverage effect, delta-neutral option portfolios should earn the risk-free rate (no premium) on average.
Until now, there is no risk-free rate available in the space of crypto assets. We establish a proxy or reference rate for a possible risk-free rate, which results from the perpetual futures written on Bitcoin and Ether.

Use Deribit and Binance perpetual futures for this, equally weighted.

(This proxy is clearly not risk-free due to counterparty risk, but we argue that a carry trade involving a short position on perpetual futures on multiple platforms in conjunction with a secured, on-chain underlying position reduces the platform-induced counterparty risk (which is the residual risk in the carry trade) as far as possible.) (This also leaves liquidity risk though!!)

All underlying code is available on \href{quantlet.com}{quantlet.com}
and a corresponding courselet is available on
\href{quantinar.com}{quantinar.com}.

\section{A Proxy for the risk-free Rate of Bitcoin and Ether}

Evidently, there is no established risk-free rate in crypto (or digital asset) markets. In the absence of riskless bond, we refer to the next best thing as a proxy for the risk free rate, which is the replication of the market-implied yield through a position which only has liquidity risk and counterparty risk. To be fair, counterparty risk also exists for risk-free bonds.

A large share of the trading volume in digital asset markets falls on the futures market. Among the futures trading, perpetual instruments have historically been and currently are the most liquid instruments (show both!). While it would be possible to use a yield from a long-short futures position (e.g. short a short-tenor future and long a longer-tenor future), we argue that in order to minimize liquidity risk, the most liquid instruments should be taken into consideration. These are the perpetual futures, which are listed on some exchanges.

As of 2024-01-07, the total Open Interest for Bitcoin Perpetual Futures contracts is \$19,343,021,437 with a total 24h Trade Volume of \$50,339,886,015.
Open Interest for Ether Perpetual Futures contracts is Total Open Interest \$14,945,055,482 and total 24h Trade Volume is \$18,097,004,167.

BTC Futures

Total Open Interest $1,405,918,350
Total 24h Trade Volume $324,820,482

ETH Futures
Total Open Interest $466,932,584
Total 24h Trade Volume $98,238,298

\begin{table}[]
\begin{tabular}{lll}
\textbf{Instrument}            & \textbf{Aggregate Open Interest (USD)} & \textbf{24 Hour Aggregate Trade Volume (USD)} \\
BTC Perpetual Futures & 19,343,021,437                & 50,339,886,015                       \\
ETH Perpetual Futures & 14,945,055,482                & 18,097,004,167                       \\
BTC Futures           & 1,405,918,350                 & 324,820,482                          \\
ETH Futures           & 466,932,584                   & 98,238,298                          
\end{tabular}
\end{table}

Source for liquidity: 
https://www.coingecko.com/en/derivatives
https://www.coingecko.com/en/derivatives/futures

% Source: https://www.tablesgenerator.com/#
% Please add the following required packages to your document preamble:
% \usepackage[table,xcdraw]{xcolor}
% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage[table,xcdraw]{xcolor}
% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}

As of 2024-01-07

Bitcoin Perpetuals with 
Total Open Interest \$19,343,021,437
Total 24h Trade Volume \$50,339,886,015

\begin{table}[]
\begin{tabular}{llrrrrrrr}
\textbf{Exchange}    & \textbf{Symbol} & \multicolumn{1}{l}{\textbf{Price}} & \multicolumn{1}{l}{\textbf{Index Price}} & \multicolumn{1}{l}{\textbf{Basis}} & \multicolumn{1}{l}{\textbf{Spread}} & \multicolumn{1}{l}{\textbf{Funding Rate}} & \multicolumn{1}{l}{\textbf{Open Interest}} & \multicolumn{1}{l}{\textbf{24h Volume}} \\
\# Binance (Futures) & BTCUSDT         & \$43,995.06                        & 44011.67 USDT                            & +0.087\%                           & 0.01\%                              & 0.010\%                                   & \$3,214,727,375                            & \$6,822,091,229.25                      \\
\# Binance (Futures) & BTCUSD\_PERP    & \$44,008.70                        & 44037.64 USD                             & +0.066\%                           & 0.01\%                              & 0.010\%                                   & \$808,036,100                              & \$973,585,131.72                        \\
\# BingX (Futures)   & BTC-USDT        & \$44,042.40                        & 44046.6 USDT                             & +0.044\%                           & 0.01\%                              & 0.010\%                                   & \$909,897,901                              & \$1,867,895,422.35                      \\
\# Bitget Futures    & BTCUSDT\_UMCBL  & \$44,049.08                        & 44033.85 USDT                            & +0.015\%                           & 0.01\%                              & 0.011\%                                   & \$1,790,553,246                            & \$5,701,072,691.79                      \\
\# Bitmart Futures   & BTCUSDT         & \$44,015.47                        & 44003.79 USDT                            & +0.024\%                           & 0.01\%                              & 0.010\%                                   & \$641,006,203                              & \$2,917,088,729.74                      \\
\# Bybit (Futures)   & BTCUSDT         & \$44,039.88                        & 44016.82 USDT                            & -0.003\%                           & 0.01\%                              & 0.010\%                                   & \$2,538,053,148                            & \$2,672,726,926.85                      \\
\# Bybit (Futures)   & BTCUSD          & \$43,995.50                        & 44037.2 USD                              & +0.095\%                           & 0.01\%                              & 0.010\%                                   & \$652,512,486                              & \$283,394,472.00                        \\
\# Deribit           & BTC-PERPETUAL   & \$44,020.00                        & 44028.17 USD                             & +0.019\%                           & 0.01\%                              & 0.004\%                                   & \$577,506,570                              & \$204,526,951.86                        \\
\# OKX (Futures)     & BTC-USDT-SWAP   & \$44,014.07                        & 43989.3 USDT                             & -0.006\%                           & 0.01\%                              & -0.009\%                                  & \$984,130,801                              & \$2,648,725,568.71                      \\
\# Prime XBT         & BTC/USD         & \$43,981.30                        & -                                        & -                                  & 0.02\%                              & -                                         & \$920,322,200                              & \$60,224,491.31                        
\end{tabular}
\end{table}

Ether Perpetuals

Total Open Interest \$14,945,055,482
Total 24h Trade Volume \$18,097,004,167

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage[table,xcdraw]{xcolor}
% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage[table,xcdraw]{xcolor}
% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}

\begin{table}[]
\begin{tabular}{llrrrrrrr}
\textbf{Exchange}         & \textbf{Symbol} & \multicolumn{1}{l}{\textbf{Price}} & \multicolumn{1}{l}{\textbf{Index Price}} & \multicolumn{1}{l}{\textbf{Basis}} & \multicolumn{1}{l}{\textbf{Spread}} & \multicolumn{1}{l}{\textbf{Funding Rate}} & \multicolumn{1}{l}{\textbf{Open Interest}} & \multicolumn{1}{l}{\textbf{24h Volume}} \\
\# Deepcoin (Derivatives) & ETHUSDT         & \$2,232.18                         & 2232.46 USDT                             & +0.047\%                           & 0.02\%                              & 0.006\%                                   & \$6,376,015,545                            & \$237,202,788.68                        \\
\# Binance (Futures)      & ETHUSDT         & \$2,227.11                         & 2227.74 USDT                             & +0.078\%                           & 0.01\%                              & 0.010\%                                   & \$1,680,165,204                            & \$2,843,630,044.97                      \\
\# Bybit (Futures)        & ETHUSDT         & \$2,228.15                         & 2228.15 USDT                             & +0.049\%                           & 0.01\%                              & 0.010\%                                   & \$1,027,647,498                            & \$540,402,664.24                        \\
\# OKX (Futures)          & ETH-USDT-SWAP   & \$2,227.96                         & 2227.07 USDT                             & +0.010\%                           & 0.01\%                              & -0.003\%                                  & \$584,652,223                              & \$1,748,266,659.63                      \\
\# Gate.io (Futures)      & ETH\_USDT       & \$2,228.62                         & 2227.03 USDT                             & -0.021\%                           & 0.01\%                              & 0.010\%                                   & \$508,608,257                              & \$189,744,908.47                        \\
\# BingX (Futures)        & ETH-USDT        & \$2,229.64                         & 2230.26 USDT                             & +0.062\%                           & 0.02\%                              & 0.010\%                                   & \$475,301,553                              & \$609,765,117.59                        \\
\# Prime XBT              & ETH/USD         & \$2,227.49                         & -                                        & -                                  & 0.04\%                              & -                                         & \$473,765,000                              & \$16,309,868.89                         \\
\# Deribit                & ETH-PERPETUAL   & \$2,228.05                         & 2228.59 USD                              & +0.024\%                           & 0.01\%                              & 0.004\%                                   & \$447,911,369                              & \$82,175,344.52                         \\
\# Binance (Futures)      & ETHUSD\_PERP    & \$2,227.69                         & 2228.74 USD                              & +0.047\%                           & 0.01\%                              & 0.010\%                                   & \$366,493,140                              & \$257,267,395.24                        \\
\# XT.COM (Derivatives)   & ETH\_USDT       & \$2,228.19                         & 2225.55 USDT                             & +0.031\%                           & 0.01\%                              & -0.010\%                                  & \$325,480,371                              & \$1,137,587,629.30                     
\end{tabular}
\end{table}

BTC Futures

with

Total Open Interest $1,405,918,350
Total 24h Trade Volume $324,820,482

\begin{table}[]
\begin{tabular}{llrrrrrrr}
\textbf{Exchange}                & \textbf{Symbol}    & \multicolumn{1}{l}{\textbf{Price}} & \multicolumn{1}{l}{\textbf{Index Price}} & \multicolumn{1}{l}{\textbf{Basis}} & \multicolumn{1}{l}{\textbf{Spread}} & \multicolumn{1}{l}{\textbf{Expires in}} & \multicolumn{1}{l}{\textbf{Open Interest}} & \multicolumn{1}{l}{\textbf{24h Volume}} \\
\# Binance (Futures)             & BTCUSD\_240329     & \$45,372.20                        & 44041.17 USD                             & -2.934\%                           & 0.01\%                              & 3 months                                & \$136,765,200                              & \$19,534,741.00                         \\
\# Binance (Futures)             & BTCUSD\_240628     & \$46,802.20                        & 44037.64 USD                             & -5.907\%                           & 0.02\%                              & 6 months                                & \$60,673,400                               & \$11,015,730.09                         \\
\# Binance (Futures)             & BTCUSDT\_240329    & \$45,340.32                        & 44011.62 USDT                            & -2.883\%                           & 0.04\%                              & -                                       & \$38,608,233                               & \$13,341,161.63                         \\
\# Binance (Futures)             & BTCUSDT\_240628    & \$46,588.03                        & 44011.62 USDT                            & -5.484\%                           & 0.02\%                              & -                                       & \$7,296,676                                & \$6,727,917.37                          \\
\# BitMEX (Derivative)           & XBTH24             & \$45,170.00                        & 44016.02 USD                             & -2.555\%                           & 0.03\%                              & 3 months                                & \$23,123,300                               & \$2,586,155.50                          \\
\# BitMEX (Derivative)           & XBTM24             & \$46,319.00                        & 44016.02 USD                             & -4.972\%                           & 0.15\%                              & 6 months                                & \$14,701,700                               & \$899,753.13                            \\
\# BitMEX (Derivative)           & XBTF24             & \$44,599.50                        & 44016.02 USD                             & -1.308\%                           & 0.18\%                              & 19 days                                 & \$3,686,400                                & \$42,073.68                             \\
\# BitMEX (Derivative)           & XBTU24             & \$47,680.00                        & 44016.02 USD                             & -7.685\%                           & 0.82\%                              & 9 months                                & \$874,700                                  & \$14,332.62                             \\
\# BitMEX (Derivative)           & XBTG24             & \$45,034.50                        & 44016.02 USD                             & -2.262\%                           & 0.2\%                               & about 2 months                          & \$149,400                                  & \$502.57                                \\
\# BitMEX (Derivative)           & XBTUSDTH24         & \$44,913.12                        & 43991.27 USDT                            & -2.003\%                           & 22.08\%                             & 3 months                                & \$67,027                                   & \$23,309.91                             \\
\# BTSE (Futures)                & BTC-240329         & \$45,346.20                        & 45337.26 USD                             & -0.020\%                           & 0.15\%                              & 3 months                                & \$70,779,838                               & \$70,829.75                             \\
\# BTSE (Futures)                & BTC-240628         & \$46,578.30                        & 46774.54 USD                             & +0.421\%                           & 0.2\%                               & 6 months                                & \$68,641,915                               & \$58,755.91                             \\
\# Bybit (Futures)               & BTCUSDH24          & \$45,515.00                        & 44050.12 USD                             & -3.218\%                           & 0.01\%                              & 3 months                                & \$31,042,370                               & \$8,992,250.00                          \\
\# Bybit (Futures)               & BTCUSDM24          & \$46,867.50                        & 44046.87 USD                             & -6.018\%                           & 0.01\%                              & 6 months                                & \$13,200,403                               & \$932,634.00                            \\
\# Bybit (Futures)               & BTC-29MAR24        & \$45,288.92                        & 44042.99 USDC                            & -2.786\%                           & 0.07\%                              & 3 months                                & -                                          & \$1,650,409.10                          \\
\# Bybit (Futures)               & BTC-28JUN24        & \$46,655.43                        & 44042.83 USDC                            & -5.633\%                           & 0.07\%                              & 6 months                                & -                                          & \$636,256.66                            \\
\# Bybit (Futures)               & BTC-12JAN24        & \$44,205.30                        & 44042.83 USDC                            & -0.403\%                           & 0.02\%                              & 5 days                                  & -                                          & \$320,758.50                            \\
\# Bybit (Futures)               & BTC-19JAN24        & \$44,341.25                        & 44042.83 USDC                            & -0.708\%                           & 0.03\%                              & 12 days                                 & -                                          & \$117,823.82                            \\
\# Bybit (Futures)               & BTC-27SEP24        & \$47,694.06                        & 44042.99 USDC                            & -7.688\%                           & 0.08\%                              & 9 months                                & -                                          & \$66,406.59                             \\
\# Bybit (Futures)               & BTC-26JAN24        & \$44,370.24                        & 44042.75 USDC                            & -0.773\%                           & 0.02\%                              & 19 days                                 & -                                          & \$47,245.15                             \\
\# Bybit (Futures)               & BTC-23FEB24        & \$44,806.59                        & 44042.99 USDC                            & -1.739\%                           & 0.03\%                              & about 2 months                          & -                                          & \$8,650.36                              \\
\# Crypto.com Exchange (Futures) & BTCUSD-240628      & \$46,384.20                        & 44003.38 USD                             & -5.133\%                           & 0.41\%                              & -                                       & \$24,557,224                               & \$15,659.31                             \\
\# Crypto.com Exchange (Futures) & BTCUSD-240329      & \$45,024.60                        & 44003.38 USD                             & -2.268\%                           & 0.46\%                              & -                                       & \$1,118,289                                & \$17,244.42                             \\
\# Crypto.com Exchange (Futures) & BTCUSD-240126      & \$44,215.10                        & 44003.38 USD                             & -0.479\%                           & 0.29\%                              & -                                       & \$9,701                                    & \$18,428.85                             \\
\# Crypto.com Exchange (Futures) & BTCUSD-240223      & \$44,624.00                        & 44003.38 USD                             & -1.391\%                           & 0.43\%                              & -                                       & \$3,539                                    & \$14,931.19                             \\
\# Deribit                       & BTC-29MAR24        & \$45,300.00                        & 44026.9 USD                              & -2.810\%                           & 0.02\%                              & 3 months                                & \$208,084,190                              & \$15,282,281.56                         \\
\# Deribit                       & BTC-26JAN24        & \$44,475.00                        & 44026.9 USD                              & -1.008\%                           & 0.1\%                               & 19 days                                 & \$89,621,140                               & \$3,132,087.87                          \\
\# Deribit                       & BTC-28JUN24        & \$46,577.50                        & 44026.9 USD                              & -5.476\%                           & 0.02\%                              & 6 months                                & \$79,314,230                               & \$7,916,251.72                          \\
\# Deribit                       & BTC-27SEP24        & \$47,715.00                        & 44026.9 USD                              & -7.729\%                           & 0.09\%                              & 9 months                                & \$32,099,570                               & \$1,034,939.35                          \\
\# Deribit                       & BTC-12JAN24        & \$44,110.00                        & 44026.9 USD                              & -0.188\%                           & 0.03\%                              & 5 days                                  & \$25,884,930                               & \$7,230,738.10                          \\
\# Deribit                       & BTC-23FEB24        & \$44,767.50                        & 44026.9 USD                              & -1.654\%                           & 0.03\%                              & about 2 months                          & \$22,571,940                               & \$744,850.73                            \\
\# Deribit                       & BTC-27DEC24        & \$49,247.50                        & 44026.9 USD                              & -10.601\%                          & 0.13\%                              & 12 months                               & \$11,398,600                               & \$600,503.46                            \\
\# Deribit                       & BTC-19JAN24        & \$44,222.50                        & 44026.9 USD                              & -0.442\%                           & 0.04\%                              & 12 days                                 & \$5,538,440                                & \$2,799,269.71                          \\
\# HTX Futures                   & BTC\_CQ            & \$45,297.68                        & 44023.11 USD                             & -2.814\%                           & 0.38\%                              & 3 months                                & \$1,002,200                                & \$1,602,716.03                          \\
\# HTX Futures                   & BTC\_CW            & \$44,095.65                        & 44030.71 USD                             & -0.147\%                           & 0.45\%                              & 4 days                                  & \$47,100                                   & \$1,011,419.90                          \\
\# HTX Futures                   & BTC\_NW            & \$44,211.45                        & 44023.11 USD                             & -0.426\%                           & 0.4\%                               & 11 days                                 & \$9,000                                    & \$967,082.71                            \\
\# Kraken (Futures)              & FI\_XBTUSD\_240329 & \$45,520.50                        & 44013.38 USD                             & -3.311\%                           & 0.08\%                              & 3 months                                & \$6,608,999                                & \$112,442.00                            \\
\# Kraken (Futures)              & FI\_XBTUSD\_240628 & \$46,571.50                        & 44013.38 USD                             & -5.493\%                           & 0.24\%                              & 6 months                                & \$1,929,982                                & \$275,529.00                            \\
\# Kraken (Futures)              & FI\_XBTUSD\_240126 & \$44,191.50                        & 44013.38 USD                             & -0.403\%                           & 0.4\%                               & 19 days                                 & \$300,856                                  & \$3.00                                  \\
\# Kraken (Futures)              & FF\_XBTUSD\_240329 & \$45,284.00                        & 44013.38 USD                             & -2.806\%                           & 0.11\%                              & 3 months                                & \$87                                       & \$22.68                                 \\
\# Kraken (Futures)              & FF\_XBTUSD\_240126 & \$44,307.00                        & 44013.38 USD                             & -0.663\%                           & 0.08\%                              & 19 days                                 & \$32                                       & \$4.61                                  \\
\# KuCoin Futures                & XBTMH24            & \$45,444.00                        & 44072.8 USD                              & -3.017\%                           & 0.19\%                              & 3 months                                & \$11,251,580                               & \$86,193.00                             \\
\# OKX (Futures)                 & BTC-USD-240329     & \$45,322.90                        & 44016.1 USD                              & -2.883\%                           & 0.01\%                              & 3 months                                & \$175,555,100                              & \$42,597,534.31                         \\
\# OKX (Futures)                 & BTC-USD-240628     & \$46,744.90                        & 44016.1 USD                              & -5.838\%                           & 0.03\%                              & 6 months                                & \$67,199,400                               & \$13,281,039.45                         \\
\# OKX (Futures)                 & BTC-USDT-240329    & \$45,345.34                        & 43989.3 USDT                             & -2.942\%                           & 0.02\%                              & 3 months                                & \$54,809,841                               & \$32,494,014.76                         \\
\# OKX (Futures)                 & BTC-USDT-240628    & \$46,824.58                        & 43989.3 USDT                             & -6.008\%                           & 0.02\%                              & 6 months                                & \$49,096,426                               & \$8,216,309.33                          \\
\# OKX (Futures)                 & BTC-USD-240126     & \$44,235.20                        & 44016.1 USD                              & -0.495\%                           & 0.04\%                              & 19 days                                 & \$20,945,300                               & \$1,424,116.88                          \\
\# OKX (Futures)                 & BTC-USD-240112     & \$44,085.20                        & 44016.1 USD                              & -0.157\%                           & 0.01\%                              & 5 days                                  & \$15,394,800                               & \$11,575,786.01                         \\
\# OKX (Futures)                 & BTC-USDT-240126    & \$44,284.50                        & 43989.3 USDT                             & -0.617\%                           & 0.06\%                              & 19 days                                 & \$8,573,427                                & \$1,552,614.65                          \\
\# OKX (Futures)                 & BTC-USD-240223     & \$44,719.10                        & 44016.1 USD                              & -1.572\%                           & 0.04\%                              & about 2 months                          & \$8,046,700                                & \$3,603,630.54                          \\
\# OKX (Futures)                 & BTC-USD-240119     & \$44,198.90                        & 44016.1 USD                              & -0.414\%                           & 0.01\%                              & 12 days                                 & \$4,957,200                                & \$6,247,231.64                          \\
\# OKX (Futures)                 & BTC-USDT-240112    & \$44,080.50                        & 43989.3 USDT                             & -0.157\%                           & 0.03\%                              & 5 days                                  & \$3,560,036                                & \$12,798,332.26                         \\
\# OKX (Futures)                 & BTC-USDT-240223    & \$44,781.65                        & 43989.3 USDT                             & -1.720\%                           & 0.08\%                              & about 2 months                          & \$2,960,896                                & \$711,132.65                            \\
\# OKX (Futures)                 & BTC-USDT-240119    & \$44,192.36                        & 43989.3 USDT                             & -0.409\%                           & 0.04\%                              & 12 days                                 & \$975,364                                  & \$4,235,395.39                          \\
\# OKX (Futures)                 & BTC-USDC-240628    & \$46,788.02                        & 44015.4 USDC                             & -5.958\%                           & 0.39\%                              & 6 months                                & -                                          & \$17,320.92                             \\
\# OKX (Futures)                 & BTC-USDC-240119    & \$44,137.13                        & 44020.7 USDC                             & -0.298\%                           & 0.1\%                               & 12 days                                 & -                                          & \$6,285.13                              \\
\# OKX (Futures)                 & BTC-USDC-240329    & \$45,832.15                        & 44020.7 USDC                             & -3.985\%                           & 0.31\%                              & 3 months                                & -                                          & \$1,874.53                              \\
\# OKX (Futures)                 & BTC-USDC-240112    & \$44,107.04                        & 44015.4 USDC                             & -0.242\%                           & 0.13\%                              & 5 days                                  & -                                          & \$9,765.30                              \\
\# XT.COM (Derivatives)          & BTC\_USD\_240329   & \$45,380.70                        & -                                        & -3.586\%                           & 0.01\%                              & 3 months                                & -                                          & \$33,113,000.00                         \\
\# XT.COM (Derivatives)          & BTC\_USD\_240628   & \$46,797.20                        & -                                        & -                                  & 0.01\%                              & 6 months                                & -                                          & \$31,000,200.00                         \\
\# XT.COM (Derivatives)          & BTC\_USDT\_240329  & \$45,369.50                        & -                                        & -                                  & 0.06\%                              & 3 months                                & -                                          & \$20,870,477.00                        
\end{tabular}
\end{table}

ETH Futures

with 
Total Open Interest $466,932,584
Total 24h Trade Volume $98,238,298

\begin{table}[]
\begin{tabular}{llrrrrrrr}
\textbf{Exchange}                & \textbf{Symbol}    & \multicolumn{1}{l}{\textbf{Price}} & \multicolumn{1}{l}{\textbf{Index Price}} & \multicolumn{1}{l}{\textbf{Basis}} & \multicolumn{1}{l}{\textbf{Spread}} & \multicolumn{1}{l}{\textbf{Expires in}} & \multicolumn{1}{l}{\textbf{Open Interest}} & \multicolumn{1}{l}{\textbf{24h Volume}} \\
\# Binance (Futures)             & ETHUSD\_240329     & \$2,300.60                         & 2228.81 USD                              & -3.121\%                           & 0.01\%                              & 3 months                                & \$72,215,240                               & \$16,326,927.83                         \\
\# Binance (Futures)             & ETHUSD\_240628     & \$2,368.79                         & 2229.03 USD                              & -5.900\%                           & 0.02\%                              & 6 months                                & \$11,183,360                               & \$4,224,925.26                          \\
\# Binance (Futures)             & ETHUSDT\_240329    & \$2,300.46                         & 2227.81 USDT                             & -3.110\%                           & 0.02\%                              & -                                       & \$35,040                                   & \$6,902,127.72                          \\
\# Binance (Futures)             & ETHUSDT\_240628    & \$2,360.43                         & 2227.81 USDT                             & -5.572\%                           & 0.03\%                              & -                                       & \$4,604                                    & \$4,303,853.27                          \\
\# BitMEX (Derivative)           & ETHUSDH24          & \$2,452.60                         & 2228.01 USD                              & -9.157\%                           & 0.02\%                              & 3 months                                & \$4,341,286                                & \$537,210.05                            \\
\# BitMEX (Derivative)           & ETHH24             & \$2,217.23                         & 0.05 XBT                                 & +0.457\%                           & 0.37\%                              & 3 months                                & \$1,575,478                                & \$69,244.08                             \\
\# BTSE (Futures)                & ETH-240628         & \$2,350.05                         & 2367.26 USD                              & +0.732\%                           & 0.75\%                              & 6 months                                & \$2,517,764                                & \$49,845.75                             \\
\# BTSE (Futures)                & ETH-240329         & \$2,297.48                         & 2298.78 USD                              & +0.057\%                           & 0.23\%                              & 3 months                                & \$2,081,180                                & \$49,130.96                             \\
\# Bybit (Futures)               & ETHUSDM24          & \$2,357.45                         & 2229.63 USD                              & -5.422\%                           & 0.05\%                              & 6 months                                & -                                          & \$448,716.00                            \\
\# Bybit (Futures)               & ETHUSDH24          & \$2,287.90                         & 2229.53 USD                              & -2.551\%                           & 0.06\%                              & 3 months                                & -                                          & \$294,007.00                            \\
\# Bybit (Futures)               & ETH-26JAN24        & \$2,243.65                         & 2229.46 USDC                             & -0.668\%                           & 0.05\%                              & 19 days                                 & -                                          & \$286,245.77                            \\
\# Bybit (Futures)               & ETH-12JAN24        & \$2,231.66                         & 2229.51 USDC                             & -0.132\%                           & 0.03\%                              & 5 days                                  & -                                          & \$99,683.25                             \\
\# Bybit (Futures)               & ETH-28JUN24        & \$2,368.66                         & 2229.46 USDC                             & -5.910\%                           & 0.1\%                               & 6 months                                & -                                          & \$97,357.35                             \\
\# Bybit (Futures)               & ETH-19JAN24        & \$2,238.21                         & 2229.51 USDC                             & -0.424\%                           & 0.06\%                              & 12 days                                 & -                                          & \$92,957.40                             \\
\# Bybit (Futures)               & ETH-23FEB24        & \$2,274.64                         & 2229.51 USDC                             & -2.019\%                           & 0.03\%                              & about 2 months                          & -                                          & \$68,610.65                             \\
\# Bybit (Futures)               & ETH-29MAR24        & \$2,297.48                         & 2229.51 USDC                             & -2.993\%                           & 0.03\%                              & 3 months                                & -                                          & \$36,926.81                             \\
\# Bybit (Futures)               & ETH-27SEP24        & \$2,420.24                         & 2229.46 USDC                             & -7.915\%                           & 0.12\%                              & 9 months                                & -                                          & \$30,008.99                             \\
\# Crypto.com Exchange (Futures) & ETHUSD-240628      & \$2,347.68                         & 2224.96 USD                              & -5.227\%                           & 0.24\%                              & -                                       & \$24,573,170                               & \$822.63                                \\
\# Crypto.com Exchange (Futures) & ETHUSD-240126      & \$2,234.13                         & 2224.96 USD                              & -0.410\%                           & 0.14\%                              & -                                       & \$79,732                                   & \$742.18                                \\
\# Crypto.com Exchange (Futures) & ETHUSD-240329      & \$2,286.18                         & 2224.96 USD                              & -2.678\%                           & 0.29\%                              & -                                       & \$58,845                                   & \$895.27                                \\
\# Crypto.com Exchange (Futures) & ETHUSD-240223      & \$2,249.95                         & 2224.92 USD                              & -1.112\%                           & 0.17\%                              & -                                       & \$27,557                                   & \$1,247.37                              \\
\# Deribit                       & ETH-29MAR24        & \$2,296.00                         & 2228.59 USD                              & -2.936\%                           & 0.02\%                              & 3 months                                & \$88,364,731                               & \$2,438,537.07                          \\
\# Deribit                       & ETH-28JUN24        & \$2,363.75                         & 2228.59 USD                              & -5.718\%                           & 0.01\%                              & 6 months                                & \$35,776,949                               & \$1,152,627.40                          \\
\# Deribit                       & ETH-26JAN24        & \$2,247.75                         & 2228.59 USD                              & -0.852\%                           & 0.02\%                              & 19 days                                 & \$34,172,788                               & \$2,694,255.34                          \\
\# Deribit                       & ETH-27SEP24        & \$2,414.50                         & 2228.59 USD                              & -7.700\%                           & 0.02\%                              & 9 months                                & \$22,332,082                               & \$781,745.26                            \\
\# Deribit                       & ETH-12JAN24        & \$2,232.25                         & 2228.59 USD                              & -0.164\%                           & 0.01\%                              & 5 days                                  & \$10,281,174                               & \$1,626,945.44                          \\
\# Deribit                       & ETH-23FEB24        & \$2,273.25                         & 2228.59 USD                              & -1.965\%                           & 0.08\%                              & about 2 months                          & \$10,187,446                               & \$484,690.68                            \\
\# Deribit                       & ETH-27DEC24        & \$2,463.00                         & 2228.59 USD                              & -9.517\%                           & 0.21\%                              & 12 months                               & \$6,513,788                                & \$630,519.38                            \\
\# Deribit                       & ETH-19JAN24        & \$2,235.50                         & 2228.59 USD                              & -0.309\%                           & 0.03\%                              & 12 days                                 & \$1,564,812                                & \$1,319,273.13                          \\
\# HTX Futures                   & ETH\_CQ            & \$2,298.00                         & 2228.66 USD                              & -3.018\%                           & 0.01\%                              & 3 months                                & \$1,130,370                                & \$562,732.95                            \\
\# HTX Futures                   & ETH\_CW            & \$2,233.11                         & 2228.54 USD                              & -0.205\%                           & 0.01\%                              & 4 days                                  & \$49,870                                   & \$481,669.51                            \\
\# HTX Futures                   & ETH\_NW            & \$2,242.45                         & 2228.66 USD                              & -0.615\%                           & 0.35\%                              & 11 days                                 & \$31,200                                   & \$302,572.28                            \\
\# Kraken (Futures)              & FI\_ETHUSD\_240329 & \$2,300.00                         & 2228 USD                                 & -3.130\%                           & 0.11\%                              & 3 months                                & \$4,415,821                                & \$29,062.00                             \\
\# Kraken (Futures)              & FI\_ETHUSD\_240628 & \$2,366.30                         & 2228 USD                                 & -5.845\%                           & 0.25\%                              & 6 months                                & \$3,541,749                                & \$40,752.00                             \\
\# Kraken (Futures)              & FF\_ETHUSD\_240329 & \$2,314.00                         & 2228 USD                                 & -3.717\%                           & 0.13\%                              & 3 months                                & \$1,584                                    & \$0.45                                  \\
\# Kraken (Futures)              & FF\_ETHUSD\_240126 & \$2,254.90                         & 2228 USD                                 & -1.193\%                           & 0.09\%                              & 19 days                                 & \$299                                      & \$36.69                                 \\
\# Kraken (Futures)              & FI\_ETHUSD\_240126 & \$2,254.15                         & 2228 USD                                 & -1.160\%                           & 0.24\%                              & 19 days                                 & \$48,594                                   & \$10,184.00                             \\
\# OKX (Futures)                 & ETH-USD-240329     & \$2,296.78                         & 2228.71 USD                              & -2.964\%                           & 0.01\%                              & 3 months                                & \$61,638,000                               & \$13,362,284.32                         \\
\# OKX (Futures)                 & ETH-USD-240628     & \$2,370.47                         & 2228.67 USD                              & -5.982\%                           & 0.03\%                              & 6 months                                & \$29,831,250                               & \$8,793,848.00                          \\
\# OKX (Futures)                 & ETH-USDT-240329    & \$2,296.43                         & 2226.75 USDT                             & -2.985\%                           & 0.02\%                              & 3 months                                & \$20,412,503                               & \$5,701,567.75                          \\
\# OKX (Futures)                 & ETH-USD-240112     & \$2,231.34                         & 2228.71 USD                              & -0.118\%                           & 0.01\%                              & 5 days                                  & \$7,456,500                                & \$8,340,344.15                          \\
\# OKX (Futures)                 & ETH-USDT-240628    & \$2,366.59                         & 2227.15 USDT                             & -5.845\%                           & 0.02\%                              & 6 months                                & \$3,885,976                                & \$3,157,980.17                          \\
\# OKX (Futures)                 & ETH-USD-240119     & \$2,233.42                         & 2228.71 USD                              & -0.211\%                           & 0.03\%                              & 12 days                                 & \$3,212,320                                & \$3,467,988.47                          \\
\# OKX (Futures)                 & ETH-USDT-240112    & \$2,229.86                         & 2226.75 USDT                             & -0.089\%                           & 0.01\%                              & 5 days                                  & \$2,298,019                                & \$3,571,348.57                          \\
\# OKX (Futures)                 & ETH-USDT-240119    & \$2,235.50                         & 2226.75 USDT                             & -0.341\%                           & 0.01\%                              & 12 days                                 & \$690,474                                  & \$2,364,707.49                          \\
\# OKX (Futures)                 & ETH-USDC-240329    & \$2,296.52                         & 2227.93 USDC                             & -3.020\%                           & 0.38\%                              & 3 months                                & -                                          & \$55,373.71                             \\
\# OKX (Futures)                 & ETH-USDC-240112    & \$2,251.63                         & 2227.89 USDC                             & -1.088\%                           & 0.09\%                              & 5 days                                  & -                                          & \$5,770.92                              \\
\# OKX (Futures)                 & ETH-USDC-240119    & \$2,251.87                         & 2227.93 USDC                             & -1.037\%                           & 0.14\%                              & 12 days                                 & -                                          & \$9.01                                  \\
\# XT.COM (Derivatives)          & ETH\_USD\_240628   & \$2,368.07                         & -                                        & -                                  & 0.04\%                              & 6 months                                & -                                          & \$2,929,170.00                          \\
\# XT.COM (Derivatives)          & ETH\_USD\_240329   & \$2,299.63                         & -                                        & -2.350\%                           & 0.01\%                              & 3 months                                & -                                          & \$2,698,030.00                          \\
\# XT.COM (Derivatives)          & ETH\_USDT\_240329  & \$2,301.93                         & -                                        & -                                  & 0.03\%                              & 3 months                                & -                                          & \$124,645.12                            \\
\# OKX (Futures)                 & BTC-USDT-240112    & \$44,080.50                        & -0.157\%                                 & 0.03\%                             & 5 days                              & \$3,560,036                             & \$12,798,332.26                            & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDT-240223    & \$44,781.65                        & -1.720\%                                 & 0.08\%                             & about 2 months                      & \$2,960,896                             & \$711,132.65                               & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDT-240119    & \$44,192.36                        & -0.409\%                                 & 0.04\%                             & 12 days                             & \$975,364                               & \$4,235,395.39                             & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDC-240628    & \$46,788.02                        & -5.958\%                                 & 0.39\%                             & 6 months                            & -                                       & \$17,320.92                                & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDC-240119    & \$44,137.13                        & -0.298\%                                 & 0.1\%                              & 12 days                             & -                                       & \$6,285.13                                 & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDC-240329    & \$45,832.15                        & -3.985\%                                 & 0.31\%                             & 3 months                            & -                                       & \$1,874.53                                 & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDC-240112    & \$44,107.04                        & -0.242\%                                 & 0.13\%                             & 5 days                              & -                                       & \$9,765.30                                 & \multicolumn{1}{l}{}                    \\
\# XT.COM (Derivatives)          & BTC\_USD\_240329   & \$45,380.70                        & -3.586\%                                 & 0.01\%                             & 3 months                            & -                                       & \$33,113,000.00                            & \multicolumn{1}{l}{}                    \\
\# XT.COM (Derivatives)          & BTC\_USD\_240628   & \$46,797.20                        & -                                        & 0.01\%                             & 6 months                            & -                                       & \$31,000,200.00                            & \multicolumn{1}{l}{}                    \\
\# XT.COM (Derivatives)          & BTC\_USDT\_240329  & \$45,369.50                        & -                                        & 0.06\%                             & 3 months                            & -                                       & \$20,870,477.00                            & \multicolumn{1}{l}{}                   
\end{tabular}
\end{table}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}

Crypto exchanges often provide a so-called perpetual future, which is a synthetic (typically cash-settled instrument) whose return is tied to the underlying via a funding rate. Most exchanges (quote here) provide perpetual futures for Bitcoin and Ethereum. 
The price of each perpetual future is linked to the underlying by adjusting a variable interest rate, called the funding rate, that fluctuates depending on the deviation between the price of the perpetual future and the underlying. The funding rate increases the cost of holding positions against the deviation, meaning that if the price of the perpetual future is higher than the underlying, the funding rate is positive which causes the market participants who are long to pay the short side in a contractually predefined frequency. Broadly said, the more extreme the deviation, the higher the funding rate and thus the higher the incentive to trade against the deviation and thus converge the price of the perpetual future towards its underlying. 
The risk-free rate is approximated using the simplest and most liquid way of realising the funding rate, a carry trade. In such a trade, the trader shorts (longs) the perpetual future if the funding rate is positive (negative) and simulateously buys (sells) the underlying such that the market delta of the combined position is zero.

The funding rate on \href{https://www.deribit.com/kb/deribit-linear-perpetual}{Deribit} is calculated as follows:

Deribit perpetual contracts are a derivative product similar to a future, however, they do not have an expiry date. The perpetual contract features funding payments. These payments have been introduced to keep the perpetual contract price as close as possible to the underlying crypto price - the relevant Deribit Index. If the perpetual contract trades at a higher price than the index, traders that have long positions need to make funding payments to the traders that have short positions. This will make the product less attractive to the long position holders and more attractive to the short position holders. This subsequently will cause a perpetual price to trade in line with the price of the index. If the perpetual trades at a price lower than the index, the short position holders will have to pay the long position holders instead.

The Deribit perpetual contract features a continuous measurement of the difference between the mark price of the contract and the relevant Deribit Index. The percentage difference between these two price levels is the basis for the 8-hourly funding rate that is applied to all outstanding perpetual contracts.

Funding payments are calculated every millisecond. The funding payments will be added to or subtracted from the realized PNL, which is also part of the available trading balance. At each daily settlement, the realized PNL will be moved to or from the cash balance, from which withdrawals can be made.

The total funding paid will show up in the transaction log in the "funding" column. This column shows the funding amount that is applied to the trader's entire net position in the period between the relevant trade and the trade before that. Put differently: the trader can see the funding paid or received on the position between position changes.

Funding Rate
When the funding rate is positive, long position holders pay funding to the short position holders; when the funding rate is negative, short position holders pay funding to the long position holders. The funding rate is expressed as an 8-hour interest rate, and is calculated at any given time as follows:

Premium Rate
Premium Rate = ((Mark Price - Deribit Index) / Deribit Index) * 100%

Funding Rate
Sequentially, the funding rate is derived from the premium rate by applying a damper.

If the premium rate is within -0.025\% and 0.025\% range, the actual funding rate will be reduced to 0.00\%.
If the premium rate is lower than -0.025\%, then the actual funding rate will be the premium rate + 0.025\%.
If the premium rate is higher than 0.025\%, then the actual funding rate will be the premium rate - 0.025\%.
Additionally, the funding rate is capped at +/ - 5\% for all linear USDC perpetuals.
Funding Rate = Maximum (0.025\%, Premium Rate) + Minimum (-0.025\%, Premium Rate)

Time Fraction
Time Fraction = Funding Rate Time Period / 8 hours

The actual funding payment is calculated by multiplying the funding rate by the position size and the time fraction.

Funding Payment = Funding Rate * Position Size * Time Fraction

In reality, the spread of the Deribit Index and the mark price changes continuously, and all changes are taken into account. Therefore, the examples above are extreme simplifications of the actual calculations. The funding paid or received is continuously added to the realized PNL and is moved to or from the cash balance at the daily settlement, at 08:00 UTC.

Fees on Funding
Deribit does not charge any fees on funding. All the funding payments are transferred between the holders of the perpetual contracts. This makes the funding a zero-sum game, where longs receive all funding from shorts, or shorts receive all funding from longs.

Mark Price
It is essential to understand how the mark price is calculated. We start by determining the "Fair Price". The fair price is calculated as the average of the fair impact bid and the fair impact ask.

Fair Impact Bid Price is an average price at which 100,000 USDC worth of contracts can be bought.

Fair Impact Ask Price is an average price at which 100,000 USDC worth of contracts can be sold.

Fair Price can never be higher than 0.1 % of best ask or lower than best bid

Fair Price = (Fair Impact Bid + Fair Impact Ask) / 2
The mark price is derived using both the Deribit Index and the fair price, by adding to the Deribit Index the 30 second exponential moving average (EMA) of the Fair Price - Deribit Index.

Mark Price = Deribit Index + 30 second EMA (Fair Price - Deribit Index)
We implement a safety setting on our mark price to not deviate by more than a certain percentage of Index. This is to prevent manipulation or to curb the impact of panic trading. When a deviation from index is genuine, then we would increase the possible deviation.

Trading outside of this bandwidth is still allowed.

The 30 second EMA is recalculated every second, so in total, there are 30 time periods in which the measurement of the latest second has a weight of 2 / (30 + 1) = 0.0645 or (6.45%).

Allowed Trading Bandwidth
Two parameters act as a boundary to the allowed trading range:

Perpetual trades are limited by Deribit Index + 1 minute EMA (Fair Price - Index) +/- 1.5%, and a fixed bandwidth of the Deribit Index of +/- 7.5%.

If the market circumstances require so, bandwidth parameters can be adjusted at the sole discretion of Deribit.

Insert Plot Funding Rate!!

\begin{equation}
C_t = e^{-r \tau} \mathbb{\sf{E}}^Q[\psi(S_T)] = e^{-r, \tau}  \int_{-\infty}^\infty \psi(u) f_t^Q(u) du
\end{equation}

where the PK \(K(\cdot)\) is defined as \(\frac{q(\cdot)}{p(\cdot)}\)
and for simplicity of notation we write \(q(\cdot)\) for
\(\frac{\partial Q}{\partial t}(\cdot)\) and drop the sequence \(r_t\).

\section{Nonparametric Estimation of State Price Densities}

\subsection{Derivation}

\subsection{Local Polynomial Estimation of the IV Surface}

Assume that the implied volatilites have some noise added (Huynh et al.
2002; Rookley 1997).

\begin{equation}
\sigma (M, \tau) = g(M, \tau) + \sigma^{*}(M, \tau)  \varepsilon
\end{equation}

with a standardized error random variable \(\varepsilon\), moneyness
\(M\), \(\tau\) and \(\varepsilon\) independent and
\(\sigma^{*}(M, \tau)\) being the scaling of the error term given \(M\)
and \(\tau\). Suppose \(g\) is smooth, i.e.~it can be approximated using
Taylor's Theorem.

Taylor expansion of \(g\) in a neighborhood of \((M_0, \tau_0)\):
\begin{equation}
\begin{aligned}
g(M, \tau) = g(M_0, \tau_0) +
\frac{\partial g}{\partial M}(M-M_0) + 
\frac{1}{2} \frac{\partial^2 g}{\partial M^2}(M-M_0)^2 + \\
\frac{\partial g}{\partial \tau}(\tau - \tau_0) + 
\frac{1}{2} \frac{\partial^2 g}{\partial \tau^2}(\tau - \tau_0)^2 + \\ 
\frac{\partial^2 g}{\partial M \partial \tau}(M-M_0)(\tau - \tau_0)
\end{aligned}
\end{equation}

The functional relationship between the IV surface \(\sigma\) and \(M\)
and \(\tau\) can now be approximated using a Weighted Least Squares
Estimator (WLSE), minimizing the objective function

\begin{equation}
\operatorname{arg}\, \underset{\beta}{\operatorname{min}}{(\sigma - X\beta)^\top W (\sigma - X\beta)}
\end{equation}

where
\(W = \text{diag}(\mathcal{K}_{h_m, h_\tau}(M_j - M_0, \tau_j - \tau_0))\)
for a Gaussian kernel \(\mathcal{K}\) with bandwidths \(h_m\) and
\(h_\tau\). \(\sigma\) is the \(n \times 1\) vector of observed implied
volatilities, \(\beta\) is the vector of the local polynomial
coefficients.

\[
X = 
\begin{pmatrix}
1 & (M_1 - M_0) & (M_1 - M_0)^2 & (\tau_1 - \tau_0) & (\tau_1 - \tau_0)^2 & (M_1 - M_0) (\tau_1 - \tau_0) \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \\
1 & (M_n - M_0) & (M_n - M_0)^2 & (\tau_n - \tau_0) & (\tau_n - \tau_0)^2 & (M_n - M_0) (\tau_n - \tau_0) \\
\end{pmatrix}
\]

The resulting WLSE is

\begin{equation}
\hat{\beta} = (X ^\top WX)^{-1}X ^\top W \sigma
\end{equation}

Following Härdle et al. (2014), a window of the last 500 daily returns
(based on \(\tau\)) is used to calculate a nonparametric Kernel Density
Estimator for the PD.\\
\href{https://quantinar.ro/course/103/statistics-of-financial-markets}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/quantinar_logo.png}}\,Chapter 24}

\section{Literature Review}

Breeden and Litzenberger (1978) derive SPDs using Arrow-Debreu prices
and Butterfly Spreads. Their paper was the cornerstone for the now
existing vast literature on estimation of SPDs. Without requiring a
parametric form for SPDs, Rookley (1997) who developed a nonparametric
estimation method. IV skews are estimated by decomposing the functional
relationship between IV, moneyness and time-to-maturity \(\tau\). In
this manner it is possible to derive the SPD at every point in a robust
way. Aı̈t-Sahalia et al. (2001) estimate PKs from S\&P500 options data
and the according return series in order to assess the efficiency of the
options market. Departures from SPD and PD are used to identify
inefficient pricing. They design a trading strategy, exploiting the
skewness and kurtosis of the densities. The strategy is shown to have a
high Sharpe Ratio. Grith et al. (2009) propose a systematic modeling
approach to study the evolution of PKs over time. With European DAX
data, a series of empirical PKs is estimated from 2003 until 2006. While
the risk-neutral density is inferred using Rookley's method, the PD is
estimated with a GARCH model. A common shape is identified and
deviations of the time varying EPKs are studied. The deviations between
individual PKs are described using a set of parameters for horizontal
and vertical shifts. The relationship between PKs and Arrow-Pratt
measure of Absolute Risk Aversion (ARA) give insight into investor's
risk aversion. In a related research, Härdle et al. (2014) derive
bootstrap-based confidence bands for nonparametrically estimated PKs.

Inverse Options, meaning options settled in kind of the underlying, are
dominant in the crypto world. Settlement of BTC options in terms of BTC
changes the payoff function of the option from \(\max(S_T - K; 0)\) to
\(\max(\frac{S_T - K}{S_T}; 0)\) and thus changes the contract's net
present value. Alexander and Imeraj (2021) adjust Black-Scholes prices
and hedge ratios to inverse options. However, it is argued that traders
are erroneously applying vanilla Black-Scholes valuation instead of the
corrected prices - perhaps because of being unaware of the concept. We
have decided to use with the standard, non-inverse Black Scholes pricing
for multiple reasons. Alexander and Imeraj (2021) argue that traders are
perhaps unaware of the difference. Since we are inferring investor
expectations, we figure that it is more appropriate to infer pricing
kernels under the same Black Scholes prices that the investors use.
Additionally, it is likely that many traders actually hedge their
overall Bitcoin exposure. This would negate the ``inverse'' part of the
option. Furthermore, reviewing the differences in pricing compared to
the adjusted prices, we find that the difference is small in the absence
of extreme moves in the underlying. This was not the case in the time
period under consideration.

Hou et al. (2020) price BTC options under an SVCJ model. Their results
emphasize the tail risk introduced by jumps in the underlying. Jumps in
particular introduce market incompleteness, but anonymity of
transactions may be relevant as well. Chen and Vinogradov (2021) derive
PKs and the impact of market incompleteness on risk premia. They state
that a key property of cryptocurrency valuation is the user's anonymity
(or pseudonymity), which is simultaneously the source of value and
incompleteness in the respective market. They argue that hedging an
anonymous transaction would require an identity disclosure. This
contradiction may introduce an effectively unhedgeable event.

\section{Deribit}

Deribit, a margin-trading platform for futures and options, has been
launched in June 2016 in the Netherlands and is currently incorporated
in Panama. As of time of the writing, Deribit is the largest BTC option
exchange (Skew.com 2022). For both types of derivatives, BTC and ETH are
the underlying as well as the currency in which settlement is conducted.
This allows to view them as inverse options.

\begin{center}
\begin{figure}
\includegraphics[width=0.9\textwidth]{/Users/julian/src/up/spd/plot/Skew_BTC_Open_Interest_cut.pdf}
\caption[Open Interest in BTC Options per Exchange.]%
{Open Interest in BTC Options per Exchange. Snapshot from \href{https://analytics.skew.com/dashboard/bitcoin-options}{skew.com} on 2022-01-16.  \color{bittersweet}{Deribit}, \color{ao}{LedgerX}, \color{bananamania}{CME}, \color{bluegray}{bit.com}. }
\label{fig: ivtest}
\end{figure}
\end{center}

\subsection{Bitcoin Option Contract Specifications}

Deribit offers cash-settled European-style options. The underlying is a
synthetic index, whose exact composition is described in the data
section. Settlement is first calculated in USD and then conducted in
kind of the underlying. E.g. a BTC call with a strike of \(10000\) and a
settlement value of \(12000\) at maturity will result in a cashflow of
\(2000\) USD, which is equivalent to \(\frac{1}{6}\) BTC (excluding
transaction cost). The settlement value is defined as the average of the
underlying BTC index for the last 30 minutes before settlement. Each
contract has a lot size of 1 BTC and is settled at 8am UTC on the
respective maturity date. Maturity dates have a daily, monthly,
quarterly and annual frequency. Deribit is the only derivatives exchange
to offer daily options. Instruments are trading 24/7 with a tick size of
0.0005 BTC. Due to the automatic usage of margin-trading, margin-based
liquidation of positions is possible. Initial margin and maintenance
margins are both zero for long positions. For short positions, initial
margin is required (Deribit 2022b). In case of liquidation, a penalty is
applied to the defaulting party, whose proceeds are paid into the
Deribit Insurance Fund (Deribit 2020a).

Trading and deliveries on Deribit are subject to fees (Deribit 2022c).
The applied fees generally vary depending on whether the order was a
liquidity maker or taker. However, for Bitcoin options they are equal.
Trading fees are 0.03\% of the underlying or 0.0003 BTC per options
contract. The fees constitute a maximum of 12.5\% of the contract's
value.

\section{Data}

\subsection{Data Structure}

The dataset consists of 8,444,664 order book snapshots, which have been
collected from Deribit in the time from 2021-04-01 until 2022-04-01. The
snapshots were captured via the Deribit API V2. The data base is
available on the
\href{https://blockchain-research-center.com}{Blockchain Research Center (BRC)}.
The full data set includes all parameters that the
\href{https://docs.deribit.com/#deribit-api-v2-1-1}{Deribit API V2}
returns at the time of collection under the methods

\begin{itemize}
\tightlist
\item
  \{\it public/get\_last\_trades\_by\_instrument\_and\_time\}
\item
  \{\it public/get\_order\_book\}
\end{itemize}

Most importantly the results include

\begin{itemize}
\tightlist
\item
  Timestamps
\item
  Greeks
\item
  Implied Volatility
\item
  Tick Direction
\item
  Order Type
\item
  Volume
\item
  Instrument Price
\item
  Strike
\item
  Spot
\end{itemize}

High-frequent order book changes and executed trades are captured for
options and futures whose underlying is BTC. Values of the underlying
synthetic BTC USD Index are also saved. The underlying is calculated as
an equally-weighted BTC/USD price of eleven major crypto exchanges,
namely Binance, Bitfinex, Bitstamp, Bittrex, Coinbase Pro, FTX, Gemini,
Huobi, Itbit, Kraken, LMAX Digital, OKEx. Individual feeds can be
excluded due to administrative decisions or in case if invalid data.
Remaining feeds are sorted, truncated around the median price and
weighted (Deribit 2020b).

\subsection{Data Integrity}

Data integrity plays a crucial role in crypto markets. Mark Carney, the
chair of the Financial Stability Board and head of the Bank of England,
warned that wash trading, pump and dumps and spoofing, known as outlawed
manipulation techniques in equity markets, are also present in crypto
and pose a risk to financial stability (Rodgers 2019). Wash trading is
used to increase trading volume and thus artificially increase demand. A
spoofer submits market non-bona fide price quotes in order to cause
artificial price volatility (Sar 2017). Pump \& Dump is a form of
securities fraud in which a group of traders rapidly and artificially
inflate a price in order to offload their previously acquired inventory.
Since Pump \& Dump schemes require low liquidity (in the underlying), we
can exclude this possibility for Bitcoin (La Morgia et al. 2020).

Aloosh and Li (2019) find evidence for exchange-driven inflated volume,
generated by wash trades, to market themselves under the guise of
liquidity. Cong et al. (2019) report that on average 70\% of volume on
decentralized exchanges is fake due to the use of wash trading.
According to Bitmex (2019), up to 95\% of trading volume on unregulated
exchanges is generated by wash trading. In order to ensure data
integrity, we exclusively use orderbooks instead of executed trades in
order to filter possible cases of wash trading.

To address the potential issue of spoofing, we adopt an approach related
to Tuccella et al. (2021). They attempt to identify spoofing on
cryptocurrency exchanges using a GRU model. Their predictors are a
function of the amount of cancelled orders relative to the cumulative
order size on each side of the order book. Since Bitcoin derivatives
order books do not have the same amount of depth as spot or futures
markets, we employ a stricter method and exclude all orders whose
lifetime does not exceed two seconds. We find that less than 5\% of
orders are filtered due to short order lifetime. We figure that
traditional spoofing techniques, based on the rapid submission and
cancellation of orders, are difficult to implement due to Deribit's rate
limits (Deribit 2023). Users with less than 1 million USD in
7-day-turnover can only post five matching engine requests. These
include request for a buy or sell order or a corresponding cancellation.
Although the limits for a market maker (who could be allowed to post 30
requests per second given a turnover of 25 million USD in 7 days) are
considerably higher, the market maker must still meet the quoting
requirements for a large range of instruments. This effectively
restricts the influence that single entities can have on certain
instruments via the rapid submission and cancellation of orders and
explains the rarity of such events in our dataset.

\subsection{Preprocessing}

Although market makers are obliged to quote most instruments for the
majority of the time, some may not be quoted at all or only be quoted at
the cost of a large spread (Deribit 2020c). For Far-Out-of-Money (FOTM)
contracts, when the minimum tick size exceeds the market IV, one can
observe bids for 0\% IV. Cases of quotes for 0\% IV and all duplicates
are excluded from the data set. Since the majority of option trading
volume concentrates on instruments with short or medium time-to-maturity
\(\tau\), we restrict \(\tau\) to be smaller or equal than one quarter
of a year. \(\tau\) is normalized on the span of a year. Instruments are
grouped according to their time-to-maturity, over a range of 0, 1, 2, 4,
8 weeks. We restrict moneyness \(M = \frac{K}{S}\) to the interval
\([0.7, 1.3]\) in order to exclude the influence of high-volatility
observations which could be unreliable (Grith et al. 2009).

Call options are exclusively used in order to estimate the IV surface in
Rookley's method. Put-Call-Parity ensures arbitrage-free call option
prices. However, we do not use Put-Call-Parity to increase our data
sample. Conversion of option prices via Put-Call-Parity could introduce
errors into our dataset due to market microstructure (MMN) noise. Our
high-frequent order book data is extracted via an API. Due to rate
limitations and a diverse amount of instruments, order book snapshots
cannot be taken simultaneously for all instruments. The data scraping
application successively iterates over all instruments and captures
changes in order books. While the resulting data base provides a clear
picture of order book data, it does not allow for a complete
reconstruction of all order books at all times. Without simultaneity,
volatility of the underlying and order competition within the spread
could change the prices of the corresponding instruments (required for
Put-Call-Parity) between the individual snapshots. A range of futures
may be available on Deribit, but only the perpetual futures contract is
actually liquid. The classic (non-perpetual) futures also price in the
settlement cost, which often causes them to trade in backwardation close
to maturity. But in spite of this, BTC options tend to be traded more
frequently when they are close to maturity. The combination of these
real-world limitations challenges the simple conversion via
Put-Call-Parity and could introduce errors into our dataset.
Furthermore, we do not see the necessity to use such a method as we have
sufficient call option data to compute the IV surface from which we
sample option prices.

\href{https://github.com/QuantLet/BitcoinOptions/blob/master/main.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,Preprocessing}

\section{Volatility}

In accordance with Masset (2011) and Eraker (2021), the list of
well-established stylised facts on volatility in traditional markets
include horizontal dependence of volatility, leverage effect, the
volatility premium and extreme events. Horizontal dependence of
volatility describes the tendency of local volatility clusters and the
tendence to mean-reversion. Both properties can be observed in the 7 day
rolling volatility. Sudden spikes in realized volatility form clusters
and are often followed by similar drops. The leverage effect can be
measured as a negative correlation between returns and volatility.
Phases of negative returns coincide with high volatility and vice versa.
The Pearson correlation between the underlying and realized volatility
is -12.59\%. The Pearson correlation between the underlying and the bid
(ask) IV is -16.11\% (-13.05\%). Due to the low absolute values of
correlation, the existence of a leverage effect is unlikely. However,
the sign matches the direction of a leverage effect. Considering the
large share of retailers with easy access to leverage, the lack of
significance is a surprising result. It is well known that, in general,
the IV exceeds the unconditional annualized standard deviation. This is
displayed in Figure \ref{fig: iv_comparison}. The difference between the
implied and realized volatility is commonly referred to as volatility
premium. The mean 7 day rolling volatility is 70.96\%, whereas the mean
IV at the bid and ask are 85.28\% and 93.57\%. Spikes in realized
volatility rarely exceed the ask of the IV. Consequently, the volatility
premium is shown to be substantial throughout the regarded time period.
Fat-tail events are priced in due to the classic volatility skew. This
is depicted in Figure \ref{fig: emp_ivsmile}, the empirical volatility
skew.

Since the data are high frequency order book snapshots with substantial
bid-ask spreads, we must address the issue of Market Microstructure
Noise (MMN). A simple estimate of realized volatility would increase
with our sample due to order competition within the spread. We resolve
the issue of MMN in our calculation of realized volatility by taking the
arithmetic mean after aggregating the implied volatility of bids and
asks on a daily base. Therefore the frequency is matching the underlying
BTC index. The formula for annualized, realized volatility of price
\(p\) over a window of \(w\) days is

\begin{equation}
\sigma_w = \sqrt{\frac{1}{w}\sum_{t=1}^w{(\log{p_t} - \log{p_{t-1})}^2}} \sqrt{\frac{365}{w}}
\end{equation}

Annualization is based on 365 days since Bitcoin options are traded
continuously on Deribit.

The term structure itself is easier to assess when regarding ATM-options
in a series of dates (Figure 6). ATM-options are defined as those
options, whose moneyness falls in the interval \([0.9, 1.1]\). A typical
term structure of the Bitcoin options reveals a high level of implied
volatility for shorted-dated options. For options with a
time-to-maturity of three days or more, implied volatility drops sharply
compared to the initial level. Naturally, IV is increasing with
time-to-maturity.

\begin{center}
\begin{figure}
\includegraphics[width=1\textwidth]{/Users/julian/src/up/spd/plot/iv_comparison_up.png}
\caption[Implied Volatiliy vs. Realized Volatility on Deribit]%
{Implied volatility vs. realized volatility. Realized volatility is annualized and regarded in a 7 day window.
IV is calculated as the average of observed orderbooks at the bid and ask. Described data filters apply. {\color{blue} Ask IV}, {\color{green} Bid IV}, {\color{red} 7 Day Rolling Volatility}.}
\label{fig: iv_comparison}
\href{https://github.com/QuantLet/BitcoinOptions/blob/master/src/vola.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,Volatility}
\end{figure}
\end{center}

\begin{centering}
\begin{figure}
\includegraphics[width=1\textwidth]{/Users/julian/src/up/spd/plot/vola_2021-05-23.png}
\caption[Empirical Volatility Surface]{Volatility Skew shown in the empirical Volatility Surface on 2021-05-23}
\label{fig: emp_ivsmile}
\end{figure}
\end{centering}

\begin{centering}
\begin{figure}
\includegraphics[width=1\textwidth]{/Users/julian/src/up/spd/plot/vola_2021-09-08.png}
\caption[Empirical Volatility Surface 2]{Volatility Skew shown in a calmer period on 2021-09-08}
\label{fig: emp_ivsmile2}
\href{https://github.com/QuantLet/BitcoinOptions/blob/master/src/vola_plots.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,IV Smile}
\end{figure}
\end{centering}

\begin{minipage}{.5\textwidth}
    \centering
    \includegraphics[width=2.8in]{/Users/julian/src/up/spd/plot/term_structure_atm_options_on_-2022-01-010.08.png}
    \centering
    \includegraphics[width=2.8in]{/Users/julian/src/up/spd/plot/term_structure_atm_options_on_-2022-01-020.08.png}
    \centering
    \includegraphics[width=2.8in]{/Users/julian/src/up/spd/plot/term_structure_atm_options_on_-2022-01-030.08.png}
    \centering
    \includegraphics[width=2.8in]{/Users/julian/src/up/spd/plot/term_structure_atm_options_on_-2022-01-040.08.png}

\end{minipage}
\begin{minipage}{0.5\textwidth}
    \centering
    \includegraphics[width=2.8in]{/Users/julian/src/up/spd/plot/term_structure_atm_options_on_-2022-01-050.08.png}
    \centering
    \includegraphics[width=2.8in]{/Users/julian/src/up/spd/plot/term_structure_atm_options_on_-2022-01-060.08.png}
    \centering
    \includegraphics[width=2.8in]{/Users/julian/src/up/spd/plot/term_structure_atm_options_on_-2022-01-070.08.png}
        \centering
    \includegraphics[width=2.8in]{/Users/julian/src/up/spd/plot/term_structure_atm_options_on_-2022-01-080.08.png}

\end{minipage}
    \captionof{figure}{ATM Term Structure from 2022-01-01 until 2022-01-08. Moneyness $M \in [0.9, 1.1]$}

\label{ATM Term Structure}
\href{https://github.com/QuantLet/BitcoinOptions/blob/master/main.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,Main}

\pagebreak
\newpage
\section{Empirical Pricing Kernels}

Instruments are grouped by time-to-maturity in order to summarize the
findings. All instruments with less than one week to maturity are
classified as having less than one, between one and two, between two and
three, between three and four and between five and eight weeks to
maturity. We typically observe U-shaped PKs for short-term maturities.
In a similar fashion to the oil market, investors perceive short-term
price changes in any direction as undesirable and prefer to hedge their
risk (Christoffersen et al. 2021). For short-dated instruemnts, option
writers are asking for substantial risk premia in order to reflect
unhedgeable risks, such as jumps in the underlying (Hou et al. 2020).

We find a differnet shape for longer-dated instruments. Typically, PKs
in traditional markets are monotonically decreasing. This results from
human risk aversion; investors are willing to insure themselves against
losses and have a preference for smooth consumption curves. However, PKs
outside of index options markets are not well studied (Cuesdeanu and
Jackwerth 2018). Grith et al. (2009) compartmentalize a pricing kernel
in two segments around a possible breakpoint. Testing an unrestricted
model against GMM estimates of the restricted one, they reject
monotonicity of the empirical PK in four out of five cases (performing a
D-test). The phenomenon of having a decreasing slope, although with
locally increasing sections in empirical PKs is called the pricing
kernel puzzle. We observe a similar behavior and extend the existing
literature on the pricing kernel puzzle to a new asset class.

Since the empirical PKs for short-dated instruments are U-shaped, we
figure that investors are willing to pay a high risk premium in order to
insure immediate price risks. With an increasing maturity of the
instrument, the shape of the empirical PK resembles a ``W'' or
``tilde''. This corresponds to a a lesser willingness to pay a risk
premium.\\
The dynamic of the empirical PKs also reveals the existence of a
time-varying risk premium in BTC markets. The observed PK shapes are
closely aligned with the results reported in Cuesdeanu and Jackwerth
(2018). However, their object of study are traditional markets, such as
major indices which are highly correlated with aggregate wealth. We
conclude that, due to the similarities of the empirical PKs, BTC is well
on its way to becoming an established asset class. It may prove to be
much rather a store of value than it is currently given credit for.

Apart from the discussion of results, deviations between PD and SPD can
be interpreted as trading opportunities. (Blaskowitz et al. 2004) design
and evaluate trading strategies based on deviations between PD and SPD.
Estimated PKs could be used by traders employing so-called Skewness or
Kurtosis trades. Nevertheless, traders must recognize the hidden cost of
employing such strategies in terms of substantial hedging cost, e.g.~in
the form of volatility and large spreads.

\begin{centering}
\begin{figure}
\includegraphics[width=1\textwidth]{/Users/julian/src/up/spd/plot/Pricing Kernel_over_time_nweeks=0.png}
\caption[Empirical Pricing Kernel - 0 Weeks to Expiration]{Pricing Kernels with less than one Week to Expiration}
\end{figure}
\end{centering}

\begin{centering}
\begin{figure}
\includegraphics[width=1\textwidth]{/Users/julian/src/up/spd/plot/Pricing Kernel_over_time_nweeks=1.png}
\caption[Empirical Pricing Kernel - 1 Week to Expiration]{Empirical Pricing Kernel - 1 Week to Expiration}
\end{figure}
\end{centering}

\begin{centering}
\begin{figure}
\includegraphics[width=1\textwidth]{/Users/julian/src/up/spd/plot/Pricing Kernel_over_time_nweeks=2.png}
\caption[Empirical Pricing Kernel - 2 Weeks to Expiration]{Empirical Pricing Kernel - 2 Weeks to Expiration}
\end{figure}
\end{centering}

\begin{centering}
\begin{figure}
\includegraphics[width=1\textwidth]{/Users/julian/src/up/spd/plot/Pricing Kernel_over_time_nweeks=3.png}
\caption[Empirical Pricing Kernel - 3 Weeks to Expiration]{Empirical Pricing Kernel - 3 Weeks to Expiration}
\end{figure}
\end{centering}

\begin{centering}
\begin{figure}
\includegraphics[width=1\textwidth]{/Users/julian/src/up/spd/plot/Pricing Kernel_over_time_nweeks=4.png}
\caption[Empirical Pricing Kernel - 4 Weeks to Expiration]{Empirical Pricing Kernel - 4 Weeks to Expiration}
\end{figure}
\end{centering}

\begin{centering}
\begin{figure}
\includegraphics[width=1\textwidth]{/Users/julian/src/up/spd/plot/Pricing Kernel_over_time_nweeks=8.png}
\caption[Empirical Pricing Kernel - 5-8 Weeks to Expiration]{Empirical Pricing Kernel - 5-8 Weeks to Expiration}
\href{https://github.com/QuantLet/BitcoinOptions/blob/master/main.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,Main}
\end{figure}
\end{centering}

\pagebreak
\newpage

\section{Bootstrap-based Confidence Bands}

Having computed a large set of empirical PKs with varying maturities, it
should be ensured that the observed features are not mere artefacts. A
useful tool to assess the validity of our estimated PKs are
bootstrap-based confidence bands. Following the derivation of Härdle et
al. (2014), confidence bands at any confidence level can be derived via
the ``wild bootstrap method''. Figure
\ref{fig:Bootstrapped Confidence Bands - 2 Days} until Figure
\ref{fig:Bootstrapped Confidence Bands - 84 Days} are depicted
exemplarily for the whole set. Confidence bands are tighter around the
PKs when time to maturity increases. PKs are also flatter with growing
time to maturity.

The concept of bootstrapping PK can be introduced by linearization of
the PK. Stochastic deviation of the PK can be linearised into a
stochastic part, containing the estimator of the SPD and a deterministic
part containing the expectation of the physical density. Convergence of
both parts can be proven separately.

\[
\widehat{K}(x) = \frac{\widehat{q}(x)}{\widehat{p}(x)} 
\]

\[
\begin{aligned}
\underset{x \in E}{\operatorname{sup}} \lvert \widehat{K}(x) - K(x) \rvert = \underset{x \in E}{\operatorname{sup}} \lvert \frac{\widehat{q}(x) - q(x)}{p(x)} - \frac{\widehat{p}(x) - p(x)}{p(x)} * \frac{q(x)}{p(x)} - \frac{\{\widehat{q}(x) - q(x)\} \{\widehat{p}(x) - p(x)\} }{p^2(x)} \rvert  \\
+ {\scriptstyle \mathcal{O}_p}[\operatorname{max}((n_p h_{n_p}/\log)^{-1/2} + h^2_{np},h_{n_q}^{-2} \{ n_q h_{n_q} / \log{n_q} \}^{-1/2} + h^2_{n_q})]
\end{aligned}
\]

Consider the leading term of \[
\underset{x \in E}{\operatorname{sup}} \lvert \frac{\widehat{q}(x) - q(x)}{p(x)} \rvert
\]

Resample data from the smoothed bivariate distribution of strike and
moneyness (X,Y)

\[
{\widehat{f}(x,y)} = \frac{\widehat{\sigma_X}}{n_q h^2_{n_q}\widehat{\sigma_Y}} 
\Sigma_{i=1}^{n_q} K
\{ 
\frac{X_i - x}{h_{n_q}}, 
\frac{(Y_i - y)\widehat{\sigma_X}}{h_{n_q}\widehat{\sigma_Y}}
\}
\]

Using the resampled data, calculate the bootstrap analogue

\[
\underset{x \in E}{\operatorname{sup}} \lvert \frac{\widehat{q}^*(x) - \widehat{q}(x)}{p(x)} \rvert
\]

where
\(\underset{x \in E}{\operatorname{sup}} \lvert\widehat{q}^*(x) - \widehat{q}(x) \rvert\)
converges to the Gumbel distribution at an unfortunately slow rate of
\(\frac{1}{\log n_q}\) (Hall 1991).

\begin{minipage}{.5\textwidth}
    \centering
    \includegraphics[width=3.0in]{/Users/julian/src/up/spd/plot/EPK_2021-12-31_0.0055.pdf}
    \captionof{figure}{2 Days to Maturity.}
    \label{fig:Bootstrapped Confidence Bands - 2 Days}
    \centering
    \includegraphics[width=3.0in]{/Users/julian/src/up/spd/plot/EPK_2021-12-31_0.0192.pdf}
    \captionof{figure}{7 Days to Maturity.}
    \label{fig:Bootstrapped Confidence Bands - 7 Days}
    \centering
    \includegraphics[width=3.0in]{/Users/julian/src/up/spd/plot/EPK_2021-12-31_0.0384.pdf}
    \captionof{figure}{14 Days to Maturity.}
    \label{fig:Bootstrapped Confidence Bands - 14 Days}
    \centering
    \includegraphics[width=3.0in]{/Users/julian/src/up/spd/plot/EPK_2021-12-31_0.0575.pdf}
    \captionof{figure}{21 Days to Maturity.}
    \label{fig:Bootstrapped Confidence Bands - 21 Days}

\end{minipage}
\begin{minipage}{0.5\textwidth}
    \centering
    \includegraphics[width=3.0in]{/Users/julian/src/up/spd/plot/EPK_2021-12-31_0.0767.pdf}
    \captionof{figure}{28 Days to Maturity.}
    \label{fig:Bootstrapped Confidence Bands - 28 Days}
    \centering
    \includegraphics[width=3.0in]{/Users/julian/src/up/spd/plot/EPK_2021-12-31_0.1534.pdf}
        \captionof{figure}{56 Days to Maturity.}
    \label{fig:Bootstrapped Confidence Bands - 56 Days}
    \centering
    \includegraphics[width=3.0in]{/Users/julian/src/up/spd/plot/EPK_2021-12-31_0.2301.pdf}
        \captionof{figure}{84 Days to Maturity. \color{red}{Empirical Pricing Kernels} with \color{blue}{bootstrapped Confidence Bands} on 2021-12-31.}
    \href{https://github.com/QuantLet/BitcoinOptions/blob/master/main.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,Main}
    \label{fig:Bootstrapped Confidence Bands - 84 Days}
\end{minipage}

\pagebreak
\newpage

\section{Conclusion}

Empirical PKs have been estimated using Rookley's Method on a dataset of
order book snapshots from Deribit, the leading BTC options market.
Bootstrap-based confidence bands have been estimated in order to
validate the results.

We assess the presence of well-established stylized facts of volatility
in the BTC market. We find horizontal dependence of volatility, meaning
that volatility tends to cluster and revert to the mean. The difference
between implied and realized volatility, commonly referred to as the
volatility premium, is substantial. Realized volatility rarely exceeds
the implied volatility. The IV term structure is found to be
traditional. Despite market participant's easy access to leverage, we
cannot confirm the existence of a leverage effect.

Our analysis extends the literature to a new asset class and contributes
to the pricing kernel puzzle. We are able to replicate empirical results
from the analysis of PKs in traditional markets, which are highly
correlated with aggregate wealth. Their similarity to PKs estimated from
BTC options indicates that BTC is becoming an established asset class.

Furthermore, our analysis sheds light on the BTC valuation and
risk-aversion of the retail traders. BTC option traders mostly consist
of retailers who are heterogeneous in their risk profile. The price of
short-dated instruments includes a high risk premium. Such instruments
reflect the anticipation of jumps in the underlying. These instruments
would be traded by either very risk-averse or risk-affine traders.
Long-dated instruments are employed as classic hedging instruments.
While a substantial risk premium is paid by traders in order to insure
themselves against falling prices, they are additionally (but to a
lesser degree) insuring themselves against sharply rising prices. Thus
investors are also hedging their risk of being priced out of a dynamic
market.

Several extensions of the presented research are available for the
future. Among those, we would find an extended dataset insightful, that
could support estimation of PKs and their evolution over a longer time
frame and possibly describe the convergence to the PKs of established
markets. It would also be interesting to analyze the evolution of PKs
around stress events in order to get further insights into investor risk
aversion, anticipation of such events and trading opportunities.

\section{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-aitsahalia01optionprice}{}%
Aı̈t-Sahalia, Y., Wang, Y., \& Yared, F. (2001). Do option markets
correctly price the probabilities of movement of the underlying asset?
\emph{Journal of Econometrics}, \emph{102}(1), 67--110.
\url{https://doi.org/https://doi.org/10.1016/S0304-4076(00)00091-9}

\leavevmode\hypertarget{ref-ait_sahalia_wang1998}{}%
Aït-Sahalia, Y., \& Lo, A. W. (1998). Nonparametric estimation of
state-price densities implicit in financial asset prices. \emph{The
Journal of Finance}, \emph{53}(2), 499--547.
\url{https://doi.org/https://doi.org/10.1111/0022-1082.215228}

\leavevmode\hypertarget{ref-alexander2022net}{}%
Alexander, C., Deng, J., Feng, J., \& Wan, H. (2022). Net buying
pressure and the information in bitcoin option trades. \emph{Journal of
Financial Markets}, 100764.

\leavevmode\hypertarget{ref-inverse_options}{}%
Alexander, C., \& Imeraj, A. (2021, July). Inverse options in a
black-scholes world.
\url{https://doi.org/https://doi.org/10.48550/arXiv.2107.12041}

\leavevmode\hypertarget{ref-aloosh}{}%
Aloosh, A., \& Li, J. (2019). Direct evidence of bitcoin wash trading.
\emph{Theory of Probability \& Its Applications, 5, 285-301}.
\url{https://doi.org/http://dx.doi.org/10.2139/ssrn.3362153}

\leavevmode\hypertarget{ref-bitmex}{}%
Bitmex. (2019). Presentation to the u.s. Securities and exchange
commission on wash trading.
\url{https://www.sec.gov/comments/sr-nysearca-2019-01/srnysearca201901-5164833-183434.pdf}.
Accessed 18 January 2022

\leavevmode\hypertarget{ref-blase}{}%
Blaskowitz, O., Härdle, W. K., \& Schmidt, P. (2004). Skewness and
kurtosis trades. In (pp. 1--14).
\url{https://doi.org/10.1007/978-0-8176-8180-7_1}

\leavevmode\hypertarget{ref-speculation}{}%
Blau, B. M. (2017). Price dynamics and speculative trading in bitcoin.
\emph{Research in International Business and Finance}, \emph{41},
493--499.
\url{https://doi.org/https://doi.org/10.1016/j.ribaf.2017.05.010}

\leavevmode\hypertarget{ref-bloomberg-retail-share}{}%
Bloomberg. (2021, November 17). Retail traders slide back below 20\% of
market's total volume.
\url{https://www.bloomberg.com/news/articles/2021-11-17/retail-traders-retreat-as-choppy-markets-challenge-easy-profits}.
Accessed 24 August 2021

\leavevmode\hypertarget{ref-breeden_litzenberger78}{}%
Breeden, D. T., \& Litzenberger, R. H. (1978). Prices of
state-contingent claims implicit in option prices. \emph{The Journal of
Business}, \emph{51}(4), 621--651.
\url{http://www.jstor.org/stable/2352653}

\leavevmode\hypertarget{ref-coins_with_benefits}{}%
Chen, C. Y., \& Vinogradov, D. (2021). Coins with benefits: On
existence, pricing kernel and risk premium of cryptocurrencies.
\emph{Available at SSRN 3959984}.
\url{https://doi.org/http://dx.doi.org/10.2139/ssrn.3864578}

\leavevmode\hypertarget{ref-choi}{}%
Choi, S., \& Shin, J. (2022). Bitcoin: An inflation hedge but not a safe
haven. \emph{Finance Research Letters}, \emph{46}, 102379.
\url{https://doi.org/https://doi.org/10.1016/j.frl.2021.102379}

\leavevmode\hypertarget{ref-oil_spd}{}%
Christoffersen, P., Jacobs, K., \& Pan, X. (. (2021). The State Price
Density Implied by Crude Oil Futures and Option Prices. \emph{The Review
of Financial Studies}, \emph{35}(2), 1064--1103.
\url{https://doi.org/10.1093/rfs/hhab011}

\leavevmode\hypertarget{ref-cryptonews_bitcoin_retail_share}{}%
Coinbase, S., Cryptonews. (2022). Institutions \& retail compete for
bitcoin - whose hands are stronger?
\url{https://cryptonews.com/exclusives/institutions-retail-compete-for-bitcoin-which-is-the-biggest-9695.htm}.
Accessed 22 March 2022

\leavevmode\hypertarget{ref-decentralized_wash_trading}{}%
Cong, L., Li, X., Tang, K., \& Yang, Y. (2019). Crypto wash trading.
\emph{SSRN Electronic Journal}.
\url{https://doi.org/10.2139/ssrn.3530220}

\leavevmode\hypertarget{ref-jackwerth_pk}{}%
Cuesdeanu, H., \& Jackwerth, J. C. (2018). The pricing kernel puzzle:
survey and outlook. \emph{Annals of Finance}, \emph{14}(3), 289--329.
\url{https://doi.org/10.1007/s10436-017-0317-9}

\leavevmode\hypertarget{ref-cross_border_payments}{}%
Deng, Q. (2020). Application analysis on blockchain technology in
cross-border payment. In.
\url{https://doi.org/10.2991/aebmr.k.200306.050}

\leavevmode\hypertarget{ref-deribit-insurance-fund}{}%
Deribit. (2020a). Deribit insurance fund.
\url{https://www.deribit.com/main\#/insurance}. Accessed 4 January 2021

\leavevmode\hypertarget{ref-synthetic-btc-index}{}%
Deribit. (2020b). Deribit btc-usd index.
\url{https://www.deribit.com/main\#/indexes}. Accessed 2 January 2021

\leavevmode\hypertarget{ref-market-maker-obligations}{}%
Deribit. (2020c). Deribit market maker obligations.
\url{https://www.deribit.com/pages/docs/options}. Accessed 2 January
2021

\leavevmode\hypertarget{ref-deribit_retail_piechart}{}%
Deribit. (2022a). Deribit volume share.
\url{https://www.deribit.com/statistics/BTC/market-data}. Accessed 25
July 2022

\leavevmode\hypertarget{ref-deribit_margin}{}%
Deribit. (2022b). Deribit margin requirements.
\url{https://legacy.deribit.com/pages/docs/options}. Accessed 2 May 2022

\leavevmode\hypertarget{ref-deribit_fees}{}%
Deribit. (2022c). Deribit margin requirements.
\url{https://www.deribit.com/kb/fees}. Accessed 2 May 2022

\leavevmode\hypertarget{ref-deribit-rate-limit}{}%
Deribit. (2023). Deribit rate limits.
\url{https://www.deribit.com/kb/deribit-rate-limits}. Accessed 16 April
2021

\leavevmode\hypertarget{ref-vola_premium}{}%
Eraker, B. (2021). The volatility premium. \emph{Quarterly Journal of
Finance (QJF)}, \emph{11}(03), 1--35.
\url{https://EconPapers.repec.org/RePEc:wsi:qjfxxx:v:11:y:2021:i:03:n:s2010139221500142}

\leavevmode\hypertarget{ref-networkeffect}{}%
Gandal, N., \& Halaburda, H. (2016). Can we predict the winner in a
market with network effects? Competition in cryptocurrency market.
\emph{Games}, \emph{7}(3). \url{https://doi.org/10.3390/g7030016}

\leavevmode\hypertarget{ref-Grith09sim}{}%
Grith, M., Härdle, W. K., \& Park, J. (2009). Shape invariant modelling
pricing kernels and risk aversion.
\url{https://doi.org/10.2139/ssrn.2894289}

\leavevmode\hypertarget{ref-hall_1991}{}%
Hall, P. (1991). Edgeworth expansions for nonparametric density
estimators, with applications. \emph{Statistics}, \emph{22}(2),
215--232. \url{https://doi.org/10.1080/02331889108802305}

\leavevmode\hypertarget{ref-hayes}{}%
Hayes, A. S. (2019). Bitcoin price and its marginal cost of production:
Support for a fundamental value. \emph{Applied Economics Letters},
\emph{26}(7), 554--560.
\url{https://doi.org/10.1080/13504851.2018.1488040}

\leavevmode\hypertarget{ref-w2yo}{}%
Härdle, W. K., Okhrin, Y., \& Wang, W. (2014). Uniform Confidence Bands
for Pricing Kernels. \emph{Journal of Financial Econometrics},
\emph{13}(2), 376--413. \url{https://doi.org/10.1093/jjfinec/nbu002}

\leavevmode\hypertarget{ref-pricing_cryptocurrency_options}{}%
Hou, A. J., Wang, W., Chen, C. Y. H., \& Härdle, W. K. (2020). Pricing
Cryptocurrency Options*. \emph{Journal of Financial Econometrics},
\emph{18}(2), 250--279. \url{https://doi.org/10.1093/jjfinec/nbaa006}

\leavevmode\hypertarget{ref-xfg_spd}{}%
Huynh, K., Kervella, P., \& Zheng, J. (2002). \emph{Estimating
state-price densities with nonparametric regression} (SFB 373 Discussion
Papers No. 2002,40). Humboldt University of Berlin, Interdisciplinary
Research Project 373: Quantification; Simulation of Economic Processes.
\url{https://EconPapers.repec.org/RePEc:zbw:sfb373:200240}

\leavevmode\hypertarget{ref-pumpndump}{}%
La Morgia, M., Mei, A., Sassi, F., \& Stefa, J. (2020). Pump and dumps
in the bitcoin era: Real time detection of cryptocurrency market
manipulations. In \emph{2020 29th international conference on computer
communications and networks (icccn)} (pp. 1--9).
\url{https://doi.org/10.1109/ICCCN49398.2020.9209660}

\leavevmode\hypertarget{ref-vola_stylized_facts}{}%
Masset, P. (2011). Volatility stylized facts. \emph{SSRN Electronic
Journal}. \url{https://doi.org/10.2139/ssrn.1804070}

\leavevmode\hypertarget{ref-satoshi2009}{}%
Nakamoto, S. (2009). Bitcoin: A peer-to-peer electronic cash system.
\url{http://www.bitcoin.org/bitcoin.pdf}

\leavevmode\hypertarget{ref-fundamental-theorem-of-asset-pricing}{}%
Pascucci, A., \& Agliardi, R. (2011). PDE and martingale methods in
option pricing. In (pp. 429--495).

\leavevmode\hypertarget{ref-rodgers2019}{}%
Rodgers, T. (2019, April 4). 95 percent of volume could be wash trading
as bitcoin price surges.
\url{https://www.forbes.com/sites/tomrodgers1/2019/04/04/99-of-volume-could-be-wash-trading-as-bitcoin-takes-back-5000/?sh=4640bcbc23d5}.
Accessed 15 April 2023

\leavevmode\hypertarget{ref-Rookley97}{}%
Rookley, C. (1997). Fully exploiting the information content of intra
day option quotes: Applications in option pricing and risk management.

\leavevmode\hypertarget{ref-sar2017}{}%
Sar, M. (2017). Dodd-frank and the spoofing prohibition in commodities
markets.

\leavevmode\hypertarget{ref-deriv_volume}{}%
Skew.com. (2022). Open interest for bitcoin options over time. Skew is a
leading crypto data analytics platform for institutional clients.
\url{https://analytics.skew.com/dashboard/bitcoin-options}. Accessed 25
July 2022

\leavevmode\hypertarget{ref-brazilian_speculation}{}%
Smaniotto, E. N., \& Neto, G. B. (2022). Speculative trading in bitcoin:
A brazilian market evidence. \emph{The Quarterly Review of Economics and
Finance}, \emph{85}, 47--54.
\url{https://doi.org/https://doi.org/10.1016/j.qref.2020.10.024}

\leavevmode\hypertarget{ref-cross_border_stellar}{}%
Stellar Development Foundation. (2022). 2022 research reveals high
awareness and growing cross-border use of cryptocurrency in four key
markets.
\url{https://www.prnewswire.com/news-releases/2022-research-reveals-high-awareness-and-growing-cross-border-use-of-cryptocurrency-in-four-key-markets-301512401.html}.
Accessed 22 July 2022

\leavevmode\hypertarget{ref-nickszabo}{}%
Szabo, N. (2008). Bit gold.
\url{https://unenumerated.blogspot.com/2005/12/bit-gold.html}. Accessed
28 July 2022

\leavevmode\hypertarget{ref-tuccella2021}{}%
Tuccella, J.-N., Nadler, P., \& Șerban, O. (2021). Protecting retail
investors from order book spoofing using a gru-based detection model.

\end{document}