### 2024-01-19 written/paper.tex
minor written

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12,
]{article}
\usepackage{lmodern}
\usepackage[table,xcdraw]{xcolor}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Empirical Option Returns and the risk-free Rate in Crypto Asset Markets},
  pdfauthor={Julian Winkel; Wolfgang K. Härdle},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin = 2.5cm]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{float}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{subfig}
\usepackage{caption}
\usepackage{eucal}
\floatplacement{figure}{H}

\title{Empirical Option Returns and the risk-free Rate in Crypto Asset Markets}
\author{Julian Winkel\footnote{International Research Training Group 1792;
  Humboldt-Universität zu Berlin,
  \href{mailto:julian.winkel@hu-berlin.de}{\nolinkurl{julian.winkel@hu-berlin.de}}} \and Wolfgang K. Härdle\footnote{BRC Blockchain Research Center,
  Humboldt-Universität zu Berlin, Berlin, Germany; Sim Kee Boon
  Institute, Singapore Management University, Singapore; WISE Wang Yanan
  Institute for Studies in Economics, Xiamen University, Xiamen, China;
  Dept. Information Management and Finance, National Chiao Tung
  University, Hsinchu, Taiwan, ROC; Dept. Mathematics and Physics,
  Charles University, Prague, Czech Republic, Grants -- DFG IRTG 1792
  and the Czech Science Foundation's grant no. 19-28231X / CAS XDA
  23020303 gratefully acknowledged,
  \href{mailto:haerdle@hu-berlin.de}{\nolinkurl{haerdle@hu-berlin.de}}}}
\date{7 January 2024}

\begin{document}
\maketitle

\newcommand{\quantnet}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,}
\newcommand{\quantinar}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/quantinar_logo.png}}\,}

\definecolor{bittersweet}{rgb}{1.0, 0.44, 0.37} 
\definecolor{ao}{rgb}{0.0, 0.5, 0.0}
\definecolor{bananamania}{rgb}{0.98, 0.91, 0.71}
\definecolor{bluegray}{rgb}{0.4, 0.6, 0.8}

\pagenumbering{roman}

\begin{abstract}

\singlespacing 
abstract

 \medskip
 \textbf{Keywords:} Bitcoin, Deribit, pricing kernel, risk aversion, speculation, hedging  \\
 \indent \textbf{JEL Classification:} C14, C50, G10  \\
 \bigskip

\end{abstract}

\pagenumbering{arabic}

\newpage

\section{Introduction}
Despite the ongoing attention that crypto assets receive and despite their ongoing integration into traditional markets, whereas the last example would be the SEC's approval of a range of Bitcoin Spot ETFs, fundamental questions about crypto asset markets remain unclear:
How efficient are crypto asset markets relative to established, traditional markets such as equities? 
Does a risk-free rate exist and if yes, how does it relate to the risk-free rate in established markets? Are derivatives, written on crypto assets, fairly priced or at least fairly priced relative to other derivatives? What do we know about the risk-return characteristics of crypto assets?
We assess these fundamental questions and find that the methods required to conduct our analysis have been out in the open for long; in fact we are employing similar methods to those, under which the efficiency of equity option markets had been assessed in their "youth" in the early 2000s. 
The common key behind the employed methods is the precise and bespoke tailoring of risk-return characteristics through the use of options, as proposed by Shumway and Bakshi et al in XZY year.

We discuss the existence of a risk-free rate in crypto asset markets and derive a proxy. We use it as a benchmark that beta-neutral portfolios should yield. 
We construct and evaluate the performance of beta-neutral portfolios and find that they do not generate their theoretical returns for particular levels of moneyness. We compare the risk-return characteristics to the late 90s equity markets and find, in a similar fashion as the results reported by Bakshi et al and Shumway et al., severe overpricings of at-the-money options. The effect holds after compensating for market crash insurance and transaction cost. 
We assess the deadweight loss due to overpricings and propose a simple trading strategy that exploits this partciular inefficiency.

We constructs and evaluate the empirical performance of option portfolios, written on crypto assets such as Bitcoin and Ether, to allow model-free insights into the existence and degree of a leverage effect, the existence and size of risk premia, resulting market inefficiencies (measurable in economic loss / money lost) and to explore the option (pun intended)/possibility of curing named inefficiencies under transaction cost and liquidity constraints.

What evidence or contradictions can we find?
The precise and bespoke tailoring of risk-return characteristics through the use of options allows promising insights into the risk-return characteristics of crypto assets, such as Bitcoin and Ether.

From our point of view, the primary question leading such research is the question of market efficiency.

Due to the youth of the crypto asset option markets, the primary question foregoing such research is that of the efficiency of named market. Do the laws that govern equity option markets also apply to crypto asset option markets? Can we find empirical evidence or contradictions?

- Is there a risk-free rate in crypto markets?
- How efficient is the crypto (option) market compared to equity option markets?

Another, so far rarely used pathway is the assessment of the risk-return characteristics of Bitcoin through the lens of the options written on it, which allow new model-free insights into the existence and degree of a leverage effect, the existence and size of risk premia, resulting market inefficiencies (measurable in economic loss / money lost) and to explore the option (pun intended)/possibility of curing named inefficiencies under transaction cost and liquidity constraints.

This reasoning follows the insights of Shumway and Bakshi, who separately evaluated option returns in equity markets by means of delta-neutral option portfolios in the early 2000s.

Bakshi and Shumway both found that expected call returns exceed those of the underlying and that expected put returns underperform the underlying. Both also found that the effect increases (decreases) with the strike. 
Bakshi found that zero-beta, at-the-money straddle positions produce average losses of approximately three percent per week and suggests that the reason may be a systematic volatility premium being priced in options.

Following Shumway, the risk-return characteristics of options can be separated in two elements, a leverage effect and a curvature effect due to the sensitivity of option payoffs to higher moments of the underlying. 
First, the holder (buyer) of an option is subject to a leverage effect since exposure to the underlying can be obtained through the payment of a premium, which is small compared to the price of the underlying. 
If there is a leverage effect, then (long) call options should outperform their underlying and (long) put options should underperform their underlying. Furthermore, the effect should increase (decrease) with the strike price for calls (puts).
This reasoning allows a simple test of existence and size of the leverage effect and thus conclusions about the level of speculation associated to Bitcoin valuation.

(The second component, related to the sensitivity of options towards higher moments, implies that options are redundant assets...)

If options were redundant assets, then apart of the leverage effect, delta-neutral option portfolios should earn the risk-free rate (no premium) on average.
Until now, there is no risk-free rate available in the space of crypto assets. We establish a proxy or reference rate for a possible risk-free rate, which results from the perpetual futures written on Bitcoin and Ether.

Use Deribit and Binance perpetual futures for this, equally weighted.

(This proxy is clearly not risk-free due to counterparty risk, but we argue that a carry trade involving a short position on perpetual futures on multiple platforms in conjunction with a secured, on-chain underlying position reduces the platform-induced counterparty risk (which is the residual risk in the carry trade) as far as possible.) (This also leaves liquidity risk though!!)

All underlying code is available on \href{quantlet.com}{quantlet.com}
and a corresponding courselet is available on
\href{quantinar.com}{quantinar.com}.

\section{Literature Review}

\section{Methodology}

- scrape 8h funding rate continuously and use as market-interest rate

- scrape all BTC, ETH transactions from Deribit 
- load data, clean
- center on settlement date for each day (10am UTC)
- for each option, find the pendant via put-call-parity
- calculate the beta of each put, call
- build zero-beta straddles by finding weights such that the sum of weights is zero
- assess empirical, daily returns of the straddle
- build boxplots over levels of moneyness
- KDE --> t-distribution
- perform t-test for population mean of 0

\section{Data}

We use the \href{blockchain-research-center.de}{Blockchain Research Center} to retrieve a complete set of all transactions conducted on Deribit's options written on Bitcoin and Ether. 
The exchange started operations in July 2017. 
Furthermore, we use Deribits API V2 to retrieve a complete set of funding rates since (approx.) July 2019.

- table of amount of transactions per group
- plot: amount of transactions over time per group (crypto)

\section{A Proxy for the risk-free Rate of Bitcoin and Ether}

Evidently, there is no established risk-free rate in crypto (or digital asset) markets. In the absence of riskless bond, we refer to the next best thing as a proxy for the risk free rate, which is the replication of the market-implied yield through a position which only has liquidity risk and counterparty risk. To be fair, counterparty risk also exists for risk-free bonds.

A large share of the trading volume in digital asset markets falls on the futures market. Among the futures trading, perpetual instruments have historically been and currently are the most liquid instruments (show both!). While it would be possible to use a yield from a long-short futures position (e.g. short a short-tenor future and long a longer-tenor future), we argue that in order to minimize liquidity risk, the most liquid instruments should be taken into consideration. These are the perpetual futures, which are listed on some exchanges.

As of 2024-01-07, the total Open Interest for Bitcoin Perpetual Futures contracts is \$19,343,021,437 with a total 24h Trade Volume of \$50,339,886,015.
Open Interest for Ether Perpetual Futurews contracts is Total Open Interest \$14,945,055,482 and total 24h Trade Volume is \$18,097,004,167.

BTC Futures

Total Open Interest $1,405,918,350
Total 24h Trade Volume $324,820,482

ETH Futures
Total Open Interest $466,932,584
Total 24h Trade Volume $98,238,298

\begin{table}[]
\begin{tabular}{lll}
\textbf{Instrument}            & \textbf{Aggregate Open Interest (USD)} & \textbf{24 Hour Aggregate Trade Volume (USD)} \\
BTC Perpetual Futures & 19,343,021,437                & 50,339,886,015                       \\
ETH Perpetual Futures & 14,945,055,482                & 18,097,004,167                       \\
BTC Futures           & 1,405,918,350                 & 324,820,482                          \\
ETH Futures           & 466,932,584                   & 98,238,298                          
\end{tabular}
\end{table}

Source for liquidity: 
https://www.coingecko.com/en/derivatives
https://www.coingecko.com/en/derivatives/futures

% Source: https://www.tablesgenerator.com/#
% Please add the following required packages to your document preamble:
% \usepackage[table,xcdraw]{xcolor}
% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage[table,xcdraw]{xcolor}
% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}

As of 2024-01-07

Bitcoin Perpetuals with 
Total Open Interest \$19,343,021,437
Total 24h Trade Volume \$50,339,886,015

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{llrrrrrrr}
\hline
\textbf{Exchange}    & \textbf{Symbol} & \multicolumn{1}{l}{\textbf{Price}} & \multicolumn{1}{l}{\textbf{Index Price}} & \multicolumn{1}{l}{\textbf{Basis}} & \multicolumn{1}{l}{\textbf{Spread}} & \multicolumn{1}{l}{\textbf{Funding Rate}} & \multicolumn{1}{l}{\textbf{Open Interest}} & \multicolumn{1}{l}{\textbf{24h Volume}} \\
\hline
\hline
\# Binance (Futures) & BTCUSDT         & \$43,995.06                        & 44011.67 USDT                            & +0.087\%                           & 0.01\%                              & 0.010\%                                   & \$3,214,727,375                            & \$6,822,091,229.25                      \\
\# Binance (Futures) & BTCUSD\_PERP    & \$44,008.70                        & 44037.64 USD                             & +0.066\%                           & 0.01\%                              & 0.010\%                                   & \$808,036,100                              & \$973,585,131.72                        \\
\# BingX (Futures)   & BTC-USDT        & \$44,042.40                        & 44046.6 USDT                             & +0.044\%                           & 0.01\%                              & 0.010\%                                   & \$909,897,901                              & \$1,867,895,422.35                      \\
\# Bitget Futures    & BTCUSDT\_UMCBL  & \$44,049.08                        & 44033.85 USDT                            & +0.015\%                           & 0.01\%                              & 0.011\%                                   & \$1,790,553,246                            & \$5,701,072,691.79                      \\
\# Bitmart Futures   & BTCUSDT         & \$44,015.47                        & 44003.79 USDT                            & +0.024\%                           & 0.01\%                              & 0.010\%                                   & \$641,006,203                              & \$2,917,088,729.74                      \\
\# Bybit (Futures)   & BTCUSDT         & \$44,039.88                        & 44016.82 USDT                            & -0.003\%                           & 0.01\%                              & 0.010\%                                   & \$2,538,053,148                            & \$2,672,726,926.85                      \\
\# Bybit (Futures)   & BTCUSD          & \$43,995.50                        & 44037.2 USD                              & +0.095\%                           & 0.01\%                              & 0.010\%                                   & \$652,512,486                              & \$283,394,472.00                        \\
\# Deribit           & BTC-PERPETUAL   & \$44,020.00                        & 44028.17 USD                             & +0.019\%                           & 0.01\%                              & 0.004\%                                   & \$577,506,570                              & \$204,526,951.86                        \\
\# OKX (Futures)     & BTC-USDT-SWAP   & \$44,014.07                        & 43989.3 USDT                             & -0.006\%                           & 0.01\%                              & -0.009\%                                  & \$984,130,801                              & \$2,648,725,568.71                      \\
\# Prime XBT         & BTC/USD         & \$43,981.30                        & -                                        & -                                  & 0.02\%                              & -                                         & \$920,322,200                              & \$60,224,491.31                        
\end{tabular}}
\end{table}

Ether Perpetuals

Total Open Interest \$14,945,055,482
Total 24h Trade Volume \$18,097,004,167

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage[table,xcdraw]{xcolor}
% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage[table,xcdraw]{xcolor}
% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}

\begin{table}[ht]
\resizebox{\textwidth}{!}{\begin{tabular}{llrrrrrrr}
\hline
\textbf{Exchange}         & \textbf{Symbol} & \multicolumn{1}{l}{\textbf{Price}} & \multicolumn{1}{l}{\textbf{Index Price}} & \multicolumn{1}{l}{\textbf{Basis}} & \multicolumn{1}{l}{\textbf{Spread}} & \multicolumn{1}{l}{\textbf{Funding Rate}} & \multicolumn{1}{l}{\textbf{Open Interest}} & \multicolumn{1}{l}{\textbf{24h Volume}} \\
\hline
\hline
\# Deepcoin (Derivatives) & ETHUSDT         & \$2,232.18                         & 2232.46 USDT                             & +0.047\%                           & 0.02\%                              & 0.006\%                                   & \$6,376,015,545                            & \$237,202,788.68                        \\
\# Binance (Futures)      & ETHUSDT         & \$2,227.11                         & 2227.74 USDT                             & +0.078\%                           & 0.01\%                              & 0.010\%                                   & \$1,680,165,204                            & \$2,843,630,044.97                      \\
\# Bybit (Futures)        & ETHUSDT         & \$2,228.15                         & 2228.15 USDT                             & +0.049\%                           & 0.01\%                              & 0.010\%                                   & \$1,027,647,498                            & \$540,402,664.24                        \\
\# OKX (Futures)          & ETH-USDT-SWAP   & \$2,227.96                         & 2227.07 USDT                             & +0.010\%                           & 0.01\%                              & -0.003\%                                  & \$584,652,223                              & \$1,748,266,659.63                      \\
\# Gate.io (Futures)      & ETH\_USDT       & \$2,228.62                         & 2227.03 USDT                             & -0.021\%                           & 0.01\%                              & 0.010\%                                   & \$508,608,257                              & \$189,744,908.47                        \\
\# BingX (Futures)        & ETH-USDT        & \$2,229.64                         & 2230.26 USDT                             & +0.062\%                           & 0.02\%                              & 0.010\%                                   & \$475,301,553                              & \$609,765,117.59                        \\
\# Prime XBT              & ETH/USD         & \$2,227.49                         & -                                        & -                                  & 0.04\%                              & -                                         & \$473,765,000                              & \$16,309,868.89                         \\
\# Deribit                & ETH-PERPETUAL   & \$2,228.05                         & 2228.59 USD                              & +0.024\%                           & 0.01\%                              & 0.004\%                                   & \$447,911,369                              & \$82,175,344.52                         \\
\# Binance (Futures)      & ETHUSD\_PERP    & \$2,227.69                         & 2228.74 USD                              & +0.047\%                           & 0.01\%                              & 0.010\%                                   & \$366,493,140                              & \$257,267,395.24                        \\
\# XT.COM (Derivatives)   & ETH\_USDT       & \$2,228.19                         & 2225.55 USDT                             & +0.031\%                           & 0.01\%                              & -0.010\%                                  & \$325,480,371                              & \$1,137,587,629.30                     
\end{tabular}}
\end{table}

BTC Futures

with

Total Open Interest $1,405,918,350
Total 24h Trade Volume $324,820,482

\begin{table}[ht]
\resizebox{\textwidth}{!}{\begin{tabular}{llrrrrrrr}
\hline
\textbf{Exchange}                & \textbf{Symbol}    & \multicolumn{1}{l}{\textbf{Price}} & \multicolumn{1}{l}{\textbf{Index Price}} & \multicolumn{1}{l}{\textbf{Basis}} & \multicolumn{1}{l}{\textbf{Spread}} & \multicolumn{1}{l}{\textbf{Expires in}} & \multicolumn{1}{l}{\textbf{Open Interest}} & \multicolumn{1}{l}{\textbf{24h Volume}} \\
\hline
\hline
\# Binance (Futures)             & BTCUSD\_240329     & \$45,372.20                        & 44041.17 USD                             & -2.934\%                           & 0.01\%                              & 3 months                                & \$136,765,200                              & \$19,534,741.00                         \\
\# Binance (Futures)             & BTCUSD\_240628     & \$46,802.20                        & 44037.64 USD                             & -5.907\%                           & 0.02\%                              & 6 months                                & \$60,673,400                               & \$11,015,730.09                         \\
\# Binance (Futures)             & BTCUSDT\_240329    & \$45,340.32                        & 44011.62 USDT                            & -2.883\%                           & 0.04\%                              & -                                       & \$38,608,233                               & \$13,341,161.63                         \\
\# Binance (Futures)             & BTCUSDT\_240628    & \$46,588.03                        & 44011.62 USDT                            & -5.484\%                           & 0.02\%                              & -                                       & \$7,296,676                                & \$6,727,917.37                          \\
\# BitMEX (Derivative)           & XBTH24             & \$45,170.00                        & 44016.02 USD                             & -2.555\%                           & 0.03\%                              & 3 months                                & \$23,123,300                               & \$2,586,155.50                          \\
\# BitMEX (Derivative)           & XBTM24             & \$46,319.00                        & 44016.02 USD                             & -4.972\%                           & 0.15\%                              & 6 months                                & \$14,701,700                               & \$899,753.13                            \\
\# BitMEX (Derivative)           & XBTF24             & \$44,599.50                        & 44016.02 USD                             & -1.308\%                           & 0.18\%                              & 19 days                                 & \$3,686,400                                & \$42,073.68                             \\
\# BitMEX (Derivative)           & XBTU24             & \$47,680.00                        & 44016.02 USD                             & -7.685\%                           & 0.82\%                              & 9 months                                & \$874,700                                  & \$14,332.62                             \\
\# BitMEX (Derivative)           & XBTG24             & \$45,034.50                        & 44016.02 USD                             & -2.262\%                           & 0.2\%                               & about 2 months                          & \$149,400                                  & \$502.57                                \\
\# BitMEX (Derivative)           & XBTUSDTH24         & \$44,913.12                        & 43991.27 USDT                            & -2.003\%                           & 22.08\%                             & 3 months                                & \$67,027                                   & \$23,309.91                             \\
\# BTSE (Futures)                & BTC-240329         & \$45,346.20                        & 45337.26 USD                             & -0.020\%                           & 0.15\%                              & 3 months                                & \$70,779,838                               & \$70,829.75                             \\
\# BTSE (Futures)                & BTC-240628         & \$46,578.30                        & 46774.54 USD                             & +0.421\%                           & 0.2\%                               & 6 months                                & \$68,641,915                               & \$58,755.91                             \\
\# Bybit (Futures)               & BTCUSDH24          & \$45,515.00                        & 44050.12 USD                             & -3.218\%                           & 0.01\%                              & 3 months                                & \$31,042,370                               & \$8,992,250.00                          \\
\# Bybit (Futures)               & BTCUSDM24          & \$46,867.50                        & 44046.87 USD                             & -6.018\%                           & 0.01\%                              & 6 months                                & \$13,200,403                               & \$932,634.00                            \\
\# Bybit (Futures)               & BTC-29MAR24        & \$45,288.92                        & 44042.99 USDC                            & -2.786\%                           & 0.07\%                              & 3 months                                & -                                          & \$1,650,409.10                          \\
\# Bybit (Futures)               & BTC-28JUN24        & \$46,655.43                        & 44042.83 USDC                            & -5.633\%                           & 0.07\%                              & 6 months                                & -                                          & \$636,256.66                            \\
\# Bybit (Futures)               & BTC-12JAN24        & \$44,205.30                        & 44042.83 USDC                            & -0.403\%                           & 0.02\%                              & 5 days                                  & -                                          & \$320,758.50                            \\
\# Bybit (Futures)               & BTC-19JAN24        & \$44,341.25                        & 44042.83 USDC                            & -0.708\%                           & 0.03\%                              & 12 days                                 & -                                          & \$117,823.82                            \\
\# Bybit (Futures)               & BTC-27SEP24        & \$47,694.06                        & 44042.99 USDC                            & -7.688\%                           & 0.08\%                              & 9 months                                & -                                          & \$66,406.59                             \\
\# Bybit (Futures)               & BTC-26JAN24        & \$44,370.24                        & 44042.75 USDC                            & -0.773\%                           & 0.02\%                              & 19 days                                 & -                                          & \$47,245.15                             \\
\# Bybit (Futures)               & BTC-23FEB24        & \$44,806.59                        & 44042.99 USDC                            & -1.739\%                           & 0.03\%                              & about 2 months                          & -                                          & \$8,650.36                              \\
\# Crypto.com Exchange (Futures) & BTCUSD-240628      & \$46,384.20                        & 44003.38 USD                             & -5.133\%                           & 0.41\%                              & -                                       & \$24,557,224                               & \$15,659.31                             \\
\# Crypto.com Exchange (Futures) & BTCUSD-240329      & \$45,024.60                        & 44003.38 USD                             & -2.268\%                           & 0.46\%                              & -                                       & \$1,118,289                                & \$17,244.42                             \\
\# Crypto.com Exchange (Futures) & BTCUSD-240126      & \$44,215.10                        & 44003.38 USD                             & -0.479\%                           & 0.29\%                              & -                                       & \$9,701                                    & \$18,428.85                             \\
\# Crypto.com Exchange (Futures) & BTCUSD-240223      & \$44,624.00                        & 44003.38 USD                             & -1.391\%                           & 0.43\%                              & -                                       & \$3,539                                    & \$14,931.19                             \\
\# Deribit                       & BTC-29MAR24        & \$45,300.00                        & 44026.9 USD                              & -2.810\%                           & 0.02\%                              & 3 months                                & \$208,084,190                              & \$15,282,281.56                         \\
\# Deribit                       & BTC-26JAN24        & \$44,475.00                        & 44026.9 USD                              & -1.008\%                           & 0.1\%                               & 19 days                                 & \$89,621,140                               & \$3,132,087.87                          \\
\# Deribit                       & BTC-28JUN24        & \$46,577.50                        & 44026.9 USD                              & -5.476\%                           & 0.02\%                              & 6 months                                & \$79,314,230                               & \$7,916,251.72                          \\
\# Deribit                       & BTC-27SEP24        & \$47,715.00                        & 44026.9 USD                              & -7.729\%                           & 0.09\%                              & 9 months                                & \$32,099,570                               & \$1,034,939.35                          \\
\# Deribit                       & BTC-12JAN24        & \$44,110.00                        & 44026.9 USD                              & -0.188\%                           & 0.03\%                              & 5 days                                  & \$25,884,930                               & \$7,230,738.10                          \\
\# Deribit                       & BTC-23FEB24        & \$44,767.50                        & 44026.9 USD                              & -1.654\%                           & 0.03\%                              & about 2 months                          & \$22,571,940                               & \$744,850.73                            \\
\# Deribit                       & BTC-27DEC24        & \$49,247.50                        & 44026.9 USD                              & -10.601\%                          & 0.13\%                              & 12 months                               & \$11,398,600                               & \$600,503.46                            \\
\# Deribit                       & BTC-19JAN24        & \$44,222.50                        & 44026.9 USD                              & -0.442\%                           & 0.04\%                              & 12 days                                 & \$5,538,440                                & \$2,799,269.71                          \\
\# HTX Futures                   & BTC\_CQ            & \$45,297.68                        & 44023.11 USD                             & -2.814\%                           & 0.38\%                              & 3 months                                & \$1,002,200                                & \$1,602,716.03                          \\
\# HTX Futures                   & BTC\_CW            & \$44,095.65                        & 44030.71 USD                             & -0.147\%                           & 0.45\%                              & 4 days                                  & \$47,100                                   & \$1,011,419.90                          \\
\# HTX Futures                   & BTC\_NW            & \$44,211.45                        & 44023.11 USD                             & -0.426\%                           & 0.4\%                               & 11 days                                 & \$9,000                                    & \$967,082.71                            \\
\# Kraken (Futures)              & FI\_XBTUSD\_240329 & \$45,520.50                        & 44013.38 USD                             & -3.311\%                           & 0.08\%                              & 3 months                                & \$6,608,999                                & \$112,442.00                            \\
\# Kraken (Futures)              & FI\_XBTUSD\_240628 & \$46,571.50                        & 44013.38 USD                             & -5.493\%                           & 0.24\%                              & 6 months                                & \$1,929,982                                & \$275,529.00                            \\
\# Kraken (Futures)              & FI\_XBTUSD\_240126 & \$44,191.50                        & 44013.38 USD                             & -0.403\%                           & 0.4\%                               & 19 days                                 & \$300,856                                  & \$3.00                                  \\
\# Kraken (Futures)              & FF\_XBTUSD\_240329 & \$45,284.00                        & 44013.38 USD                             & -2.806\%                           & 0.11\%                              & 3 months                                & \$87                                       & \$22.68                                 \\
\# Kraken (Futures)              & FF\_XBTUSD\_240126 & \$44,307.00                        & 44013.38 USD                             & -0.663\%                           & 0.08\%                              & 19 days                                 & \$32                                       & \$4.61                                  \\
\# KuCoin Futures                & XBTMH24            & \$45,444.00                        & 44072.8 USD                              & -3.017\%                           & 0.19\%                              & 3 months                                & \$11,251,580                               & \$86,193.00                             \\
\# OKX (Futures)                 & BTC-USD-240329     & \$45,322.90                        & 44016.1 USD                              & -2.883\%                           & 0.01\%                              & 3 months                                & \$175,555,100                              & \$42,597,534.31                         \\
\# OKX (Futures)                 & BTC-USD-240628     & \$46,744.90                        & 44016.1 USD                              & -5.838\%                           & 0.03\%                              & 6 months                                & \$67,199,400                               & \$13,281,039.45                         \\
\# OKX (Futures)                 & BTC-USDT-240329    & \$45,345.34                        & 43989.3 USDT                             & -2.942\%                           & 0.02\%                              & 3 months                                & \$54,809,841                               & \$32,494,014.76                         \\
\# OKX (Futures)                 & BTC-USDT-240628    & \$46,824.58                        & 43989.3 USDT                             & -6.008\%                           & 0.02\%                              & 6 months                                & \$49,096,426                               & \$8,216,309.33                          \\
\# OKX (Futures)                 & BTC-USD-240126     & \$44,235.20                        & 44016.1 USD                              & -0.495\%                           & 0.04\%                              & 19 days                                 & \$20,945,300                               & \$1,424,116.88                          \\
\# OKX (Futures)                 & BTC-USD-240112     & \$44,085.20                        & 44016.1 USD                              & -0.157\%                           & 0.01\%                              & 5 days                                  & \$15,394,800                               & \$11,575,786.01                         \\
\# OKX (Futures)                 & BTC-USDT-240126    & \$44,284.50                        & 43989.3 USDT                             & -0.617\%                           & 0.06\%                              & 19 days                                 & \$8,573,427                                & \$1,552,614.65                          \\
\# OKX (Futures)                 & BTC-USD-240223     & \$44,719.10                        & 44016.1 USD                              & -1.572\%                           & 0.04\%                              & about 2 months                          & \$8,046,700                                & \$3,603,630.54                          \\
\# OKX (Futures)                 & BTC-USD-240119     & \$44,198.90                        & 44016.1 USD                              & -0.414\%                           & 0.01\%                              & 12 days                                 & \$4,957,200                                & \$6,247,231.64                          \\
\# OKX (Futures)                 & BTC-USDT-240112    & \$44,080.50                        & 43989.3 USDT                             & -0.157\%                           & 0.03\%                              & 5 days                                  & \$3,560,036                                & \$12,798,332.26                         \\
\# OKX (Futures)                 & BTC-USDT-240223    & \$44,781.65                        & 43989.3 USDT                             & -1.720\%                           & 0.08\%                              & about 2 months                          & \$2,960,896                                & \$711,132.65                            \\
\# OKX (Futures)                 & BTC-USDT-240119    & \$44,192.36                        & 43989.3 USDT                             & -0.409\%                           & 0.04\%                              & 12 days                                 & \$975,364                                  & \$4,235,395.39                          \\
\# OKX (Futures)                 & BTC-USDC-240628    & \$46,788.02                        & 44015.4 USDC                             & -5.958\%                           & 0.39\%                              & 6 months                                & -                                          & \$17,320.92                             \\
\# OKX (Futures)                 & BTC-USDC-240119    & \$44,137.13                        & 44020.7 USDC                             & -0.298\%                           & 0.1\%                               & 12 days                                 & -                                          & \$6,285.13                              \\
\# OKX (Futures)                 & BTC-USDC-240329    & \$45,832.15                        & 44020.7 USDC                             & -3.985\%                           & 0.31\%                              & 3 months                                & -                                          & \$1,874.53                              \\
\# OKX (Futures)                 & BTC-USDC-240112    & \$44,107.04                        & 44015.4 USDC                             & -0.242\%                           & 0.13\%                              & 5 days                                  & -                                          & \$9,765.30                              \\
\# XT.COM (Derivatives)          & BTC\_USD\_240329   & \$45,380.70                        & -                                        & -3.586\%                           & 0.01\%                              & 3 months                                & -                                          & \$33,113,000.00                         \\
\# XT.COM (Derivatives)          & BTC\_USD\_240628   & \$46,797.20                        & -                                        & -                                  & 0.01\%                              & 6 months                                & -                                          & \$31,000,200.00                         \\
\# XT.COM (Derivatives)          & BTC\_USDT\_240329  & \$45,369.50                        & -                                        & -                                  & 0.06\%                              & 3 months                                & -                                          & \$20,870,477.00                        
\end{tabular}}
\end{table}

ETH Futures

with 
Total Open Interest $466,932,584
Total 24h Trade Volume $98,238,298

\begin{table}[ht]
\resizebox{\textwidth}{!}{\begin{tabular}{llrrrrrrr}
\hline
\textbf{Exchange}                & \textbf{Symbol}    & \multicolumn{1}{l}{\textbf{Price}} & \multicolumn{1}{l}{\textbf{Index Price}} & \multicolumn{1}{l}{\textbf{Basis}} & \multicolumn{1}{l}{\textbf{Spread}} & \multicolumn{1}{l}{\textbf{Expires in}} & \multicolumn{1}{l}{\textbf{Open Interest}} & \multicolumn{1}{l}{\textbf{24h Volume}} \\
\hline
\hline
\# Binance (Futures)             & ETHUSD\_240329     & \$2,300.60                         & 2228.81 USD                              & -3.121\%                           & 0.01\%                              & 3 months                                & \$72,215,240                               & \$16,326,927.83                         \\
\# Binance (Futures)             & ETHUSD\_240628     & \$2,368.79                         & 2229.03 USD                              & -5.900\%                           & 0.02\%                              & 6 months                                & \$11,183,360                               & \$4,224,925.26                          \\
\# Binance (Futures)             & ETHUSDT\_240329    & \$2,300.46                         & 2227.81 USDT                             & -3.110\%                           & 0.02\%                              & -                                       & \$35,040                                   & \$6,902,127.72                          \\
\# Binance (Futures)             & ETHUSDT\_240628    & \$2,360.43                         & 2227.81 USDT                             & -5.572\%                           & 0.03\%                              & -                                       & \$4,604                                    & \$4,303,853.27                          \\
\# BitMEX (Derivative)           & ETHUSDH24          & \$2,452.60                         & 2228.01 USD                              & -9.157\%                           & 0.02\%                              & 3 months                                & \$4,341,286                                & \$537,210.05                            \\
\# BitMEX (Derivative)           & ETHH24             & \$2,217.23                         & 0.05 XBT                                 & +0.457\%                           & 0.37\%                              & 3 months                                & \$1,575,478                                & \$69,244.08                             \\
\# BTSE (Futures)                & ETH-240628         & \$2,350.05                         & 2367.26 USD                              & +0.732\%                           & 0.75\%                              & 6 months                                & \$2,517,764                                & \$49,845.75                             \\
\# BTSE (Futures)                & ETH-240329         & \$2,297.48                         & 2298.78 USD                              & +0.057\%                           & 0.23\%                              & 3 months                                & \$2,081,180                                & \$49,130.96                             \\
\# Bybit (Futures)               & ETHUSDM24          & \$2,357.45                         & 2229.63 USD                              & -5.422\%                           & 0.05\%                              & 6 months                                & -                                          & \$448,716.00                            \\
\# Bybit (Futures)               & ETHUSDH24          & \$2,287.90                         & 2229.53 USD                              & -2.551\%                           & 0.06\%                              & 3 months                                & -                                          & \$294,007.00                            \\
\# Bybit (Futures)               & ETH-26JAN24        & \$2,243.65                         & 2229.46 USDC                             & -0.668\%                           & 0.05\%                              & 19 days                                 & -                                          & \$286,245.77                            \\
\# Bybit (Futures)               & ETH-12JAN24        & \$2,231.66                         & 2229.51 USDC                             & -0.132\%                           & 0.03\%                              & 5 days                                  & -                                          & \$99,683.25                             \\
\# Bybit (Futures)               & ETH-28JUN24        & \$2,368.66                         & 2229.46 USDC                             & -5.910\%                           & 0.1\%                               & 6 months                                & -                                          & \$97,357.35                             \\
\# Bybit (Futures)               & ETH-19JAN24        & \$2,238.21                         & 2229.51 USDC                             & -0.424\%                           & 0.06\%                              & 12 days                                 & -                                          & \$92,957.40                             \\
\# Bybit (Futures)               & ETH-23FEB24        & \$2,274.64                         & 2229.51 USDC                             & -2.019\%                           & 0.03\%                              & about 2 months                          & -                                          & \$68,610.65                             \\
\# Bybit (Futures)               & ETH-29MAR24        & \$2,297.48                         & 2229.51 USDC                             & -2.993\%                           & 0.03\%                              & 3 months                                & -                                          & \$36,926.81                             \\
\# Bybit (Futures)               & ETH-27SEP24        & \$2,420.24                         & 2229.46 USDC                             & -7.915\%                           & 0.12\%                              & 9 months                                & -                                          & \$30,008.99                             \\
\# Crypto.com Exchange (Futures) & ETHUSD-240628      & \$2,347.68                         & 2224.96 USD                              & -5.227\%                           & 0.24\%                              & -                                       & \$24,573,170                               & \$822.63                                \\
\# Crypto.com Exchange (Futures) & ETHUSD-240126      & \$2,234.13                         & 2224.96 USD                              & -0.410\%                           & 0.14\%                              & -                                       & \$79,732                                   & \$742.18                                \\
\# Crypto.com Exchange (Futures) & ETHUSD-240329      & \$2,286.18                         & 2224.96 USD                              & -2.678\%                           & 0.29\%                              & -                                       & \$58,845                                   & \$895.27                                \\
\# Crypto.com Exchange (Futures) & ETHUSD-240223      & \$2,249.95                         & 2224.92 USD                              & -1.112\%                           & 0.17\%                              & -                                       & \$27,557                                   & \$1,247.37                              \\
\# Deribit                       & ETH-29MAR24        & \$2,296.00                         & 2228.59 USD                              & -2.936\%                           & 0.02\%                              & 3 months                                & \$88,364,731                               & \$2,438,537.07                          \\
\# Deribit                       & ETH-28JUN24        & \$2,363.75                         & 2228.59 USD                              & -5.718\%                           & 0.01\%                              & 6 months                                & \$35,776,949                               & \$1,152,627.40                          \\
\# Deribit                       & ETH-26JAN24        & \$2,247.75                         & 2228.59 USD                              & -0.852\%                           & 0.02\%                              & 19 days                                 & \$34,172,788                               & \$2,694,255.34                          \\
\# Deribit                       & ETH-27SEP24        & \$2,414.50                         & 2228.59 USD                              & -7.700\%                           & 0.02\%                              & 9 months                                & \$22,332,082                               & \$781,745.26                            \\
\# Deribit                       & ETH-12JAN24        & \$2,232.25                         & 2228.59 USD                              & -0.164\%                           & 0.01\%                              & 5 days                                  & \$10,281,174                               & \$1,626,945.44                          \\
\# Deribit                       & ETH-23FEB24        & \$2,273.25                         & 2228.59 USD                              & -1.965\%                           & 0.08\%                              & about 2 months                          & \$10,187,446                               & \$484,690.68                            \\
\# Deribit                       & ETH-27DEC24        & \$2,463.00                         & 2228.59 USD                              & -9.517\%                           & 0.21\%                              & 12 months                               & \$6,513,788                                & \$630,519.38                            \\
\# Deribit                       & ETH-19JAN24        & \$2,235.50                         & 2228.59 USD                              & -0.309\%                           & 0.03\%                              & 12 days                                 & \$1,564,812                                & \$1,319,273.13                          \\
\# HTX Futures                   & ETH\_CQ            & \$2,298.00                         & 2228.66 USD                              & -3.018\%                           & 0.01\%                              & 3 months                                & \$1,130,370                                & \$562,732.95                            \\
\# HTX Futures                   & ETH\_CW            & \$2,233.11                         & 2228.54 USD                              & -0.205\%                           & 0.01\%                              & 4 days                                  & \$49,870                                   & \$481,669.51                            \\
\# HTX Futures                   & ETH\_NW            & \$2,242.45                         & 2228.66 USD                              & -0.615\%                           & 0.35\%                              & 11 days                                 & \$31,200                                   & \$302,572.28                            \\
\# Kraken (Futures)              & FI\_ETHUSD\_240329 & \$2,300.00                         & 2228 USD                                 & -3.130\%                           & 0.11\%                              & 3 months                                & \$4,415,821                                & \$29,062.00                             \\
\# Kraken (Futures)              & FI\_ETHUSD\_240628 & \$2,366.30                         & 2228 USD                                 & -5.845\%                           & 0.25\%                              & 6 months                                & \$3,541,749                                & \$40,752.00                             \\
\# Kraken (Futures)              & FF\_ETHUSD\_240329 & \$2,314.00                         & 2228 USD                                 & -3.717\%                           & 0.13\%                              & 3 months                                & \$1,584                                    & \$0.45                                  \\
\# Kraken (Futures)              & FF\_ETHUSD\_240126 & \$2,254.90                         & 2228 USD                                 & -1.193\%                           & 0.09\%                              & 19 days                                 & \$299                                      & \$36.69                                 \\
\# Kraken (Futures)              & FI\_ETHUSD\_240126 & \$2,254.15                         & 2228 USD                                 & -1.160\%                           & 0.24\%                              & 19 days                                 & \$48,594                                   & \$10,184.00                             \\
\# OKX (Futures)                 & ETH-USD-240329     & \$2,296.78                         & 2228.71 USD                              & -2.964\%                           & 0.01\%                              & 3 months                                & \$61,638,000                               & \$13,362,284.32                         \\
\# OKX (Futures)                 & ETH-USD-240628     & \$2,370.47                         & 2228.67 USD                              & -5.982\%                           & 0.03\%                              & 6 months                                & \$29,831,250                               & \$8,793,848.00                          \\
\# OKX (Futures)                 & ETH-USDT-240329    & \$2,296.43                         & 2226.75 USDT                             & -2.985\%                           & 0.02\%                              & 3 months                                & \$20,412,503                               & \$5,701,567.75                          \\
\# OKX (Futures)                 & ETH-USD-240112     & \$2,231.34                         & 2228.71 USD                              & -0.118\%                           & 0.01\%                              & 5 days                                  & \$7,456,500                                & \$8,340,344.15                          \\
\# OKX (Futures)                 & ETH-USDT-240628    & \$2,366.59                         & 2227.15 USDT                             & -5.845\%                           & 0.02\%                              & 6 months                                & \$3,885,976                                & \$3,157,980.17                          \\
\# OKX (Futures)                 & ETH-USD-240119     & \$2,233.42                         & 2228.71 USD                              & -0.211\%                           & 0.03\%                              & 12 days                                 & \$3,212,320                                & \$3,467,988.47                          \\
\# OKX (Futures)                 & ETH-USDT-240112    & \$2,229.86                         & 2226.75 USDT                             & -0.089\%                           & 0.01\%                              & 5 days                                  & \$2,298,019                                & \$3,571,348.57                          \\
\# OKX (Futures)                 & ETH-USDT-240119    & \$2,235.50                         & 2226.75 USDT                             & -0.341\%                           & 0.01\%                              & 12 days                                 & \$690,474                                  & \$2,364,707.49                          \\
\# OKX (Futures)                 & ETH-USDC-240329    & \$2,296.52                         & 2227.93 USDC                             & -3.020\%                           & 0.38\%                              & 3 months                                & -                                          & \$55,373.71                             \\
\# OKX (Futures)                 & ETH-USDC-240112    & \$2,251.63                         & 2227.89 USDC                             & -1.088\%                           & 0.09\%                              & 5 days                                  & -                                          & \$5,770.92                              \\
\# OKX (Futures)                 & ETH-USDC-240119    & \$2,251.87                         & 2227.93 USDC                             & -1.037\%                           & 0.14\%                              & 12 days                                 & -                                          & \$9.01                                  \\
\# XT.COM (Derivatives)          & ETH\_USD\_240628   & \$2,368.07                         & -                                        & -                                  & 0.04\%                              & 6 months                                & -                                          & \$2,929,170.00                          \\
\# XT.COM (Derivatives)          & ETH\_USD\_240329   & \$2,299.63                         & -                                        & -2.350\%                           & 0.01\%                              & 3 months                                & -                                          & \$2,698,030.00                          \\
\# XT.COM (Derivatives)          & ETH\_USDT\_240329  & \$2,301.93                         & -                                        & -                                  & 0.03\%                              & 3 months                                & -                                          & \$124,645.12                            \\
\# OKX (Futures)                 & BTC-USDT-240112    & \$44,080.50                        & -0.157\%                                 & 0.03\%                             & 5 days                              & \$3,560,036                             & \$12,798,332.26                            & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDT-240223    & \$44,781.65                        & -1.720\%                                 & 0.08\%                             & about 2 months                      & \$2,960,896                             & \$711,132.65                               & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDT-240119    & \$44,192.36                        & -0.409\%                                 & 0.04\%                             & 12 days                             & \$975,364                               & \$4,235,395.39                             & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDC-240628    & \$46,788.02                        & -5.958\%                                 & 0.39\%                             & 6 months                            & -                                       & \$17,320.92                                & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDC-240119    & \$44,137.13                        & -0.298\%                                 & 0.1\%                              & 12 days                             & -                                       & \$6,285.13                                 & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDC-240329    & \$45,832.15                        & -3.985\%                                 & 0.31\%                             & 3 months                            & -                                       & \$1,874.53                                 & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDC-240112    & \$44,107.04                        & -0.242\%                                 & 0.13\%                             & 5 days                              & -                                       & \$9,765.30                                 & \multicolumn{1}{l}{}                    \\
\# XT.COM (Derivatives)          & BTC\_USD\_240329   & \$45,380.70                        & -3.586\%                                 & 0.01\%                             & 3 months                            & -                                       & \$33,113,000.00                            & \multicolumn{1}{l}{}                    \\
\# XT.COM (Derivatives)          & BTC\_USD\_240628   & \$46,797.20                        & -                                        & 0.01\%                             & 6 months                            & -                                       & \$31,000,200.00                            & \multicolumn{1}{l}{}                    \\
\# XT.COM (Derivatives)          & BTC\_USDT\_240329  & \$45,369.50                        & -                                        & 0.06\%                             & 3 months                            & -                                       & \$20,870,477.00                            & \multicolumn{1}{l}{}                   
\end{tabular}}
\end{table}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}

Crypto exchanges often provide a so-called perpetual future, which is a synthetic (typically cash-settled instrument) whose return is tied to the underlying via a funding rate. Most exchanges (quote here) provide perpetual futures for Bitcoin and Ethereum. 
The price of each perpetual future is linked to the underlying by adjusting a variable interest rate, called the funding rate, that fluctuates depending on the deviation between the price of the perpetual future and the underlying. The funding rate increases the cost of holding positions against the deviation, meaning that if the price of the perpetual future is higher than the underlying, the funding rate is positive which causes the market participants who are long to pay the short side in a contractually predefined frequency. Broadly said, the more extreme the deviation, the higher the funding rate and thus the higher the incentive to trade against the deviation and thus converge the price of the perpetual future towards its underlying. 
The risk-free rate is approximated using the simplest and most liquid way of realising the funding rate, a carry trade. In such a trade, the trader shorts (longs) the perpetual future if the funding rate is positive (negative) and simulateously buys (sells) the underlying such that the market delta of the combined position is zero.

The funding rate on \href{https://www.deribit.com/kb/deribit-linear-perpetual}{Deribit} is calculated as follows:

Deribit perpetual contracts are a derivative product similar to a future, however, they do not have an expiry date. The perpetual contract features funding payments. These payments have been introduced to keep the perpetual contract price as close as possible to the underlying crypto price - the relevant Deribit Index. If the perpetual contract trades at a higher price than the index, traders that have long positions need to make funding payments to the traders that have short positions. This will make the product less attractive to the long position holders and more attractive to the short position holders. This subsequently will cause a perpetual price to trade in line with the price of the index. If the perpetual trades at a price lower than the index, the short position holders will have to pay the long position holders instead.

The Deribit perpetual contract features a continuous measurement of the difference between the mark price of the contract and the relevant Deribit Index. The percentage difference between these two price levels is the basis for the 8-hourly funding rate that is applied to all outstanding perpetual contracts.

Funding payments are calculated every millisecond. The funding payments will be added to or subtracted from the realized PNL, which is also part of the available trading balance. At each daily settlement, the realized PNL will be moved to or from the cash balance, from which withdrawals can be made.

The total funding paid will show up in the transaction log in the "funding" column. This column shows the funding amount that is applied to the trader's entire net position in the period between the relevant trade and the trade before that. Put differently: the trader can see the funding paid or received on the position between position changes.

Funding Rate
When the funding rate is positive, long position holders pay funding to the short position holders; when the funding rate is negative, short position holders pay funding to the long position holders. The funding rate is expressed as an 8-hour interest rate, and is calculated at any given time as follows:

Premium Rate
Premium Rate = ((Mark Price - Deribit Index) / Deribit Index) * 100%

Funding Rate
Sequentially, the funding rate is derived from the premium rate by applying a damper.

If the premium rate is within -0.025\% and 0.025\% range, the actual funding rate will be reduced to 0.00\%.
If the premium rate is lower than -0.025\%, then the actual funding rate will be the premium rate + 0.025\%.
If the premium rate is higher than 0.025\%, then the actual funding rate will be the premium rate - 0.025\%.
Additionally, the funding rate is capped at +/ - 5\% for all linear USDC perpetuals.
Funding Rate = Maximum (0.025\%, Premium Rate) + Minimum (-0.025\%, Premium Rate)

Time Fraction
Time Fraction = Funding Rate Time Period / 8 hours

The actual funding payment is calculated by multiplying the funding rate by the position size and the time fraction.

Funding Payment = Funding Rate * Position Size * Time Fraction

In reality, the spread of the Deribit Index and the mark price changes continuously, and all changes are taken into account. Therefore, the examples above are extreme simplifications of the actual calculations. The funding paid or received is continuously added to the realized PNL and is moved to or from the cash balance at the daily settlement, at 08:00 UTC.

Fees on Funding
Deribit does not charge any fees on funding. All the funding payments are transferred between the holders of the perpetual contracts. This makes the funding a zero-sum game, where longs receive all funding from shorts, or shorts receive all funding from longs.

Mark Price
It is essential to understand how the mark price is calculated. We start by determining the "Fair Price". The fair price is calculated as the average of the fair impact bid and the fair impact ask.

Fair Impact Bid Price is an average price at which 100,000 USDC worth of contracts can be bought.

Fair Impact Ask Price is an average price at which 100,000 USDC worth of contracts can be sold.

Fair Price can never be higher than 0.1 % of best ask or lower than best bid

Fair Price = (Fair Impact Bid + Fair Impact Ask) / 2
The mark price is derived using both the Deribit Index and the fair price, by adding to the Deribit Index the 30 second exponential moving average (EMA) of the Fair Price - Deribit Index.

Mark Price = Deribit Index + 30 second EMA (Fair Price - Deribit Index)
We implement a safety setting on our mark price to not deviate by more than a certain percentage of Index. This is to prevent manipulation or to curb the impact of panic trading. When a deviation from index is genuine, then we would increase the possible deviation.

Trading outside of this bandwidth is still allowed.

The 30 second EMA is recalculated every second, so in total, there are 30 time periods in which the measurement of the latest second has a weight of 2 / (30 + 1) = 0.0645 or (6.45%).

Allowed Trading Bandwidth
Two parameters act as a boundary to the allowed trading range:

Perpetual trades are limited by Deribit Index + 1 minute EMA (Fair Price - Index) +/- 1.5%, and a fixed bandwidth of the Deribit Index of +/- 7.5%.

If the market circumstances require so, bandwidth parameters can be adjusted at the sole discretion of Deribit.

Solve for interest rates from

\begin{equation}
F_t = e^{r \tau} S_t
\end{equation}

funding_btc_vs_eth_annualized_zoom.png
funding_btc_vs_eth_zoom.png
funding_btc_vs_eth_annualized.png
funding_btc_vs_eth.png

\section{Results}

We find the returns of the zero-beta-straddles to be t-distributed and conduct a left-sided t-test under the null hypothesis of a population mean greater or equal to zero. 
The null hypothesis is not rejected for all models, vanilla and crash-resistant, for ITM and OTM options. Conversely, the null hypothesis is firmly rejected on a high significance level for ATM options. 
Therefore, we accept the alternative hypothesis of an underperformance for ATM options. Consequently, we find a similar result to Shumway and Bakshi - that is the existence of a premium for ATM contracts. 
We figure the premium to be a volatility premium since a straddle's payoff depends on the relationship of Implied Volatility to Realized Volatility. 
Unsurprisingly, the size of the premium in crypto markets is higher than the counterpart in equity markets - we figure that the reason is the higher risk in crypto markets, particularly via volatility and counterparty risk. 
Furthermore, there are mostly retailers present in the crypto option market. Some professional liquidity providers exist, but they remain few and therefore another reason for the premium's size could be illiquidity.

- the premium is larger when tau is small.

\begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/plots/funding_btc_vs_eth_annualized.png}
    \captionof{figure}{Annualized Funding Rate \color{blue}{BTC} vs. \color{orange}{ETH}}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/plots/funding_btc_vs_eth_annualized_zoom.png}

     \captionof{figure}{(Zoomed) Annualized Funding Rate \color{blue}{BTC} vs. \color{orange}{ETH}}

\end{minipage}
\begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/plots/funding_btc_vs_eth.png}
    \captionof{figure}{Funding Rate \color{blue}{BTC} vs. \color{orange}{ETH}}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/plots/funding_btc_vs_eth_zoom.png}
    \captionof{figure}{(Zoomed) Funding Rate \color{blue}{BTC} vs. \color{orange}{ETH}}

\end{minipage}
   % \captionof{figure}{ATM Term Structure from 2022-01-01 until 2022-01-08. Moneyness $M \in [0.9, 1.1]$}

\label{ATM Term Structure}
\href{https://github.com/QuantLet/BitcoinOptions/blob/master/main.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,Main}

\pagebreak
\newpage

\begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/plots/vanilla/no_fees/boxplot_combined_ret_daily_over_moneyness_.png}
    \captionof{figure}{Performance - Vanilla - No Fees}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/plots/vanilla/fees/boxplot_combined_ret_daily_over_moneyness_.png}
     \captionof{figure}{Performance - Vanilla - Fees}

\end{minipage}
\begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/plots/crash_resistant/no_fees/boxplot_combined_ret_daily_over_moneyness_.png}
    \captionof{figure}{Performance - Crash Resistant - No Fees}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/plots/crash_resistant/fees/boxplot_combined_ret_daily_over_moneyness_.png}
     \captionof{figure}{Performance - Crash Resistant - Fees}

\end{minipage}
   % \captionof{figure}{ATM Term Structure from 2022-01-01 until 2022-01-08. Moneyness $M \in [0.9, 1.1]$}

\label{ATM Term Structure}
\href{https://github.com/QuantLet/BitcoinOptions/blob/master/main.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,Main}

We conduct a left-sided t-test to test for a population mean of zero. Values are rounded to 4 digits.

\section{Conclusion}

Further research: 
- Evidence for leverage effect

%\begin{minipage}{\textwidth}
%    \centering
  %  \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/out/vanilla/no_fees/density/moneyness_between_95_and_105.png}
    %\captionof{figure}{Performance - Vanilla - No Fees}
    %\centering
    %\includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/out/vanilla/no_fees/density/moneyness_between_95_and_105.png}
     %\captionof{figure}{Performance - Vanilla - Fees}

%\end{minipage}
%\begin{minipage}{\textwidth}
%    \centering
%    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/out/crash_resistant/no_fees/moneyness_between_95_and_105.png}
%    \captionof{figure}{Performance - Crash Resistant - No Fees}
 %   \centering
 %   \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/out/crash_resistant/no_fees/moneyness_between_95_and_105.png}
 %    \captionof{figure}{Performance - Crash Resistant - Fees}

%\end{minipage}
   % \captionof{figure}{ATM Term Structure from 2022-01-01 until 2022-01-08. Moneyness $M \in [0.9, 1.1]$}

%\label{ATM Term Structure}
%\href{https://github.com/QuantLet/BitcoinOptions/blob/master/main.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,Main}

\begin{table}[ht]
\hline
\resizebox{\textwidth}{!}{\begin{tabular}{lllll}
  Moneyness    & Vanilla ex Fee & Vanilla cum Fee & Crash Resistant ex Fee & Crash Resistant cum Fee \\
\hline
\hline
0.7  & 0.6088         & 0.5854          & 0.5845                 & 0.6078                  \\
0.71 & 0.7993         & 0.7769          & 0.7762                 & 0.7987                  \\
0.72 & 0.9943         & 0.9925          & 0.9925                 & 0.9943                  \\
0.73 & 0.9348         & 0.9222          & 0.9222                 & 0.9348                  \\
0.74 & 0.9997         & 0.9996          & 0.9996                 & 0.9997                  \\
0.75 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.76 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.77 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.78 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.79 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.8  & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.81 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.82 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.83 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.84 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.85 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.86 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.87 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.88 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.89 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.9  & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.91 & 1.0            & 0.9998          & 0.9998                 & 1.0                     \\
0.92 & 1.0            & 0.9968          & 0.9968                 & 1.0                     \\
0.93 & 0.9984         & 0.9437          & 0.9437                 & 0.9984                  \\
0.94 & 0.9669         & 0.6277          & 0.6277                 & 0.9669                  \\
0.95 & 0.8817         & 0.3093          & 0.3093                 & 0.8817                  \\
0.96 & 0.3229         & 0.0129          & 0.0129                 & 0.3229                  \\
0.97 & 0.2143         & 0.0047          & 0.0047                 & 0.2143                  \\
0.98 & 0.0082         & 0.0             & 0.0                    & 0.0082                  \\
0.99 & 0.0002         & 0.0             & 0.0                    & 0.0002                  \\
1.0  & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.01 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.02 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.03 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.04 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.05 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.06 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.07 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.08 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.09 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.1  & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.11 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.12 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.13 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.14 & 0.0172         & 0.0106          & 0.0106                 & 0.0172                  \\
1.15 & 0.0484         & 0.0347          & 0.0347                 & 0.0484                  \\
1.16 & 0.0563         & 0.0421          & 0.0421                 & 0.0564                  \\
1.17 & 0.1889         & 0.1588          & 0.1587                 & 0.1887                  \\
1.18 & 0.8503         & 0.8119          & 0.8115                 & 0.8499                  \\
1.19 & 0.9906         & 0.9861          & 0.9859                 & 0.9905                  \\
1.2  & 0.8888         & 0.8593          & 0.8562                 & 0.8862                  \\
1.21 & 0.9796         & 0.9713          & 0.9701                 & 0.9787                  \\
1.22 & 0.986          & 0.9805          & 0.9793                 & 0.985                   \\
1.23 & 0.9627         & 0.951           & 0.9464                 & 0.959                   \\
1.24 & 0.9969         & 0.9954          & 0.9945                 & 0.9962                  \\
1.25 & 0.9976         & 0.9966          & 0.9957                 & 0.9969                  \\
1.26 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
1.27 & 0.9999         & 0.9999          & 0.9998                 & 0.9999                  \\
1.28 & 0.9996         & 0.9995          & 0.9993                 & 0.9995                  \\
1.29 & 0.9939         & 0.9932          & 0.9919                 & 0.9927                  \\
1.3  & 0.9436         & 0.9403          & 0.9334                 & 0.937                  
\end{tabular}}
\end{table}

\pagebreak
\newpage

NOTE: 
USE THE TRADED VOLUME * INSTRUMENT PRICE TO FIND THE AMOUNT OF VALUE LOST DUE TO THE VOLA PREMIUM

\section{Deribit}

Deribit, a margin-trading platform for futures and options, has been
launched in June 2016 in the Netherlands and is currently incorporated
in Panama. As of time of the writing, Deribit is the largest BTC option
exchange (Skew.com 2022). For both types of derivatives, BTC and ETH are
the underlying as well as the currency in which settlement is conducted.
This allows to view them as inverse options.

\begin{center}
\begin{figure}
\includegraphics[width=0.9\textwidth]{/Users/julian/src/up/spd/plot/Skew_BTC_Open_Interest_cut.pdf}
\caption[Open Interest in BTC Options per Exchange.]%
{Open Interest in BTC Options per Exchange. Snapshot from \href{https://analytics.skew.com/dashboard/bitcoin-options}{skew.com} on 2022-01-16.  \color{bittersweet}{Deribit}, \color{ao}{LedgerX}, \color{bananamania}{CME}, \color{bluegray}{bit.com}. }
\label{fig: ivtest}
\end{figure}
\end{center}

\subsection{Bitcoin Option Contract Specifications}

Deribit offers cash-settled European-style options. The underlying is a
synthetic index, whose exact composition is described in the data
section. Settlement is first calculated in USD and then conducted in
kind of the underlying. E.g. a BTC call with a strike of \(10000\) and a
settlement value of \(12000\) at maturity will result in a cashflow of
\(2000\) USD, which is equivalent to \(\frac{1}{6}\) BTC (excluding
transaction cost). The settlement value is defined as the average of the
underlying BTC index for the last 30 minutes before settlement. Each
contract has a lot size of 1 BTC and is settled at 8am UTC on the
respective maturity date. Maturity dates have a daily, monthly,
quarterly and annual frequency. Deribit is the only derivatives exchange
to offer daily options. Instruments are trading 24/7 with a tick size of
0.0005 BTC. Due to the automatic usage of margin-trading, margin-based
liquidation of positions is possible. Initial margin and maintenance
margins are both zero for long positions. For short positions, initial
margin is required (Deribit 2022b). In case of liquidation, a penalty is
applied to the defaulting party, whose proceeds are paid into the
Deribit Insurance Fund (Deribit 2020a).

Trading and deliveries on Deribit are subject to fees (Deribit 2022c).
The applied fees generally vary depending on whether the order was a
liquidity maker or taker. However, for Bitcoin options they are equal.
Trading fees are 0.03\% of the underlying or 0.0003 BTC per options
contract. The fees constitute a maximum of 12.5\% of the contract's
value.

\section{Data}

\subsection{Data Structure}

The dataset consists of 8,444,664 order book snapshots, which have been
collected from Deribit in the time from 2021-04-01 until 2022-04-01. The
snapshots were captured via the Deribit API V2. The data base is
available on the
\href{https://blockchain-research-center.com}{Blockchain Research Center (BRC)}.
The full data set includes all parameters that the
\href{https://docs.deribit.com/#deribit-api-v2-1-1}{Deribit API V2}
returns at the time of collection under the methods

\begin{itemize}
\tightlist
\item
  \{\it public/get\_last\_trades\_by\_instrument\_and\_time\}
\item
  \{\it public/get\_order\_book\}
\end{itemize}

Most importantly the results include

\begin{itemize}
\tightlist
\item
  Timestamps
\item
  Greeks
\item
  Implied Volatility
\item
  Tick Direction
\item
  Order Type
\item
  Volume
\item
  Instrument Price
\item
  Strike
\item
  Spot
\end{itemize}

High-frequent order book changes and executed trades are captured for
options and futures whose underlying is BTC. Values of the underlying
synthetic BTC USD Index are also saved. The underlying is calculated as
an equally-weighted BTC/USD price of eleven major crypto exchanges,
namely Binance, Bitfinex, Bitstamp, Bittrex, Coinbase Pro, FTX, Gemini,
Huobi, Itbit, Kraken, LMAX Digital, OKEx. Individual feeds can be
excluded due to administrative decisions or in case if invalid data.
Remaining feeds are sorted, truncated around the median price and
weighted (Deribit 2020b).

\subsection{Data Integrity}

Data integrity plays a crucial role in crypto markets. Mark Carney, the
chair of the Financial Stability Board and head of the Bank of England,
warned that wash trading, pump and dumps and spoofing, known as outlawed
manipulation techniques in equity markets, are also present in crypto
and pose a risk to financial stability (Rodgers 2019). Wash trading is
used to increase trading volume and thus artificially increase demand. A
spoofer submits market non-bona fide price quotes in order to cause
artificial price volatility (Sar 2017). Pump \& Dump is a form of
securities fraud in which a group of traders rapidly and artificially
inflate a price in order to offload their previously acquired inventory.
Since Pump \& Dump schemes require low liquidity (in the underlying), we
can exclude this possibility for Bitcoin (La Morgia et al. 2020).

Aloosh and Li (2019) find evidence for exchange-driven inflated volume,
generated by wash trades, to market themselves under the guise of
liquidity. Cong et al. (2019) report that on average 70\% of volume on
decentralized exchanges is fake due to the use of wash trading.
According to Bitmex (2019), up to 95\% of trading volume on unregulated
exchanges is generated by wash trading. In order to ensure data
integrity, we exclusively use orderbooks instead of executed trades in
order to filter possible cases of wash trading.

To address the potential issue of spoofing, we adopt an approach related
to Tuccella et al. (2021). They attempt to identify spoofing on
cryptocurrency exchanges using a GRU model. Their predictors are a
function of the amount of cancelled orders relative to the cumulative
order size on each side of the order book. Since Bitcoin derivatives
order books do not have the same amount of depth as spot or futures
markets, we employ a stricter method and exclude all orders whose
lifetime does not exceed two seconds. We find that less than 5\% of
orders are filtered due to short order lifetime. We figure that
traditional spoofing techniques, based on the rapid submission and
cancellation of orders, are difficult to implement due to Deribit's rate
limits (Deribit 2023). Users with less than 1 million USD in
7-day-turnover can only post five matching engine requests. These
include request for a buy or sell order or a corresponding cancellation.
Although the limits for a market maker (who could be allowed to post 30
requests per second given a turnover of 25 million USD in 7 days) are
considerably higher, the market maker must still meet the quoting
requirements for a large range of instruments. This effectively
restricts the influence that single entities can have on certain
instruments via the rapid submission and cancellation of orders and
explains the rarity of such events in our dataset.

\subsection{Preprocessing}

Although market makers are obliged to quote most instruments for the
majority of the time, some may not be quoted at all or only be quoted at
the cost of a large spread (Deribit 2020c). For Far-Out-of-Money (FOTM)
contracts, when the minimum tick size exceeds the market IV, one can
observe bids for 0\% IV. Cases of quotes for 0\% IV and all duplicates
are excluded from the data set. Since the majority of option trading
volume concentrates on instruments with short or medium time-to-maturity
\(\tau\), we restrict \(\tau\) to be smaller or equal than one quarter
of a year. \(\tau\) is normalized on the span of a year. Instruments are
grouped according to their time-to-maturity, over a range of 0, 1, 2, 4,
8 weeks. We restrict moneyness \(M = \frac{K}{S}\) to the interval
\([0.7, 1.3]\) in order to exclude the influence of high-volatility
observations which could be unreliable (Grith et al. 2009).

Call options are exclusively used in order to estimate the IV surface in
Rookley's method. Put-Call-Parity ensures arbitrage-free call option
prices. However, we do not use Put-Call-Parity to increase our data
sample. Conversion of option prices via Put-Call-Parity could introduce
errors into our dataset due to market microstructure (MMN) noise. Our
high-frequent order book data is extracted via an API. Due to rate
limitations and a diverse amount of instruments, order book snapshots
cannot be taken simultaneously for all instruments. The data scraping
application successively iterates over all instruments and captures
changes in order books. While the resulting data base provides a clear
picture of order book data, it does not allow for a complete
reconstruction of all order books at all times. Without simultaneity,
volatility of the underlying and order competition within the spread
could change the prices of the corresponding instruments (required for
Put-Call-Parity) between the individual snapshots. A range of futures
may be available on Deribit, but only the perpetual futures contract is
actually liquid. The classic (non-perpetual) futures also price in the
settlement cost, which often causes them to trade in backwardation close
to maturity. But in spite of this, BTC options tend to be traded more
frequently when they are close to maturity. The combination of these
real-world limitations challenges the simple conversion via
Put-Call-Parity and could introduce errors into our dataset.
Furthermore, we do not see the necessity to use such a method as we have
sufficient call option data to compute the IV surface from which we
sample option prices.

\href{https://github.com/QuantLet/BitcoinOptions/blob/master/main.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,Preprocessing}

\section{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-aitsahalia01optionprice}{}%
Aı̈t-Sahalia, Y., Wang, Y., \& Yared, F. (2001). Do option markets
correctly price the probabilities of movement of the underlying asset?
\emph{Journal of Econometrics}, \emph{102}(1), 67--110.
\url{https://doi.org/https://doi.org/10.1016/S0304-4076(00)00091-9}

\leavevmode\hypertarget{ref-ait_sahalia_wang1998}{}%
Aït-Sahalia, Y., \& Lo, A. W. (1998). Nonparametric estimation of
state-price densities implicit in financial asset prices. \emph{The
Journal of Finance}, \emph{53}(2), 499--547.
\url{https://doi.org/https://doi.org/10.1111/0022-1082.215228}

\leavevmode\hypertarget{ref-alexander2022net}{}%
Alexander, C., Deng, J., Feng, J., \& Wan, H. (2022). Net buying
pressure and the information in bitcoin option trades. \emph{Journal of
Financial Markets}, 100764.

\leavevmode\hypertarget{ref-inverse_options}{}%
Alexander, C., \& Imeraj, A. (2021, July). Inverse options in a
black-scholes world.
\url{https://doi.org/https://doi.org/10.48550/arXiv.2107.12041}

\leavevmode\hypertarget{ref-aloosh}{}%
Aloosh, A., \& Li, J. (2019). Direct evidence of bitcoin wash trading.
\emph{Theory of Probability \& Its Applications, 5, 285-301}.
\url{https://doi.org/http://dx.doi.org/10.2139/ssrn.3362153}

\leavevmode\hypertarget{ref-bitmex}{}%
Bitmex. (2019). Presentation to the u.s. Securities and exchange
commission on wash trading.
\url{https://www.sec.gov/comments/sr-nysearca-2019-01/srnysearca201901-5164833-183434.pdf}.
Accessed 18 January 2022

\leavevmode\hypertarget{ref-blase}{}%
Blaskowitz, O., Härdle, W. K., \& Schmidt, P. (2004). Skewness and
kurtosis trades. In (pp. 1--14).
\url{https://doi.org/10.1007/978-0-8176-8180-7_1}

\leavevmode\hypertarget{ref-speculation}{}%
Blau, B. M. (2017). Price dynamics and speculative trading in bitcoin.
\emph{Research in International Business and Finance}, \emph{41},
493--499.
\url{https://doi.org/https://doi.org/10.1016/j.ribaf.2017.05.010}

\leavevmode\hypertarget{ref-bloomberg-retail-share}{}%
Bloomberg. (2021, November 17). Retail traders slide back below 20\% of
market's total volume.
\url{https://www.bloomberg.com/news/articles/2021-11-17/retail-traders-retreat-as-choppy-markets-challenge-easy-profits}.
Accessed 24 August 2021

\leavevmode\hypertarget{ref-breeden_litzenberger78}{}%
Breeden, D. T., \& Litzenberger, R. H. (1978). Prices of
state-contingent claims implicit in option prices. \emph{The Journal of
Business}, \emph{51}(4), 621--651.
\url{http://www.jstor.org/stable/2352653}

\leavevmode\hypertarget{ref-coins_with_benefits}{}%
Chen, C. Y., \& Vinogradov, D. (2021). Coins with benefits: On
existence, pricing kernel and risk premium of cryptocurrencies.
\emph{Available at SSRN 3959984}.
\url{https://doi.org/http://dx.doi.org/10.2139/ssrn.3864578}

\leavevmode\hypertarget{ref-choi}{}%
Choi, S., \& Shin, J. (2022). Bitcoin: An inflation hedge but not a safe
haven. \emph{Finance Research Letters}, \emph{46}, 102379.
\url{https://doi.org/https://doi.org/10.1016/j.frl.2021.102379}

\leavevmode\hypertarget{ref-oil_spd}{}%
Christoffersen, P., Jacobs, K., \& Pan, X. (. (2021). The State Price
Density Implied by Crude Oil Futures and Option Prices. \emph{The Review
of Financial Studies}, \emph{35}(2), 1064--1103.
\url{https://doi.org/10.1093/rfs/hhab011}

\leavevmode\hypertarget{ref-cryptonews_bitcoin_retail_share}{}%
Coinbase, S., Cryptonews. (2022). Institutions \& retail compete for
bitcoin - whose hands are stronger?
\url{https://cryptonews.com/exclusives/institutions-retail-compete-for-bitcoin-which-is-the-biggest-9695.htm}.
Accessed 22 March 2022

\leavevmode\hypertarget{ref-decentralized_wash_trading}{}%
Cong, L., Li, X., Tang, K., \& Yang, Y. (2019). Crypto wash trading.
\emph{SSRN Electronic Journal}.
\url{https://doi.org/10.2139/ssrn.3530220}

\leavevmode\hypertarget{ref-jackwerth_pk}{}%
Cuesdeanu, H., \& Jackwerth, J. C. (2018). The pricing kernel puzzle:
survey and outlook. \emph{Annals of Finance}, \emph{14}(3), 289--329.
\url{https://doi.org/10.1007/s10436-017-0317-9}

\leavevmode\hypertarget{ref-cross_border_payments}{}%
Deng, Q. (2020). Application analysis on blockchain technology in
cross-border payment. In.
\url{https://doi.org/10.2991/aebmr.k.200306.050}

\leavevmode\hypertarget{ref-deribit-insurance-fund}{}%
Deribit. (2020a). Deribit insurance fund.
\url{https://www.deribit.com/main\#/insurance}. Accessed 4 January 2021

\leavevmode\hypertarget{ref-synthetic-btc-index}{}%
Deribit. (2020b). Deribit btc-usd index.
\url{https://www.deribit.com/main\#/indexes}. Accessed 2 January 2021

\leavevmode\hypertarget{ref-market-maker-obligations}{}%
Deribit. (2020c). Deribit market maker obligations.
\url{https://www.deribit.com/pages/docs/options}. Accessed 2 January
2021

\leavevmode\hypertarget{ref-deribit_retail_piechart}{}%
Deribit. (2022a). Deribit volume share.
\url{https://www.deribit.com/statistics/BTC/market-data}. Accessed 25
July 2022

\leavevmode\hypertarget{ref-deribit_margin}{}%
Deribit. (2022b). Deribit margin requirements.
\url{https://legacy.deribit.com/pages/docs/options}. Accessed 2 May 2022

\leavevmode\hypertarget{ref-deribit_fees}{}%
Deribit. (2022c). Deribit margin requirements.
\url{https://www.deribit.com/kb/fees}. Accessed 2 May 2022

\leavevmode\hypertarget{ref-deribit-rate-limit}{}%
Deribit. (2023). Deribit rate limits.
\url{https://www.deribit.com/kb/deribit-rate-limits}. Accessed 16 April
2021

\leavevmode\hypertarget{ref-vola_premium}{}%
Eraker, B. (2021). The volatility premium. \emph{Quarterly Journal of
Finance (QJF)}, \emph{11}(03), 1--35.
\url{https://EconPapers.repec.org/RePEc:wsi:qjfxxx:v:11:y:2021:i:03:n:s2010139221500142}

\leavevmode\hypertarget{ref-networkeffect}{}%
Gandal, N., \& Halaburda, H. (2016). Can we predict the winner in a
market with network effects? Competition in cryptocurrency market.
\emph{Games}, \emph{7}(3). \url{https://doi.org/10.3390/g7030016}

\leavevmode\hypertarget{ref-Grith09sim}{}%
Grith, M., Härdle, W. K., \& Park, J. (2009). Shape invariant modelling
pricing kernels and risk aversion.
\url{https://doi.org/10.2139/ssrn.2894289}

\leavevmode\hypertarget{ref-hall_1991}{}%
Hall, P. (1991). Edgeworth expansions for nonparametric density
estimators, with applications. \emph{Statistics}, \emph{22}(2),
215--232. \url{https://doi.org/10.1080/02331889108802305}

\leavevmode\hypertarget{ref-hayes}{}%
Hayes, A. S. (2019). Bitcoin price and its marginal cost of production:
Support for a fundamental value. \emph{Applied Economics Letters},
\emph{26}(7), 554--560.
\url{https://doi.org/10.1080/13504851.2018.1488040}

\leavevmode\hypertarget{ref-w2yo}{}%
Härdle, W. K., Okhrin, Y., \& Wang, W. (2014). Uniform Confidence Bands
for Pricing Kernels. \emph{Journal of Financial Econometrics},
\emph{13}(2), 376--413. \url{https://doi.org/10.1093/jjfinec/nbu002}

\leavevmode\hypertarget{ref-pricing_cryptocurrency_options}{}%
Hou, A. J., Wang, W., Chen, C. Y. H., \& Härdle, W. K. (2020). Pricing
Cryptocurrency Options*. \emph{Journal of Financial Econometrics},
\emph{18}(2), 250--279. \url{https://doi.org/10.1093/jjfinec/nbaa006}

\leavevmode\hypertarget{ref-xfg_spd}{}%
Huynh, K., Kervella, P., \& Zheng, J. (2002). \emph{Estimating
state-price densities with nonparametric regression} (SFB 373 Discussion
Papers No. 2002,40). Humboldt University of Berlin, Interdisciplinary
Research Project 373: Quantification; Simulation of Economic Processes.
\url{https://EconPapers.repec.org/RePEc:zbw:sfb373:200240}

\leavevmode\hypertarget{ref-pumpndump}{}%
La Morgia, M., Mei, A., Sassi, F., \& Stefa, J. (2020). Pump and dumps
in the bitcoin era: Real time detection of cryptocurrency market
manipulations. In \emph{2020 29th international conference on computer
communications and networks (icccn)} (pp. 1--9).
\url{https://doi.org/10.1109/ICCCN49398.2020.9209660}

\leavevmode\hypertarget{ref-vola_stylized_facts}{}%
Masset, P. (2011). Volatility stylized facts. \emph{SSRN Electronic
Journal}. \url{https://doi.org/10.2139/ssrn.1804070}

\leavevmode\hypertarget{ref-satoshi2009}{}%
Nakamoto, S. (2009). Bitcoin: A peer-to-peer electronic cash system.
\url{http://www.bitcoin.org/bitcoin.pdf}

\leavevmode\hypertarget{ref-fundamental-theorem-of-asset-pricing}{}%
Pascucci, A., \& Agliardi, R. (2011). PDE and martingale methods in
option pricing. In (pp. 429--495).

\leavevmode\hypertarget{ref-rodgers2019}{}%
Rodgers, T. (2019, April 4). 95 percent of volume could be wash trading
as bitcoin price surges.
\url{https://www.forbes.com/sites/tomrodgers1/2019/04/04/99-of-volume-could-be-wash-trading-as-bitcoin-takes-back-5000/?sh=4640bcbc23d5}.
Accessed 15 April 2023

\leavevmode\hypertarget{ref-Rookley97}{}%
Rookley, C. (1997). Fully exploiting the information content of intra
day option quotes: Applications in option pricing and risk management.

\leavevmode\hypertarget{ref-sar2017}{}%
Sar, M. (2017). Dodd-frank and the spoofing prohibition in commodities
markets.

\leavevmode\hypertarget{ref-deriv_volume}{}%
Skew.com. (2022). Open interest for bitcoin options over time. Skew is a
leading crypto data analytics platform for institutional clients.
\url{https://analytics.skew.com/dashboard/bitcoin-options}. Accessed 25
July 2022

\leavevmode\hypertarget{ref-brazilian_speculation}{}%
Smaniotto, E. N., \& Neto, G. B. (2022). Speculative trading in bitcoin:
A brazilian market evidence. \emph{The Quarterly Review of Economics and
Finance}, \emph{85}, 47--54.
\url{https://doi.org/https://doi.org/10.1016/j.qref.2020.10.024}

\leavevmode\hypertarget{ref-cross_border_stellar}{}%
Stellar Development Foundation. (2022). 2022 research reveals high
awareness and growing cross-border use of cryptocurrency in four key
markets.
\url{https://www.prnewswire.com/news-releases/2022-research-reveals-high-awareness-and-growing-cross-border-use-of-cryptocurrency-in-four-key-markets-301512401.html}.
Accessed 22 July 2022

\leavevmode\hypertarget{ref-nickszabo}{}%
Szabo, N. (2008). Bit gold.
\url{https://unenumerated.blogspot.com/2005/12/bit-gold.html}. Accessed
28 July 2022

\leavevmode\hypertarget{ref-tuccella2021}{}%
Tuccella, J.-N., Nadler, P., \& Șerban, O. (2021). Protecting retail
investors from order book spoofing using a gru-based detection model.

\end{document}