### 2024-02-13 written/paper.tex
writing until risk free rate section

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12,
]{article}
\usepackage{lmodern}
\usepackage[table,xcdraw]{xcolor}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Empirical Option Returns and the risk-free Rate in Crypto Asset Markets},
  pdfauthor={Julian Winkel; Wolfgang K. Härdle},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin = 2.5cm]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{float}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{subfig}
\usepackage{caption}
\usepackage{eucal}
\floatplacement{figure}{H}

\title{Empirical Option Returns and the risk-free Rate in Crypto Asset Markets}
\author{Julian Winkel\footnote{International Research Training Group 1792;
  Humboldt-Universität zu Berlin,
  \href{mailto:julian.winkel@hu-berlin.de}{\nolinkurl{julian.winkel@hu-berlin.de}}} \and Wolfgang K. Härdle\footnote{BRC Blockchain Research Center,
  Humboldt-Universität zu Berlin, Berlin, Germany; Sim Kee Boon
  Institute, Singapore Management University, Singapore; WISE Wang Yanan
  Institute for Studies in Economics, Xiamen University, Xiamen, China;
  Dept. Information Management and Finance, National Chiao Tung
  University, Hsinchu, Taiwan, ROC; Dept. Mathematics and Physics,
  Charles University, Prague, Czech Republic, Grants -- DFG IRTG 1792
  and the Czech Science Foundation's grant no. 19-28231X / CAS XDA
  23020303 gratefully acknowledged,
  \href{mailto:haerdle@hu-berlin.de}{\nolinkurl{haerdle@hu-berlin.de}}}}
\date{12 February 2024}

\begin{document}
\maketitle

\newcommand{\quantnet}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,}
\newcommand{\quantinar}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/quantinar_logo.png}}\,}

\definecolor{bittersweet}{rgb}{1.0, 0.44, 0.37} 
\definecolor{ao}{rgb}{0.0, 0.5, 0.0}
\definecolor{bananamania}{rgb}{0.98, 0.91, 0.71}
\definecolor{bluegray}{rgb}{0.4, 0.6, 0.8}

\pagenumbering{roman}

\begin{abstract}

\singlespacing 
abstract

 \medskip
 \textbf{Keywords:} Bitcoin, Ether, Deribit, risk, risk premium, volatility premium, option, hedging  \\
 \indent \textbf{JEL Classification:} C14, C50, G10  \\
 \bigskip

\end{abstract}

\pagenumbering{arabic}

\newpage

\section{Introduction}
Despite the ongoing integration into established markets (whereas the last example would be the SEC's approval of a range of Bitcoin Spot ETFs), fundamental questions about crypto asset markets remain unclear:
How efficient are crypto asset markets relative to established, traditional markets such as equities? 
Does a risk-free rate exist and if yes, how does it relate to the risk-free rate in established markets? Are derivatives, written on crypto assets, fairly priced or at least fairly priced relative to other derivatives? 
We assess these fundamental questions and find that the methods required to conduct our analysis have been readily available; in fact we are employing similar methods to those, under which the efficiency of equity option markets had been assessed in their "youth" in the early 2000s. 
The common key behind the employed methods is the precise and bespoke tailoring of risk-return characteristics through the use of options, as proposed by Shumway and Bakshi et al in (@quote).

We discuss the existence of a risk-free rate in crypto asset markets and derive a proxy. The proxy serves as a benchmark that beta-neutral portfolios should yield. 
Subsequently, we construct and evaluate the performance of beta-neutral option portfolios using a unique data set of executed transactions from Derbit, the largest Crypto Options Exchange.  We find that the beta-neutral option portfolios, written on Bitcoin and Ether, are fairly priced for most levels of moneyness. However, the portfolios do not generate the appropriate theoretical returns for at-the-money instruments. We compare the risk-return characteristics to the late 90s equity markets and find, in a similar fashion as the results reported by Bakshi et al and Shumway et al., the existence of high premia.for at-the-money options. The effect holds after compensating for tail risk (market crash insurance) and transaction cost.

%
%We construct and evaluate the empirical performance of option portfolios, written on crypto assets such as Bitcoin and Ether, to allow model-free insights into the existence and degree of a leverage effect, the existence and size of risk premia, resulting market inefficiencies (measurable in economic loss / money lost) and to explore the option (pun intended)/possibility of curing named inefficiencies under transaction cost and liquidity constraints. 
%
%Another, so far rarely used pathway is the assessment of the risk-return characteristics of Bitcoin through the lens of the options written on it, which allow new model-free insights into the existence and degree of a leverage effect, the existence and size of risk premia, resulting market inefficiencies (measurable in economic loss / money lost) and to explore the option (pun intended)/possibility of curing named inefficiencies under transaction cost and liquidity constraints. 
%
%This reasoning follows the insights of Shumway and Bakshi, who separately evaluated option returns in equity markets by means of delta-neutral option portfolios in the early 2000s. 
%
%Bakshi and Shumway both found that expected call returns exceed those of the underlying and that expected put returns underperform the underlying. Both also found that the effect increases (decreases) with the strike. 
%Bakshi found that zero-beta, at-the-money straddle positions produce average losses of approximately three percent per week and suggests that the reason may be a systematic volatility premium being priced in options. 
%
%Following Shumway, the risk-return characteristics of options can be separated in two elements, a leverage effect and a curvature effect due to the sensitivity of option payoffs to higher moments of the underlying. 
%First, the holder (buyer) of an option is subject to a leverage effect since exposure to the underlying can be obtained through the payment of a premium, which is small compared to the price of the underlying. 
%If there is a leverage effect, then (long) call options should outperform their underlying and (long) put options should underperform their underlying. Furthermore, the effect should increase (decrease) with the strike price for calls (puts).
%This reasoning allows a simple test of existence and size of the leverage effect and thus conclusions about the level of speculation associated to Bitcoin valuation. 
%
%(The second component, related to the sensitivity of options towards higher moments, implies that options are redundant assets...)
%
%If options were redundant assets, then apart of the leverage effect, delta-neutral option portfolios should earn the risk-free rate (no premium) on average.
%Until now, there is no risk-free rate available in the space of crypto assets. We establish a proxy or reference rate for a possible risk-free rate, which results from the perpetual futures written on Bitcoin and Ether. 
%
%Use Deribit and Binance perpetual futures for this, equally weighted. 
%
%(This proxy is clearly not risk-free due to counterparty risk, but we argue that a carry trade involving a short position on perpetual futures on multiple platforms in conjunction with a secured, on-chain underlying position reduces the platform-induced counterparty risk (which is the residual risk in the carry trade) as far as possible.) (This also leaves liquidity risk though!!)
%
%All underlying code is available on \href{quantlet.com}{quantlet.com}
%and a corresponding courselet is available on
%\href{quantinar.com}{quantinar.com}.
%

\section{A Proxy for the risk-free Rate of Bitcoin and Ether}

One of the most fundamental concepts in mathematical finance is the existence of a risk-free rate and a risky asset. 
While crypto assets are well-known for their riskiness, the existence of a risk-free rate has not been established in this new market. 
The risk-free rate in established markets derives from the practice of lending out money to low-risk, typically government institutions which owe an interest rate on top of the principal to the lender. Despite the naming convention, government issued bonds are not entirely risk-free due to the possibility of default.
Although default risk is commonly mitigated by a government's possibility to create money, the chance of a sovereign debt default is non-zero and hence not entirely risk-free.

In order to apply concepts of mathematical finance onto crypto assets, it is paramount to develop a risk-free rate - or at least a low-risk rate whose residual risk may converge to zero in time. 
An interesting candidate for a low-risk bond in the crypto asset space would be staking, which is the validation of transactions under the risk of losing a principal (the stake). Evidently, only crypto assets which follow a Proof-of-Stake (PoS) consensus mechanism instead of a Proof-of-Work (PoW) consensus mechanism can be used for staking. At the time, this is still a minor amount of cryptocurrencies. We include one of those, Ether, in our experiment. However, we figure that a multitude of risks is still present in staking, such as the risk of hacks, software bugs, slashing or blockchain congestions and subsequent reduction of stakes.
Another approach to develop a risk-free rate would be lending in the DeFi space. Yet, this space may be vulnerable to similar risks as staking through hacks or software bugs in smart contracts.

In the absence of riskless bond, we construct as a proxy for the risk free rate via the replication of the market-implied yield through a position in futures contracts, whose risks mainly arise due to liquidity risk and counterparty risk.
A large share of the trading volume in digital asset markets falls on the futures market. Among the futures trading, perpetual instruments have historically been and currently are the most liquid instruments. While it would be possible to use a yield from a long-short futures position (e.g. short a short-tenor future and long a longer-tenor future), we argue that in order to minimize liquidity risk, the most liquid instruments should be taken into consideration.

Solve for interest rates from Futures price $F$ Spot price $S$ at time $t$ and time-to-maturity $\tau$:

\begin{equation}
F_t = e^{r \tau} S_t
\end{equation}

Crypto exchanges often provide a so-called perpetual future, which is a synthetic (typically cash-settled instrument) whose return is tied to the underlying via a funding rate. Most exchanges (quote here) provide perpetual futures for Bitcoin and Ether. 
The price of each perpetual future is linked to the underlying by adjusting a variable interest rate, called the funding rate, that fluctuates depending on the deviation between the price of the perpetual future and the underlying. The funding rate increases the cost of holding positions against the deviation, meaning that if the price of the perpetual future is higher than the underlying, the funding rate is positive which causes the market participants who are long to pay the short side in a predefined frequency. Broadly said, the more extreme the deviation, the higher the funding rate and thus the higher the incentive to trade against the deviation and thus converge the price of the perpetual future towards its underlying. 
The risk-free rate is approximated using the simplest and most liquid way of realising the funding rate, a carry trade. In such a trade, the trader shorts (longs) the perpetual future if the funding rate is positive (negative) and simulateously buys (sells) the underlying such that the market delta of the combined position is zero.

The funding rate on \href{https://www.deribit.com/kb/deribit-linear-perpetual}{Deribit} is calculated as follows:

Instruments, such as inverse options, futures and perpetual futures, have a Mark Price $M$. The Mark Price at a time is defined as the average of the highest bid and lowest ask price within an unspecified bandwidth (quote: 
Mark price of an options contract is the current value of the option as calculated by the Deribit risk management system. Usually, this is the average of the best bid and best ask price. However, for risk management purposes, there is price bandwidth in place.
). The Deribit Index $D$, which is an equally weighted Index of BTC/USD pairs across different exchanges, is used in conjunction with the Mark Price to calculate the Premium Rate $P$ of the Perpetual Future $F$ over the Deribit Index.

\begin{equation}
P = \frac{M - D}{D}
\end{equation}

The Funding Rate results as an 8-hour interest rate from the interval in which the Premium Rate falls.

\begin{equation}
F = \max{0.025\%, P} + \min{-0.025\%, P}
\end{equation}

The actual funding payment results $f$ from the product of the Funding Rate, Position Size and Holding Time. It is continuously added to the PnL account and settled daily at 8:00 UTC.

%@Todo: Summarize!
%Mark Price
%It is essential to understand how the mark price is calculated. We start by determining the "Fair Price". The fair price is calculated as the average of the fair impact bid and the fair impact ask.
%Fair Impact Bid Price is an average price at which 100,000 USDC worth of contracts can be bought.=
%Fair Impact Ask Price is an average price at which 100,000 USDC worth of contracts can be sold.
%Fair Price can never be higher than 0.1\% of best ask or lower than best bid
%Fair Price = (Fair Impact Bid + Fair Impact Ask) / 2
%The mark price is derived using both the Deribit Index and the fair price, by adding to the Deribit Index the 30 second exponential moving average (EMA) of the Fair Price - Deribit Index.
%
%Mark Price = Deribit Index + 30 second EMA (Fair Price - Deribit Index)
%We implement a safety setting on our mark price to not deviate by more than a certain percentage of Index. This is to prevent manipulation or to curb the impact of panic trading. When a deviation from index is genuine, then we would increase the possible deviation.

The actual annualized returns of such a carry trade are summarized in the table below.

Table:

Year, Currency, Exchange, Annual Return

funding_btc_vs_eth_annualized_zoom.png
funding_btc_vs_eth_zoom.png
funding_btc_vs_eth_annualized.png
funding_btc_vs_eth.png

As of 2024-01-07, the total Open Interest for Bitcoin Perpetual Futures contracts is \$19,343,021,437 with a total 24h Trade Volume of \$50,339,886,015.
Open Interest for Ether Perpetual Futures contracts is Total Open Interest \$14,945,055,482 and total 24h Trade Volume is \$18,097,004,167.

BTC Futures

Total Open Interest $1,405,918,350
Total 24h Trade Volume $324,820,482

ETH Futures
Total Open Interest $466,932,584
Total 24h Trade Volume $98,238,298

\begin{table}[]
\begin{tabular}{lll}
\textbf{Instrument}            & \textbf{Aggregate Open Interest (USD)} & \textbf{24 Hour Aggregate Trade Volume (USD)} \\
BTC Perpetual Futures & 19,343,021,437                & 50,339,886,015                       \\
ETH Perpetual Futures & 14,945,055,482                & 18,097,004,167                       \\
BTC Futures           & 1,405,918,350                 & 324,820,482                          \\
ETH Futures           & 466,932,584                   & 98,238,298                          
\end{tabular}
\end{table}

Source for liquidity: 
https://www.coingecko.com/en/derivatives
https://www.coingecko.com/en/derivatives/futures

% Source: https://www.tablesgenerator.com/#
% Please add the following required packages to your document preamble:
% \usepackage[table,xcdraw]{xcolor}
% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage[table,xcdraw]{xcolor}
% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}

As of 2024-01-07

Bitcoin Perpetuals with 
Total Open Interest \$19,343,021,437
Total 24h Trade Volume \$50,339,886,015

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{llrrrrrrr}
\hline
\textbf{Exchange}    & \textbf{Symbol} & \multicolumn{1}{l}{\textbf{Price}} & \multicolumn{1}{l}{\textbf{Index Price}} & \multicolumn{1}{l}{\textbf{Basis}} & \multicolumn{1}{l}{\textbf{Spread}} & \multicolumn{1}{l}{\textbf{Funding Rate}} & \multicolumn{1}{l}{\textbf{Open Interest}} & \multicolumn{1}{l}{\textbf{24h Volume}} \\
\hline
\hline
\# Binance (Futures) & BTCUSDT         & \$43,995.06                        & 44011.67 USDT                            & +0.087\%                           & 0.01\%                              & 0.010\%                                   & \$3,214,727,375                            & \$6,822,091,229.25                      \\
\# Binance (Futures) & BTCUSD\_PERP    & \$44,008.70                        & 44037.64 USD                             & +0.066\%                           & 0.01\%                              & 0.010\%                                   & \$808,036,100                              & \$973,585,131.72                        \\
\# BingX (Futures)   & BTC-USDT        & \$44,042.40                        & 44046.6 USDT                             & +0.044\%                           & 0.01\%                              & 0.010\%                                   & \$909,897,901                              & \$1,867,895,422.35                      \\
\# Bitget Futures    & BTCUSDT\_UMCBL  & \$44,049.08                        & 44033.85 USDT                            & +0.015\%                           & 0.01\%                              & 0.011\%                                   & \$1,790,553,246                            & \$5,701,072,691.79                      \\
\# Bitmart Futures   & BTCUSDT         & \$44,015.47                        & 44003.79 USDT                            & +0.024\%                           & 0.01\%                              & 0.010\%                                   & \$641,006,203                              & \$2,917,088,729.74                      \\
\# Bybit (Futures)   & BTCUSDT         & \$44,039.88                        & 44016.82 USDT                            & -0.003\%                           & 0.01\%                              & 0.010\%                                   & \$2,538,053,148                            & \$2,672,726,926.85                      \\
\# Bybit (Futures)   & BTCUSD          & \$43,995.50                        & 44037.2 USD                              & +0.095\%                           & 0.01\%                              & 0.010\%                                   & \$652,512,486                              & \$283,394,472.00                        \\
\# Deribit           & BTC-PERPETUAL   & \$44,020.00                        & 44028.17 USD                             & +0.019\%                           & 0.01\%                              & 0.004\%                                   & \$577,506,570                              & \$204,526,951.86                        \\
\# OKX (Futures)     & BTC-USDT-SWAP   & \$44,014.07                        & 43989.3 USDT                             & -0.006\%                           & 0.01\%                              & -0.009\%                                  & \$984,130,801                              & \$2,648,725,568.71                      \\
\# Prime XBT         & BTC/USD         & \$43,981.30                        & -                                        & -                                  & 0.02\%                              & -                                         & \$920,322,200                              & \$60,224,491.31                        
\end{tabular}}
\end{table}

Ether Perpetuals

Total Open Interest \$14,945,055,482
Total 24h Trade Volume \$18,097,004,167

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage[table,xcdraw]{xcolor}
% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage[table,xcdraw]{xcolor}
% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}

\begin{table}[ht]
\resizebox{\textwidth}{!}{\begin{tabular}{llrrrrrrr}
\hline
\textbf{Exchange}         & \textbf{Symbol} & \multicolumn{1}{l}{\textbf{Price}} & \multicolumn{1}{l}{\textbf{Index Price}} & \multicolumn{1}{l}{\textbf{Basis}} & \multicolumn{1}{l}{\textbf{Spread}} & \multicolumn{1}{l}{\textbf{Funding Rate}} & \multicolumn{1}{l}{\textbf{Open Interest}} & \multicolumn{1}{l}{\textbf{24h Volume}} \\
\hline
\hline
\# Deepcoin (Derivatives) & ETHUSDT         & \$2,232.18                         & 2232.46 USDT                             & +0.047\%                           & 0.02\%                              & 0.006\%                                   & \$6,376,015,545                            & \$237,202,788.68                        \\
\# Binance (Futures)      & ETHUSDT         & \$2,227.11                         & 2227.74 USDT                             & +0.078\%                           & 0.01\%                              & 0.010\%                                   & \$1,680,165,204                            & \$2,843,630,044.97                      \\
\# Bybit (Futures)        & ETHUSDT         & \$2,228.15                         & 2228.15 USDT                             & +0.049\%                           & 0.01\%                              & 0.010\%                                   & \$1,027,647,498                            & \$540,402,664.24                        \\
\# OKX (Futures)          & ETH-USDT-SWAP   & \$2,227.96                         & 2227.07 USDT                             & +0.010\%                           & 0.01\%                              & -0.003\%                                  & \$584,652,223                              & \$1,748,266,659.63                      \\
\# Gate.io (Futures)      & ETH\_USDT       & \$2,228.62                         & 2227.03 USDT                             & -0.021\%                           & 0.01\%                              & 0.010\%                                   & \$508,608,257                              & \$189,744,908.47                        \\
\# BingX (Futures)        & ETH-USDT        & \$2,229.64                         & 2230.26 USDT                             & +0.062\%                           & 0.02\%                              & 0.010\%                                   & \$475,301,553                              & \$609,765,117.59                        \\
\# Prime XBT              & ETH/USD         & \$2,227.49                         & -                                        & -                                  & 0.04\%                              & -                                         & \$473,765,000                              & \$16,309,868.89                         \\
\# Deribit                & ETH-PERPETUAL   & \$2,228.05                         & 2228.59 USD                              & +0.024\%                           & 0.01\%                              & 0.004\%                                   & \$447,911,369                              & \$82,175,344.52                         \\
\# Binance (Futures)      & ETHUSD\_PERP    & \$2,227.69                         & 2228.74 USD                              & +0.047\%                           & 0.01\%                              & 0.010\%                                   & \$366,493,140                              & \$257,267,395.24                        \\
\# XT.COM (Derivatives)   & ETH\_USDT       & \$2,228.19                         & 2225.55 USDT                             & +0.031\%                           & 0.01\%                              & -0.010\%                                  & \$325,480,371                              & \$1,137,587,629.30                     
\end{tabular}}
\end{table}

BTC Futures

with

Total Open Interest $1,405,918,350
Total 24h Trade Volume $324,820,482

\begin{table}[ht]
\resizebox{\textwidth}{!}{\begin{tabular}{llrrrrrrr}
\hline
\textbf{Exchange}                & \textbf{Symbol}    & \multicolumn{1}{l}{\textbf{Price}} & \multicolumn{1}{l}{\textbf{Index Price}} & \multicolumn{1}{l}{\textbf{Basis}} & \multicolumn{1}{l}{\textbf{Spread}} & \multicolumn{1}{l}{\textbf{Expires in}} & \multicolumn{1}{l}{\textbf{Open Interest}} & \multicolumn{1}{l}{\textbf{24h Volume}} \\
\hline
\hline
\# Binance (Futures)             & BTCUSD\_240329     & \$45,372.20                        & 44041.17 USD                             & -2.934\%                           & 0.01\%                              & 3 months                                & \$136,765,200                              & \$19,534,741.00                         \\
\# Binance (Futures)             & BTCUSD\_240628     & \$46,802.20                        & 44037.64 USD                             & -5.907\%                           & 0.02\%                              & 6 months                                & \$60,673,400                               & \$11,015,730.09                         \\
\# Binance (Futures)             & BTCUSDT\_240329    & \$45,340.32                        & 44011.62 USDT                            & -2.883\%                           & 0.04\%                              & -                                       & \$38,608,233                               & \$13,341,161.63                         \\
\# Binance (Futures)             & BTCUSDT\_240628    & \$46,588.03                        & 44011.62 USDT                            & -5.484\%                           & 0.02\%                              & -                                       & \$7,296,676                                & \$6,727,917.37                          \\
\# BitMEX (Derivative)           & XBTH24             & \$45,170.00                        & 44016.02 USD                             & -2.555\%                           & 0.03\%                              & 3 months                                & \$23,123,300                               & \$2,586,155.50                          \\
\# BitMEX (Derivative)           & XBTM24             & \$46,319.00                        & 44016.02 USD                             & -4.972\%                           & 0.15\%                              & 6 months                                & \$14,701,700                               & \$899,753.13                            \\
\# BitMEX (Derivative)           & XBTF24             & \$44,599.50                        & 44016.02 USD                             & -1.308\%                           & 0.18\%                              & 19 days                                 & \$3,686,400                                & \$42,073.68                             \\
\# BitMEX (Derivative)           & XBTU24             & \$47,680.00                        & 44016.02 USD                             & -7.685\%                           & 0.82\%                              & 9 months                                & \$874,700                                  & \$14,332.62                             \\
\# BitMEX (Derivative)           & XBTG24             & \$45,034.50                        & 44016.02 USD                             & -2.262\%                           & 0.2\%                               & about 2 months                          & \$149,400                                  & \$502.57                                \\
\# BitMEX (Derivative)           & XBTUSDTH24         & \$44,913.12                        & 43991.27 USDT                            & -2.003\%                           & 22.08\%                             & 3 months                                & \$67,027                                   & \$23,309.91                             \\
\# BTSE (Futures)                & BTC-240329         & \$45,346.20                        & 45337.26 USD                             & -0.020\%                           & 0.15\%                              & 3 months                                & \$70,779,838                               & \$70,829.75                             \\
\# BTSE (Futures)                & BTC-240628         & \$46,578.30                        & 46774.54 USD                             & +0.421\%                           & 0.2\%                               & 6 months                                & \$68,641,915                               & \$58,755.91                             \\
\# Bybit (Futures)               & BTCUSDH24          & \$45,515.00                        & 44050.12 USD                             & -3.218\%                           & 0.01\%                              & 3 months                                & \$31,042,370                               & \$8,992,250.00                          \\
\# Bybit (Futures)               & BTCUSDM24          & \$46,867.50                        & 44046.87 USD                             & -6.018\%                           & 0.01\%                              & 6 months                                & \$13,200,403                               & \$932,634.00                            \\
\# Bybit (Futures)               & BTC-29MAR24        & \$45,288.92                        & 44042.99 USDC                            & -2.786\%                           & 0.07\%                              & 3 months                                & -                                          & \$1,650,409.10                          \\
\# Bybit (Futures)               & BTC-28JUN24        & \$46,655.43                        & 44042.83 USDC                            & -5.633\%                           & 0.07\%                              & 6 months                                & -                                          & \$636,256.66                            \\
\# Bybit (Futures)               & BTC-12JAN24        & \$44,205.30                        & 44042.83 USDC                            & -0.403\%                           & 0.02\%                              & 5 days                                  & -                                          & \$320,758.50                            \\
\# Bybit (Futures)               & BTC-19JAN24        & \$44,341.25                        & 44042.83 USDC                            & -0.708\%                           & 0.03\%                              & 12 days                                 & -                                          & \$117,823.82                            \\
\# Bybit (Futures)               & BTC-27SEP24        & \$47,694.06                        & 44042.99 USDC                            & -7.688\%                           & 0.08\%                              & 9 months                                & -                                          & \$66,406.59                             \\
\# Bybit (Futures)               & BTC-26JAN24        & \$44,370.24                        & 44042.75 USDC                            & -0.773\%                           & 0.02\%                              & 19 days                                 & -                                          & \$47,245.15                             \\
\# Bybit (Futures)               & BTC-23FEB24        & \$44,806.59                        & 44042.99 USDC                            & -1.739\%                           & 0.03\%                              & about 2 months                          & -                                          & \$8,650.36                              \\
\# Crypto.com Exchange (Futures) & BTCUSD-240628      & \$46,384.20                        & 44003.38 USD                             & -5.133\%                           & 0.41\%                              & -                                       & \$24,557,224                               & \$15,659.31                             \\
\# Crypto.com Exchange (Futures) & BTCUSD-240329      & \$45,024.60                        & 44003.38 USD                             & -2.268\%                           & 0.46\%                              & -                                       & \$1,118,289                                & \$17,244.42                             \\
\# Crypto.com Exchange (Futures) & BTCUSD-240126      & \$44,215.10                        & 44003.38 USD                             & -0.479\%                           & 0.29\%                              & -                                       & \$9,701                                    & \$18,428.85                             \\
\# Crypto.com Exchange (Futures) & BTCUSD-240223      & \$44,624.00                        & 44003.38 USD                             & -1.391\%                           & 0.43\%                              & -                                       & \$3,539                                    & \$14,931.19                             \\
\# Deribit                       & BTC-29MAR24        & \$45,300.00                        & 44026.9 USD                              & -2.810\%                           & 0.02\%                              & 3 months                                & \$208,084,190                              & \$15,282,281.56                         \\
\# Deribit                       & BTC-26JAN24        & \$44,475.00                        & 44026.9 USD                              & -1.008\%                           & 0.1\%                               & 19 days                                 & \$89,621,140                               & \$3,132,087.87                          \\
\# Deribit                       & BTC-28JUN24        & \$46,577.50                        & 44026.9 USD                              & -5.476\%                           & 0.02\%                              & 6 months                                & \$79,314,230                               & \$7,916,251.72                          \\
\# Deribit                       & BTC-27SEP24        & \$47,715.00                        & 44026.9 USD                              & -7.729\%                           & 0.09\%                              & 9 months                                & \$32,099,570                               & \$1,034,939.35                          \\
\# Deribit                       & BTC-12JAN24        & \$44,110.00                        & 44026.9 USD                              & -0.188\%                           & 0.03\%                              & 5 days                                  & \$25,884,930                               & \$7,230,738.10                          \\
\# Deribit                       & BTC-23FEB24        & \$44,767.50                        & 44026.9 USD                              & -1.654\%                           & 0.03\%                              & about 2 months                          & \$22,571,940                               & \$744,850.73                            \\
\# Deribit                       & BTC-27DEC24        & \$49,247.50                        & 44026.9 USD                              & -10.601\%                          & 0.13\%                              & 12 months                               & \$11,398,600                               & \$600,503.46                            \\
\# Deribit                       & BTC-19JAN24        & \$44,222.50                        & 44026.9 USD                              & -0.442\%                           & 0.04\%                              & 12 days                                 & \$5,538,440                                & \$2,799,269.71                          \\
\# HTX Futures                   & BTC\_CQ            & \$45,297.68                        & 44023.11 USD                             & -2.814\%                           & 0.38\%                              & 3 months                                & \$1,002,200                                & \$1,602,716.03                          \\
\# HTX Futures                   & BTC\_CW            & \$44,095.65                        & 44030.71 USD                             & -0.147\%                           & 0.45\%                              & 4 days                                  & \$47,100                                   & \$1,011,419.90                          \\
\# HTX Futures                   & BTC\_NW            & \$44,211.45                        & 44023.11 USD                             & -0.426\%                           & 0.4\%                               & 11 days                                 & \$9,000                                    & \$967,082.71                            \\
\# Kraken (Futures)              & FI\_XBTUSD\_240329 & \$45,520.50                        & 44013.38 USD                             & -3.311\%                           & 0.08\%                              & 3 months                                & \$6,608,999                                & \$112,442.00                            \\
\# Kraken (Futures)              & FI\_XBTUSD\_240628 & \$46,571.50                        & 44013.38 USD                             & -5.493\%                           & 0.24\%                              & 6 months                                & \$1,929,982                                & \$275,529.00                            \\
\# Kraken (Futures)              & FI\_XBTUSD\_240126 & \$44,191.50                        & 44013.38 USD                             & -0.403\%                           & 0.4\%                               & 19 days                                 & \$300,856                                  & \$3.00                                  \\
\# Kraken (Futures)              & FF\_XBTUSD\_240329 & \$45,284.00                        & 44013.38 USD                             & -2.806\%                           & 0.11\%                              & 3 months                                & \$87                                       & \$22.68                                 \\
\# Kraken (Futures)              & FF\_XBTUSD\_240126 & \$44,307.00                        & 44013.38 USD                             & -0.663\%                           & 0.08\%                              & 19 days                                 & \$32                                       & \$4.61                                  \\
\# KuCoin Futures                & XBTMH24            & \$45,444.00                        & 44072.8 USD                              & -3.017\%                           & 0.19\%                              & 3 months                                & \$11,251,580                               & \$86,193.00                             \\
\# OKX (Futures)                 & BTC-USD-240329     & \$45,322.90                        & 44016.1 USD                              & -2.883\%                           & 0.01\%                              & 3 months                                & \$175,555,100                              & \$42,597,534.31                         \\
\# OKX (Futures)                 & BTC-USD-240628     & \$46,744.90                        & 44016.1 USD                              & -5.838\%                           & 0.03\%                              & 6 months                                & \$67,199,400                               & \$13,281,039.45                         \\
\# OKX (Futures)                 & BTC-USDT-240329    & \$45,345.34                        & 43989.3 USDT                             & -2.942\%                           & 0.02\%                              & 3 months                                & \$54,809,841                               & \$32,494,014.76                         \\
\# OKX (Futures)                 & BTC-USDT-240628    & \$46,824.58                        & 43989.3 USDT                             & -6.008\%                           & 0.02\%                              & 6 months                                & \$49,096,426                               & \$8,216,309.33                          \\
\# OKX (Futures)                 & BTC-USD-240126     & \$44,235.20                        & 44016.1 USD                              & -0.495\%                           & 0.04\%                              & 19 days                                 & \$20,945,300                               & \$1,424,116.88                          \\
\# OKX (Futures)                 & BTC-USD-240112     & \$44,085.20                        & 44016.1 USD                              & -0.157\%                           & 0.01\%                              & 5 days                                  & \$15,394,800                               & \$11,575,786.01                         \\
\# OKX (Futures)                 & BTC-USDT-240126    & \$44,284.50                        & 43989.3 USDT                             & -0.617\%                           & 0.06\%                              & 19 days                                 & \$8,573,427                                & \$1,552,614.65                          \\
\# OKX (Futures)                 & BTC-USD-240223     & \$44,719.10                        & 44016.1 USD                              & -1.572\%                           & 0.04\%                              & about 2 months                          & \$8,046,700                                & \$3,603,630.54                          \\
\# OKX (Futures)                 & BTC-USD-240119     & \$44,198.90                        & 44016.1 USD                              & -0.414\%                           & 0.01\%                              & 12 days                                 & \$4,957,200                                & \$6,247,231.64                          \\
\# OKX (Futures)                 & BTC-USDT-240112    & \$44,080.50                        & 43989.3 USDT                             & -0.157\%                           & 0.03\%                              & 5 days                                  & \$3,560,036                                & \$12,798,332.26                         \\
\# OKX (Futures)                 & BTC-USDT-240223    & \$44,781.65                        & 43989.3 USDT                             & -1.720\%                           & 0.08\%                              & about 2 months                          & \$2,960,896                                & \$711,132.65                            \\
\# OKX (Futures)                 & BTC-USDT-240119    & \$44,192.36                        & 43989.3 USDT                             & -0.409\%                           & 0.04\%                              & 12 days                                 & \$975,364                                  & \$4,235,395.39                          \\
\# OKX (Futures)                 & BTC-USDC-240628    & \$46,788.02                        & 44015.4 USDC                             & -5.958\%                           & 0.39\%                              & 6 months                                & -                                          & \$17,320.92                             \\
\# OKX (Futures)                 & BTC-USDC-240119    & \$44,137.13                        & 44020.7 USDC                             & -0.298\%                           & 0.1\%                               & 12 days                                 & -                                          & \$6,285.13                              \\
\# OKX (Futures)                 & BTC-USDC-240329    & \$45,832.15                        & 44020.7 USDC                             & -3.985\%                           & 0.31\%                              & 3 months                                & -                                          & \$1,874.53                              \\
\# OKX (Futures)                 & BTC-USDC-240112    & \$44,107.04                        & 44015.4 USDC                             & -0.242\%                           & 0.13\%                              & 5 days                                  & -                                          & \$9,765.30                              \\
\# XT.COM (Derivatives)          & BTC\_USD\_240329   & \$45,380.70                        & -                                        & -3.586\%                           & 0.01\%                              & 3 months                                & -                                          & \$33,113,000.00                         \\
\# XT.COM (Derivatives)          & BTC\_USD\_240628   & \$46,797.20                        & -                                        & -                                  & 0.01\%                              & 6 months                                & -                                          & \$31,000,200.00                         \\
\# XT.COM (Derivatives)          & BTC\_USDT\_240329  & \$45,369.50                        & -                                        & -                                  & 0.06\%                              & 3 months                                & -                                          & \$20,870,477.00                        
\end{tabular}}
\end{table}

ETH Futures

with 
Total Open Interest $466,932,584
Total 24h Trade Volume $98,238,298

\begin{table}[ht]
\resizebox{\textwidth}{!}{\begin{tabular}{llrrrrrrr}
\hline
\textbf{Exchange}                & \textbf{Symbol}    & \multicolumn{1}{l}{\textbf{Price}} & \multicolumn{1}{l}{\textbf{Index Price}} & \multicolumn{1}{l}{\textbf{Basis}} & \multicolumn{1}{l}{\textbf{Spread}} & \multicolumn{1}{l}{\textbf{Expires in}} & \multicolumn{1}{l}{\textbf{Open Interest}} & \multicolumn{1}{l}{\textbf{24h Volume}} \\
\hline
\hline
\# Binance (Futures)             & ETHUSD\_240329     & \$2,300.60                         & 2228.81 USD                              & -3.121\%                           & 0.01\%                              & 3 months                                & \$72,215,240                               & \$16,326,927.83                         \\
\# Binance (Futures)             & ETHUSD\_240628     & \$2,368.79                         & 2229.03 USD                              & -5.900\%                           & 0.02\%                              & 6 months                                & \$11,183,360                               & \$4,224,925.26                          \\
\# Binance (Futures)             & ETHUSDT\_240329    & \$2,300.46                         & 2227.81 USDT                             & -3.110\%                           & 0.02\%                              & -                                       & \$35,040                                   & \$6,902,127.72                          \\
\# Binance (Futures)             & ETHUSDT\_240628    & \$2,360.43                         & 2227.81 USDT                             & -5.572\%                           & 0.03\%                              & -                                       & \$4,604                                    & \$4,303,853.27                          \\
\# BitMEX (Derivative)           & ETHUSDH24          & \$2,452.60                         & 2228.01 USD                              & -9.157\%                           & 0.02\%                              & 3 months                                & \$4,341,286                                & \$537,210.05                            \\
\# BitMEX (Derivative)           & ETHH24             & \$2,217.23                         & 0.05 XBT                                 & +0.457\%                           & 0.37\%                              & 3 months                                & \$1,575,478                                & \$69,244.08                             \\
\# BTSE (Futures)                & ETH-240628         & \$2,350.05                         & 2367.26 USD                              & +0.732\%                           & 0.75\%                              & 6 months                                & \$2,517,764                                & \$49,845.75                             \\
\# BTSE (Futures)                & ETH-240329         & \$2,297.48                         & 2298.78 USD                              & +0.057\%                           & 0.23\%                              & 3 months                                & \$2,081,180                                & \$49,130.96                             \\
\# Bybit (Futures)               & ETHUSDM24          & \$2,357.45                         & 2229.63 USD                              & -5.422\%                           & 0.05\%                              & 6 months                                & -                                          & \$448,716.00                            \\
\# Bybit (Futures)               & ETHUSDH24          & \$2,287.90                         & 2229.53 USD                              & -2.551\%                           & 0.06\%                              & 3 months                                & -                                          & \$294,007.00                            \\
\# Bybit (Futures)               & ETH-26JAN24        & \$2,243.65                         & 2229.46 USDC                             & -0.668\%                           & 0.05\%                              & 19 days                                 & -                                          & \$286,245.77                            \\
\# Bybit (Futures)               & ETH-12JAN24        & \$2,231.66                         & 2229.51 USDC                             & -0.132\%                           & 0.03\%                              & 5 days                                  & -                                          & \$99,683.25                             \\
\# Bybit (Futures)               & ETH-28JUN24        & \$2,368.66                         & 2229.46 USDC                             & -5.910\%                           & 0.1\%                               & 6 months                                & -                                          & \$97,357.35                             \\
\# Bybit (Futures)               & ETH-19JAN24        & \$2,238.21                         & 2229.51 USDC                             & -0.424\%                           & 0.06\%                              & 12 days                                 & -                                          & \$92,957.40                             \\
\# Bybit (Futures)               & ETH-23FEB24        & \$2,274.64                         & 2229.51 USDC                             & -2.019\%                           & 0.03\%                              & about 2 months                          & -                                          & \$68,610.65                             \\
\# Bybit (Futures)               & ETH-29MAR24        & \$2,297.48                         & 2229.51 USDC                             & -2.993\%                           & 0.03\%                              & 3 months                                & -                                          & \$36,926.81                             \\
\# Bybit (Futures)               & ETH-27SEP24        & \$2,420.24                         & 2229.46 USDC                             & -7.915\%                           & 0.12\%                              & 9 months                                & -                                          & \$30,008.99                             \\
\# Crypto.com Exchange (Futures) & ETHUSD-240628      & \$2,347.68                         & 2224.96 USD                              & -5.227\%                           & 0.24\%                              & -                                       & \$24,573,170                               & \$822.63                                \\
\# Crypto.com Exchange (Futures) & ETHUSD-240126      & \$2,234.13                         & 2224.96 USD                              & -0.410\%                           & 0.14\%                              & -                                       & \$79,732                                   & \$742.18                                \\
\# Crypto.com Exchange (Futures) & ETHUSD-240329      & \$2,286.18                         & 2224.96 USD                              & -2.678\%                           & 0.29\%                              & -                                       & \$58,845                                   & \$895.27                                \\
\# Crypto.com Exchange (Futures) & ETHUSD-240223      & \$2,249.95                         & 2224.92 USD                              & -1.112\%                           & 0.17\%                              & -                                       & \$27,557                                   & \$1,247.37                              \\
\# Deribit                       & ETH-29MAR24        & \$2,296.00                         & 2228.59 USD                              & -2.936\%                           & 0.02\%                              & 3 months                                & \$88,364,731                               & \$2,438,537.07                          \\
\# Deribit                       & ETH-28JUN24        & \$2,363.75                         & 2228.59 USD                              & -5.718\%                           & 0.01\%                              & 6 months                                & \$35,776,949                               & \$1,152,627.40                          \\
\# Deribit                       & ETH-26JAN24        & \$2,247.75                         & 2228.59 USD                              & -0.852\%                           & 0.02\%                              & 19 days                                 & \$34,172,788                               & \$2,694,255.34                          \\
\# Deribit                       & ETH-27SEP24        & \$2,414.50                         & 2228.59 USD                              & -7.700\%                           & 0.02\%                              & 9 months                                & \$22,332,082                               & \$781,745.26                            \\
\# Deribit                       & ETH-12JAN24        & \$2,232.25                         & 2228.59 USD                              & -0.164\%                           & 0.01\%                              & 5 days                                  & \$10,281,174                               & \$1,626,945.44                          \\
\# Deribit                       & ETH-23FEB24        & \$2,273.25                         & 2228.59 USD                              & -1.965\%                           & 0.08\%                              & about 2 months                          & \$10,187,446                               & \$484,690.68                            \\
\# Deribit                       & ETH-27DEC24        & \$2,463.00                         & 2228.59 USD                              & -9.517\%                           & 0.21\%                              & 12 months                               & \$6,513,788                                & \$630,519.38                            \\
\# Deribit                       & ETH-19JAN24        & \$2,235.50                         & 2228.59 USD                              & -0.309\%                           & 0.03\%                              & 12 days                                 & \$1,564,812                                & \$1,319,273.13                          \\
\# HTX Futures                   & ETH\_CQ            & \$2,298.00                         & 2228.66 USD                              & -3.018\%                           & 0.01\%                              & 3 months                                & \$1,130,370                                & \$562,732.95                            \\
\# HTX Futures                   & ETH\_CW            & \$2,233.11                         & 2228.54 USD                              & -0.205\%                           & 0.01\%                              & 4 days                                  & \$49,870                                   & \$481,669.51                            \\
\# HTX Futures                   & ETH\_NW            & \$2,242.45                         & 2228.66 USD                              & -0.615\%                           & 0.35\%                              & 11 days                                 & \$31,200                                   & \$302,572.28                            \\
\# Kraken (Futures)              & FI\_ETHUSD\_240329 & \$2,300.00                         & 2228 USD                                 & -3.130\%                           & 0.11\%                              & 3 months                                & \$4,415,821                                & \$29,062.00                             \\
\# Kraken (Futures)              & FI\_ETHUSD\_240628 & \$2,366.30                         & 2228 USD                                 & -5.845\%                           & 0.25\%                              & 6 months                                & \$3,541,749                                & \$40,752.00                             \\
\# Kraken (Futures)              & FF\_ETHUSD\_240329 & \$2,314.00                         & 2228 USD                                 & -3.717\%                           & 0.13\%                              & 3 months                                & \$1,584                                    & \$0.45                                  \\
\# Kraken (Futures)              & FF\_ETHUSD\_240126 & \$2,254.90                         & 2228 USD                                 & -1.193\%                           & 0.09\%                              & 19 days                                 & \$299                                      & \$36.69                                 \\
\# Kraken (Futures)              & FI\_ETHUSD\_240126 & \$2,254.15                         & 2228 USD                                 & -1.160\%                           & 0.24\%                              & 19 days                                 & \$48,594                                   & \$10,184.00                             \\
\# OKX (Futures)                 & ETH-USD-240329     & \$2,296.78                         & 2228.71 USD                              & -2.964\%                           & 0.01\%                              & 3 months                                & \$61,638,000                               & \$13,362,284.32                         \\
\# OKX (Futures)                 & ETH-USD-240628     & \$2,370.47                         & 2228.67 USD                              & -5.982\%                           & 0.03\%                              & 6 months                                & \$29,831,250                               & \$8,793,848.00                          \\
\# OKX (Futures)                 & ETH-USDT-240329    & \$2,296.43                         & 2226.75 USDT                             & -2.985\%                           & 0.02\%                              & 3 months                                & \$20,412,503                               & \$5,701,567.75                          \\
\# OKX (Futures)                 & ETH-USD-240112     & \$2,231.34                         & 2228.71 USD                              & -0.118\%                           & 0.01\%                              & 5 days                                  & \$7,456,500                                & \$8,340,344.15                          \\
\# OKX (Futures)                 & ETH-USDT-240628    & \$2,366.59                         & 2227.15 USDT                             & -5.845\%                           & 0.02\%                              & 6 months                                & \$3,885,976                                & \$3,157,980.17                          \\
\# OKX (Futures)                 & ETH-USD-240119     & \$2,233.42                         & 2228.71 USD                              & -0.211\%                           & 0.03\%                              & 12 days                                 & \$3,212,320                                & \$3,467,988.47                          \\
\# OKX (Futures)                 & ETH-USDT-240112    & \$2,229.86                         & 2226.75 USDT                             & -0.089\%                           & 0.01\%                              & 5 days                                  & \$2,298,019                                & \$3,571,348.57                          \\
\# OKX (Futures)                 & ETH-USDT-240119    & \$2,235.50                         & 2226.75 USDT                             & -0.341\%                           & 0.01\%                              & 12 days                                 & \$690,474                                  & \$2,364,707.49                          \\
\# OKX (Futures)                 & ETH-USDC-240329    & \$2,296.52                         & 2227.93 USDC                             & -3.020\%                           & 0.38\%                              & 3 months                                & -                                          & \$55,373.71                             \\
\# OKX (Futures)                 & ETH-USDC-240112    & \$2,251.63                         & 2227.89 USDC                             & -1.088\%                           & 0.09\%                              & 5 days                                  & -                                          & \$5,770.92                              \\
\# OKX (Futures)                 & ETH-USDC-240119    & \$2,251.87                         & 2227.93 USDC                             & -1.037\%                           & 0.14\%                              & 12 days                                 & -                                          & \$9.01                                  \\
\# XT.COM (Derivatives)          & ETH\_USD\_240628   & \$2,368.07                         & -                                        & -                                  & 0.04\%                              & 6 months                                & -                                          & \$2,929,170.00                          \\
\# XT.COM (Derivatives)          & ETH\_USD\_240329   & \$2,299.63                         & -                                        & -2.350\%                           & 0.01\%                              & 3 months                                & -                                          & \$2,698,030.00                          \\
\# XT.COM (Derivatives)          & ETH\_USDT\_240329  & \$2,301.93                         & -                                        & -                                  & 0.03\%                              & 3 months                                & -                                          & \$124,645.12                            \\
\# OKX (Futures)                 & BTC-USDT-240112    & \$44,080.50                        & -0.157\%                                 & 0.03\%                             & 5 days                              & \$3,560,036                             & \$12,798,332.26                            & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDT-240223    & \$44,781.65                        & -1.720\%                                 & 0.08\%                             & about 2 months                      & \$2,960,896                             & \$711,132.65                               & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDT-240119    & \$44,192.36                        & -0.409\%                                 & 0.04\%                             & 12 days                             & \$975,364                               & \$4,235,395.39                             & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDC-240628    & \$46,788.02                        & -5.958\%                                 & 0.39\%                             & 6 months                            & -                                       & \$17,320.92                                & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDC-240119    & \$44,137.13                        & -0.298\%                                 & 0.1\%                              & 12 days                             & -                                       & \$6,285.13                                 & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDC-240329    & \$45,832.15                        & -3.985\%                                 & 0.31\%                             & 3 months                            & -                                       & \$1,874.53                                 & \multicolumn{1}{l}{}                    \\
\# OKX (Futures)                 & BTC-USDC-240112    & \$44,107.04                        & -0.242\%                                 & 0.13\%                             & 5 days                              & -                                       & \$9,765.30                                 & \multicolumn{1}{l}{}                    \\
\# XT.COM (Derivatives)          & BTC\_USD\_240329   & \$45,380.70                        & -3.586\%                                 & 0.01\%                             & 3 months                            & -                                       & \$33,113,000.00                            & \multicolumn{1}{l}{}                    \\
\# XT.COM (Derivatives)          & BTC\_USD\_240628   & \$46,797.20                        & -                                        & 0.01\%                             & 6 months                            & -                                       & \$31,000,200.00                            & \multicolumn{1}{l}{}                    \\
\# XT.COM (Derivatives)          & BTC\_USDT\_240329  & \$45,369.50                        & -                                        & 0.06\%                             & 3 months                            & -                                       & \$20,870,477.00                            & \multicolumn{1}{l}{}                   
\end{tabular}}
\end{table}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}

\section{Methodology}

We follow the argument that any beta-neutral portfolio should yield the risk-free rate. We construct zero-beta portfolios by pairing a call option and a put option to a straddle and adjusting the weights so that the sum of option betas yields zero. The straddle's payoff $r_v$ is depicted in Equation 1, consisting of the payoff of the call $r_c$ and the payoff of the put $r_p$. The weight $\theta$ is found by solving Equation 2 through Equation 3. The (market) beta of a call (put) can be calculated through knowledge of the spot price $s$, the option delta $\delta$, the Black-Scholes option price $C$ and the market beta $\beta_m (=1)$ as depicted in Equation 4. 
The underlying data are executed transactions. All option greeks remain constant except time-to-maturity. Any transaction has a time-to-maturity (tau) "reset"/synced to settlement time. All options are bought and held until settlement at 10am UTC. 
We match our universe of straddles as follows:
Per definition, a straddle exists of a call and a put with similar strike and maturity. We construct the set of empirical straddles by iterating over our list of executed calls and puts. For each call (put), we use Put-Call-Parity to complete the straddle. 
The empirical performance of the resulting zero-beta straddles is assessed on a daily base. We inspect the densities (via Kernel Density Estimation) of the daily returns and argue that they are t-distributed. For each level of moneyness m, we construct an interval $[m, m + 0.05]$. We condition daily returns on over levels of the interval and we conduct a left-sided t-test on the null hypothesis of a population mean greater or equal to zero. We define a structural (or significant?) non-zero return of the zero-beta straddle returns as (volatility risk) premium.

Straddle Return $r_v$￼, Share of Straddle’s value in Call Options ￼$\theta$, Call and Put market Beta ￼$\beta_c$, $\beta_p$.

\begin{equation}
r_v = \theta r_c + (1-\theta)r_p
\end{equation}

\begin{equation}
\theta \beta_c + (1-\theta)\beta_p = 0\\
\end{equation}

\begin{equation}
\theta = \frac{-\beta_p}{\beta_c - \beta_p}\\
\end{equation}

\begin{equation}
\beta_c = \frac{s}{C} \Delta \beta_s\\
\end{equation}

We consider the possibility that if a volatility premium is found in the vanilla Zero-Beta-Straddles, then it may attributed to the risk of Black Swan events. Therefore we repeat the analysis for a set of "crash-resistant" zero-beta straddles. 
Hedge Straddle with FOTM Option in order to exclude influence of Fat Tail Events
Close Call (Put) leg 20\% away from Strike. We suppose that a such a leg would come at a 30\% premium compared to the ATM volatility (which is more than empirically justified.)
Maybe make a summary of the relationship between ATM and 20\% OTM premium.

The performance of the vanilla and crash-resistant straddles is evaluated with and without fees. Clearly, application of the fees constitutes simply a linear shift, which is well-visible in the graphs. 
The return of the straddles until settlement is re-scaled on a daily level. It is plotted as a boxplot of daily returns for each level of moneyness in the interval [0.7, 1.3]. 
Fees are standard option fees on Deribit. We assume using an option combo fee, which makes the a spread's cheaper leg free of charge. 
Consequently, the trading fee for one option spread would be the lesser of 0.0003 Bitcoin or 12.5\% of the more expensive option price. All fees are converted into USD based on the BTC/USDT exchange rate around settlement time.

\section{Data}

We use the \href{blockchain-research-center.de}{Blockchain Research Center} to retrieve a complete set of all transactions conducted on Deribit's options written on Bitcoin and Ether. 
The exchange started operations in July 2017. 
Furthermore, we use Deribits API V2 to retrieve a complete set of funding rates for the perpetual future of Bitcoin and Ether since (approx.) July 2019.

Two conditions are imposed on the option data:
1) The maximum IV is 2.5 and the minimum IV is 0.1.
2) In case of multiple occurrences of an instrument per day, we exclusively use the first one. 
3) Furthermore, we restrict our analysis to the interval of Moneyness in [0.7, 1.3]
4) We use data exlusively past 2019-05-01, the date of first observations for the Bitcoin Perpetual Future

The purpose of the first restriction is to impose a fair bound to the options that is independent of liquidity levels. For illiquid options, occurrences of bid prices at 0\% IV and ask prices of 500\% may occur.  
The purpose the second restriction of unique usage is to prevent a bias due to the possible more frequent trading of certain options. Generally, we observe that instruments with short maturity are traded more frequently. Yet, this could also be understood as cheaper instruments trading more frequently since (Black-Scholes) option prices increase with time-to-maturity.

Bitcoin only:
We drop 0.9\% of observations due to restriction 1)
We drop 17\% of observations due to restriction 3)
Overall, we have 5,028,945 observations. Post filtering, 4,150,778 observations remain in the set.

Ethereum: 
Amount of Observations outside of allowed IV range:  0.017042133327261764
Amount of Observations outside of allowed Tau range:  0.20172172747210984
Overall Observations: 2283869
Amount of Observations in our set post filtering:  1809762

Regarding data authenticity, we refer to our previous paper Risk Premia and Pricing Kernels for Bitcoin. 
Indeed, using actual transaction data instead of orderbook data removes the threat of wash trading (etc) to our data. The most likely negative factor that could affect our data quality is wash trading. 
We assess the risk of wash trading as follows:

"A wash trade occurs when there is an act of entering into, or purporting to enter into, transactions with no intent to obtain a bona fide market position or activity that gives the false appearance of an executed transaction(s), but does not subject the Principal to any market risk or change in position or aid in price discovery." (from: https://www.ice.com/publicdocs/futures_us/Wash_Sale_FAQ.pdf)

We asses the risk of wash trading contaminating our data set as low. The reasoning behind the low risk assessment is that we are solely dealing with actual transactions associated with transaction cost. At the time of the writing, the transaction cost is 0.0003 BTC which is per current spot 12 USD. Such fees, even without consideration of spreads, make wash trading from a trader's perspective entirely infeasible.

- table of amount of transactions per group
- plot: amount of transactions over time per group (crypto)

\section{Results}

We find the returns of the zero-beta-straddles to be t-distributed and conduct a left-sided t-test under the null hypothesis of a population mean greater or equal to zero. 
The null hypothesis is not rejected for all models, vanilla and crash-resistant, for ITM and OTM options. Conversely, the null hypothesis is firmly rejected on a high significance level for ATM options. 
Therefore, we accept the alternative hypothesis of an underperformance for ATM options. Consequently, we find a similar result to Shumway and Bakshi - that is the existence of a premium for ATM contracts. 
We figure the premium to be a volatility premium since a straddle's payoff depends on the relationship of Implied Volatility to Realized Volatility. 
Unsurprisingly, the size of the premium in crypto markets is higher than the counterpart in equity markets - we figure that the reason is the higher risk in crypto markets, particularly via volatility and counterparty risk. 
Furthermore, there are mostly retailers present in the crypto option market. Some professional liquidity providers exist, but they remain few and therefore another reason for the premium's size could be illiquidity.

- the premium is larger when tau is small.

- report average returns for options grouped per strike (and overall)

\begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/btc/plots/funding_btc_vs_eth_annualized.png}
    \captionof{figure}{Annualized Funding Rate \color{blue}{BTC} vs. \color{orange}{ETH}}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/btc/plots/funding_btc_vs_eth_annualized_zoom.png}

     \captionof{figure}{(Zoomed) Annualized Funding Rate \color{blue}{BTC} vs. \color{orange}{ETH}}

\end{minipage}
\begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/btc/plots/funding_btc_vs_eth.png}
    \captionof{figure}{Funding Rate \color{blue}{BTC} vs. \color{orange}{ETH}}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/btc/plots/funding_btc_vs_eth_zoom.png}
    \captionof{figure}{(Zoomed) Funding Rate \color{blue}{BTC} vs. \color{orange}{ETH}}

\end{minipage}
   % \captionof{figure}{ATM Term Structure from 2022-01-01 until 2022-01-08. Moneyness $M \in [0.9, 1.1]$}

\label{ATM Term Structure}
\href{https://github.com/QuantLet/BitcoinOptions/blob/master/main.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,Main}

\pagebreak
\newpage

\begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/btc/plots/vanilla/no_fees/boxplot_combined_ret_daily_over_moneyness_.png}
    \captionof{figure}{BTC - Performance - Vanilla - No Fees}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/btc/plots/vanilla/fees/boxplot_combined_ret_daily_over_moneyness_.png}
     \captionof{figure}{BTC - Performance - Vanilla - Fees}

\end{minipage}
\begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/btc/plots/crash_resistant/no_fees/boxplot_combined_ret_daily_over_moneyness_.png}
    \captionof{figure}{BTC - Performance - Crash Resistant - No Fees}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/btc/plots/crash_resistant/fees/boxplot_combined_ret_daily_over_moneyness_.png}
     \captionof{figure}{BTC - Performance - Crash Resistant - Fees}

\end{minipage}
   % \captionof{figure}{ATM Term Structure from 2022-01-01 until 2022-01-08. Moneyness $M \in [0.9, 1.1]$}

\label{ATM Term Structure}
\href{https://github.com/QuantLet/BitcoinOptions/blob/master/main.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,Main}

\begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/eth/plots/vanilla/no_fees/boxplot_combined_ret_daily_over_moneyness_.png}
    \captionof{figure}{ETH - Performance - Vanilla - No Fees}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/eth/plots/vanilla/fees/boxplot_combined_ret_daily_over_moneyness_.png}
     \captionof{figure}{ETH - Performance - Vanilla - Fees}

\end{minipage}
\begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/eth/plots/crash_resistant/no_fees/boxplot_combined_ret_daily_over_moneyness_.png}
    \captionof{figure}{ETH - Performance - Crash Resistant - No Fees}
    \centering
    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/eth/plots/crash_resistant/fees/boxplot_combined_ret_daily_over_moneyness_.png}
     \captionof{figure}{ETH - Performance - Crash Resistant - Fees}

\end{minipage}
   % \captionof{figure}{ATM Term Structure from 2022-01-01 until 2022-01-08. Moneyness $M \in [0.9, 1.1]$}

\label{ATM Term Structure}
\href{https://github.com/QuantLet/BitcoinOptions/blob/master/main.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,Main}

We conduct a left-sided t-test to test for a population mean of zero. Values are rounded to 4 digits.

Compare to weekly 0.5\% loss observed in equity options.

\section{Conclusion}

Further research: 
- Evidence for leverage effect

%\begin{minipage}{\textwidth}
%    \centering
  %  \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/out/vanilla/no_fees/density/moneyness_between_95_and_105.png}
    %\captionof{figure}{Performance - Vanilla - No Fees}
    %\centering
    %\includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/out/vanilla/no_fees/density/moneyness_between_95_and_105.png}
     %\captionof{figure}{Performance - Vanilla - Fees}

%\end{minipage}
%\begin{minipage}{\textwidth}
%    \centering
%    \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/out/crash_resistant/no_fees/moneyness_between_95_and_105.png}
%    \captionof{figure}{Performance - Crash Resistant - No Fees}
 %   \centering
 %   \includegraphics[width=5.6in]{/Users/julian/src/bitcoin-vola-premium/out/crash_resistant/no_fees/moneyness_between_95_and_105.png}
 %    \captionof{figure}{Performance - Crash Resistant - Fees}

%\end{minipage}
   % \captionof{figure}{ATM Term Structure from 2022-01-01 until 2022-01-08. Moneyness $M \in [0.9, 1.1]$}

%\label{ATM Term Structure}
%\href{https://github.com/QuantLet/BitcoinOptions/blob/master/main.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,Main}

\begin{table}[ht]
\hline
\resizebox{\textwidth}{!}{\begin{tabular}{lllll}
  Moneyness    & Vanilla ex Fee & Vanilla cum Fee & Crash Resistant ex Fee & Crash Resistant cum Fee \\
\hline
\hline
0.7  & 0.6088         & 0.5854          & 0.5845                 & 0.6078                  \\
0.71 & 0.7993         & 0.7769          & 0.7762                 & 0.7987                  \\
0.72 & 0.9943         & 0.9925          & 0.9925                 & 0.9943                  \\
0.73 & 0.9348         & 0.9222          & 0.9222                 & 0.9348                  \\
0.74 & 0.9997         & 0.9996          & 0.9996                 & 0.9997                  \\
0.75 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.76 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.77 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.78 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.79 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.8  & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.81 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.82 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.83 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.84 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.85 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.86 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.87 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.88 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.89 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.9  & 1.0            & 1.0             & 1.0                    & 1.0                     \\
0.91 & 1.0            & 0.9998          & 0.9998                 & 1.0                     \\
0.92 & 1.0            & 0.9968          & 0.9968                 & 1.0                     \\
0.93 & 0.9984         & 0.9437          & 0.9437                 & 0.9984                  \\
0.94 & 0.9669         & 0.6277          & 0.6277                 & 0.9669                  \\
0.95 & 0.8817         & 0.3093          & 0.3093                 & 0.8817                  \\
0.96 & 0.3229         & 0.0129          & 0.0129                 & 0.3229                  \\
0.97 & 0.2143         & 0.0047          & 0.0047                 & 0.2143                  \\
0.98 & 0.0082         & 0.0             & 0.0                    & 0.0082                  \\
0.99 & 0.0002         & 0.0             & 0.0                    & 0.0002                  \\
1.0  & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.01 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.02 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.03 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.04 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.05 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.06 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.07 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.08 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.09 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.1  & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.11 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.12 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.13 & 0.0            & 0.0             & 0.0                    & 0.0                     \\
1.14 & 0.0172         & 0.0106          & 0.0106                 & 0.0172                  \\
1.15 & 0.0484         & 0.0347          & 0.0347                 & 0.0484                  \\
1.16 & 0.0563         & 0.0421          & 0.0421                 & 0.0564                  \\
1.17 & 0.1889         & 0.1588          & 0.1587                 & 0.1887                  \\
1.18 & 0.8503         & 0.8119          & 0.8115                 & 0.8499                  \\
1.19 & 0.9906         & 0.9861          & 0.9859                 & 0.9905                  \\
1.2  & 0.8888         & 0.8593          & 0.8562                 & 0.8862                  \\
1.21 & 0.9796         & 0.9713          & 0.9701                 & 0.9787                  \\
1.22 & 0.986          & 0.9805          & 0.9793                 & 0.985                   \\
1.23 & 0.9627         & 0.951           & 0.9464                 & 0.959                   \\
1.24 & 0.9969         & 0.9954          & 0.9945                 & 0.9962                  \\
1.25 & 0.9976         & 0.9966          & 0.9957                 & 0.9969                  \\
1.26 & 1.0            & 1.0             & 1.0                    & 1.0                     \\
1.27 & 0.9999         & 0.9999          & 0.9998                 & 0.9999                  \\
1.28 & 0.9996         & 0.9995          & 0.9993                 & 0.9995                  \\
1.29 & 0.9939         & 0.9932          & 0.9919                 & 0.9927                  \\
1.3  & 0.9436         & 0.9403          & 0.9334                 & 0.937                  
\end{tabular}}
\end{table}

%Deribit perpetual contracts are a derivative product similar to a future, however, they do not have an expiry date. The perpetual contract features funding payments. These payments have been introduced to keep the perpetual contract price as close as possible to the underlying crypto price - the relevant Deribit Index. If the perpetual contract trades at a higher price than the index, traders that have long positions need to make funding payments to the traders that have short positions. This will make the product less attractive to the long position holders and more attractive to the short position holders. This subsequently will cause a perpetual price to trade in line with the price of the index. If the perpetual trades at a price lower than the index, the short position holders will have to pay the long position holders instead.
%
%The Deribit perpetual contract features a continuous measurement of the difference between the mark price of the contract and the relevant Deribit Index. The percentage difference between these two price levels is the basis for the 8-hourly funding rate that is applied to all outstanding perpetual contracts.
%
%Funding payments are calculated every millisecond. The funding payments will be added to or subtracted from the realized PNL, which is also part of the available trading balance. At each daily settlement, the realized PNL will be moved to or from the cash balance, from which withdrawals can be made.
%
%The total funding paid will show up in the transaction log in the "funding" column. This column shows the funding amount that is applied to the trader's entire net position in the period between the relevant trade and the trade before that. Put differently: the trader can see the funding paid or received on the position between position changes.
%
%Funding Rate
%When the funding rate is positive, long position holders pay funding to the short position holders; when the funding rate is negative, short position holders pay funding to the long position holders. The funding rate is expressed as an 8-hour interest rate, and is calculated at any given time as follows:
%
%Premium Rate
%Premium Rate = ((Mark Price - Deribit Index) / Deribit Index) * 100%
%
%Funding Rate
%Sequentially, the funding rate is derived from the premium rate by applying a damper.
%
%If the premium rate is within -0.025\% and 0.025\% range, the actual funding rate will be reduced to 0.00\%.
%If the premium rate is lower than -0.025\%, then the actual funding rate will be the premium rate + 0.025\%.
%If the premium rate is higher than 0.025\%, then the actual funding rate will be the premium rate - 0.025\%.
%Additionally, the funding rate is capped at +/ - 5\% for all linear USDC perpetuals.
%Funding Rate = Maximum (0.025\%, Premium Rate) + Minimum (-0.025\%, Premium Rate)
%
%Time Fraction
%Time Fraction = Funding Rate Time Period / 8 hours
%
%The actual funding payment is calculated by multiplying the funding rate by the position size and the time fraction.
%
%Funding Payment = Funding Rate * Position Size * Time Fraction
%
%In reality, the spread of the Deribit Index and the mark price changes continuously, and all changes are taken into account. Therefore, the examples above are extreme simplifications of the actual calculations. The funding paid or received is continuously added to the realized PNL and is moved to or from the cash balance at the daily settlement, at 08:00 UTC.

\section{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-aitsahalia01optionprice}{}%
Aı̈t-Sahalia, Y., Wang, Y., \& Yared, F. (2001). Do option markets
correctly price the probabilities of movement of the underlying asset?
\emph{Journal of Econometrics}, \emph{102}(1), 67--110.
\url{https://doi.org/https://doi.org/10.1016/S0304-4076(00)00091-9}

\leavevmode\hypertarget{ref-ait_sahalia_wang1998}{}%
Aït-Sahalia, Y., \& Lo, A. W. (1998). Nonparametric estimation of
state-price densities implicit in financial asset prices. \emph{The
Journal of Finance}, \emph{53}(2), 499--547.
\url{https://doi.org/https://doi.org/10.1111/0022-1082.215228}

\leavevmode\hypertarget{ref-alexander2022net}{}%
Alexander, C., Deng, J., Feng, J., \& Wan, H. (2022). Net buying
pressure and the information in bitcoin option trades. \emph{Journal of
Financial Markets}, 100764.

\leavevmode\hypertarget{ref-inverse_options}{}%
Alexander, C., \& Imeraj, A. (2021, July). Inverse options in a
black-scholes world.
\url{https://doi.org/https://doi.org/10.48550/arXiv.2107.12041}

\leavevmode\hypertarget{ref-aloosh}{}%
Aloosh, A., \& Li, J. (2019). Direct evidence of bitcoin wash trading.
\emph{Theory of Probability \& Its Applications, 5, 285-301}.
\url{https://doi.org/http://dx.doi.org/10.2139/ssrn.3362153}

\leavevmode\hypertarget{ref-bitmex}{}%
Bitmex. (2019). Presentation to the u.s. Securities and exchange
commission on wash trading.
\url{https://www.sec.gov/comments/sr-nysearca-2019-01/srnysearca201901-5164833-183434.pdf}.
Accessed 18 January 2022

\leavevmode\hypertarget{ref-blase}{}%
Blaskowitz, O., Härdle, W. K., \& Schmidt, P. (2004). Skewness and
kurtosis trades. In (pp. 1--14).
\url{https://doi.org/10.1007/978-0-8176-8180-7_1}

\leavevmode\hypertarget{ref-speculation}{}%
Blau, B. M. (2017). Price dynamics and speculative trading in bitcoin.
\emph{Research in International Business and Finance}, \emph{41},
493--499.
\url{https://doi.org/https://doi.org/10.1016/j.ribaf.2017.05.010}

\leavevmode\hypertarget{ref-bloomberg-retail-share}{}%
Bloomberg. (2021, November 17). Retail traders slide back below 20\% of
market's total volume.
\url{https://www.bloomberg.com/news/articles/2021-11-17/retail-traders-retreat-as-choppy-markets-challenge-easy-profits}.
Accessed 24 August 2021

\leavevmode\hypertarget{ref-breeden_litzenberger78}{}%
Breeden, D. T., \& Litzenberger, R. H. (1978). Prices of
state-contingent claims implicit in option prices. \emph{The Journal of
Business}, \emph{51}(4), 621--651.
\url{http://www.jstor.org/stable/2352653}

\leavevmode\hypertarget{ref-coins_with_benefits}{}%
Chen, C. Y., \& Vinogradov, D. (2021). Coins with benefits: On
existence, pricing kernel and risk premium of cryptocurrencies.
\emph{Available at SSRN 3959984}.
\url{https://doi.org/http://dx.doi.org/10.2139/ssrn.3864578}

\leavevmode\hypertarget{ref-choi}{}%
Choi, S., \& Shin, J. (2022). Bitcoin: An inflation hedge but not a safe
haven. \emph{Finance Research Letters}, \emph{46}, 102379.
\url{https://doi.org/https://doi.org/10.1016/j.frl.2021.102379}

\leavevmode\hypertarget{ref-oil_spd}{}%
Christoffersen, P., Jacobs, K., \& Pan, X. (. (2021). The State Price
Density Implied by Crude Oil Futures and Option Prices. \emph{The Review
of Financial Studies}, \emph{35}(2), 1064--1103.
\url{https://doi.org/10.1093/rfs/hhab011}

\leavevmode\hypertarget{ref-cryptonews_bitcoin_retail_share}{}%
Coinbase, S., Cryptonews. (2022). Institutions \& retail compete for
bitcoin - whose hands are stronger?
\url{https://cryptonews.com/exclusives/institutions-retail-compete-for-bitcoin-which-is-the-biggest-9695.htm}.
Accessed 22 March 2022

\leavevmode\hypertarget{ref-decentralized_wash_trading}{}%
Cong, L., Li, X., Tang, K., \& Yang, Y. (2019). Crypto wash trading.
\emph{SSRN Electronic Journal}.
\url{https://doi.org/10.2139/ssrn.3530220}

\leavevmode\hypertarget{ref-jackwerth_pk}{}%
Cuesdeanu, H., \& Jackwerth, J. C. (2018). The pricing kernel puzzle:
survey and outlook. \emph{Annals of Finance}, \emph{14}(3), 289--329.
\url{https://doi.org/10.1007/s10436-017-0317-9}

\leavevmode\hypertarget{ref-cross_border_payments}{}%
Deng, Q. (2020). Application analysis on blockchain technology in
cross-border payment. In.
\url{https://doi.org/10.2991/aebmr.k.200306.050}

\leavevmode\hypertarget{ref-deribit-insurance-fund}{}%
Deribit. (2020a). Deribit insurance fund.
\url{https://www.deribit.com/main\#/insurance}. Accessed 4 January 2021

\leavevmode\hypertarget{ref-synthetic-btc-index}{}%
Deribit. (2020b). Deribit btc-usd index.
\url{https://www.deribit.com/main\#/indexes}. Accessed 2 January 2021

\leavevmode\hypertarget{ref-market-maker-obligations}{}%
Deribit. (2020c). Deribit market maker obligations.
\url{https://www.deribit.com/pages/docs/options}. Accessed 2 January
2021

\leavevmode\hypertarget{ref-deribit_retail_piechart}{}%
Deribit. (2022a). Deribit volume share.
\url{https://www.deribit.com/statistics/BTC/market-data}. Accessed 25
July 2022

\leavevmode\hypertarget{ref-deribit_margin}{}%
Deribit. (2022b). Deribit margin requirements.
\url{https://legacy.deribit.com/pages/docs/options}. Accessed 2 May 2022

\leavevmode\hypertarget{ref-deribit_fees}{}%
Deribit. (2022c). Deribit margin requirements.
\url{https://www.deribit.com/kb/fees}. Accessed 2 May 2022

\leavevmode\hypertarget{ref-deribit-rate-limit}{}%
Deribit. (2023). Deribit rate limits.
\url{https://www.deribit.com/kb/deribit-rate-limits}. Accessed 16 April
2021

\leavevmode\hypertarget{ref-vola_premium}{}%
Eraker, B. (2021). The volatility premium. \emph{Quarterly Journal of
Finance (QJF)}, \emph{11}(03), 1--35.
\url{https://EconPapers.repec.org/RePEc:wsi:qjfxxx:v:11:y:2021:i:03:n:s2010139221500142}

\leavevmode\hypertarget{ref-networkeffect}{}%
Gandal, N., \& Halaburda, H. (2016). Can we predict the winner in a
market with network effects? Competition in cryptocurrency market.
\emph{Games}, \emph{7}(3). \url{https://doi.org/10.3390/g7030016}

\leavevmode\hypertarget{ref-Grith09sim}{}%
Grith, M., Härdle, W. K., \& Park, J. (2009). Shape invariant modelling
pricing kernels and risk aversion.
\url{https://doi.org/10.2139/ssrn.2894289}

\leavevmode\hypertarget{ref-hall_1991}{}%
Hall, P. (1991). Edgeworth expansions for nonparametric density
estimators, with applications. \emph{Statistics}, \emph{22}(2),
215--232. \url{https://doi.org/10.1080/02331889108802305}

\leavevmode\hypertarget{ref-hayes}{}%
Hayes, A. S. (2019). Bitcoin price and its marginal cost of production:
Support for a fundamental value. \emph{Applied Economics Letters},
\emph{26}(7), 554--560.
\url{https://doi.org/10.1080/13504851.2018.1488040}

\leavevmode\hypertarget{ref-w2yo}{}%
Härdle, W. K., Okhrin, Y., \& Wang, W. (2014). Uniform Confidence Bands
for Pricing Kernels. \emph{Journal of Financial Econometrics},
\emph{13}(2), 376--413. \url{https://doi.org/10.1093/jjfinec/nbu002}

\leavevmode\hypertarget{ref-pricing_cryptocurrency_options}{}%
Hou, A. J., Wang, W., Chen, C. Y. H., \& Härdle, W. K. (2020). Pricing
Cryptocurrency Options*. \emph{Journal of Financial Econometrics},
\emph{18}(2), 250--279. \url{https://doi.org/10.1093/jjfinec/nbaa006}

\leavevmode\hypertarget{ref-xfg_spd}{}%
Huynh, K., Kervella, P., \& Zheng, J. (2002). \emph{Estimating
state-price densities with nonparametric regression} (SFB 373 Discussion
Papers No. 2002,40). Humboldt University of Berlin, Interdisciplinary
Research Project 373: Quantification; Simulation of Economic Processes.
\url{https://EconPapers.repec.org/RePEc:zbw:sfb373:200240}

\leavevmode\hypertarget{ref-pumpndump}{}%
La Morgia, M., Mei, A., Sassi, F., \& Stefa, J. (2020). Pump and dumps
in the bitcoin era: Real time detection of cryptocurrency market
manipulations. In \emph{2020 29th international conference on computer
communications and networks (icccn)} (pp. 1--9).
\url{https://doi.org/10.1109/ICCCN49398.2020.9209660}

\leavevmode\hypertarget{ref-vola_stylized_facts}{}%
Masset, P. (2011). Volatility stylized facts. \emph{SSRN Electronic
Journal}. \url{https://doi.org/10.2139/ssrn.1804070}

\leavevmode\hypertarget{ref-satoshi2009}{}%
Nakamoto, S. (2009). Bitcoin: A peer-to-peer electronic cash system.
\url{http://www.bitcoin.org/bitcoin.pdf}

\leavevmode\hypertarget{ref-fundamental-theorem-of-asset-pricing}{}%
Pascucci, A., \& Agliardi, R. (2011). PDE and martingale methods in
option pricing. In (pp. 429--495).

\leavevmode\hypertarget{ref-rodgers2019}{}%
Rodgers, T. (2019, April 4). 95 percent of volume could be wash trading
as bitcoin price surges.
\url{https://www.forbes.com/sites/tomrodgers1/2019/04/04/99-of-volume-could-be-wash-trading-as-bitcoin-takes-back-5000/?sh=4640bcbc23d5}.
Accessed 15 April 2023

\leavevmode\hypertarget{ref-Rookley97}{}%
Rookley, C. (1997). Fully exploiting the information content of intra
day option quotes: Applications in option pricing and risk management.

\leavevmode\hypertarget{ref-sar2017}{}%
Sar, M. (2017). Dodd-frank and the spoofing prohibition in commodities
markets.

\leavevmode\hypertarget{ref-deriv_volume}{}%
Skew.com. (2022). Open interest for bitcoin options over time. Skew is a
leading crypto data analytics platform for institutional clients.
\url{https://analytics.skew.com/dashboard/bitcoin-options}. Accessed 25
July 2022

\leavevmode\hypertarget{ref-brazilian_speculation}{}%
Smaniotto, E. N., \& Neto, G. B. (2022). Speculative trading in bitcoin:
A brazilian market evidence. \emph{The Quarterly Review of Economics and
Finance}, \emph{85}, 47--54.
\url{https://doi.org/https://doi.org/10.1016/j.qref.2020.10.024}

\leavevmode\hypertarget{ref-cross_border_stellar}{}%
Stellar Development Foundation. (2022). 2022 research reveals high
awareness and growing cross-border use of cryptocurrency in four key
markets.
\url{https://www.prnewswire.com/news-releases/2022-research-reveals-high-awareness-and-growing-cross-border-use-of-cryptocurrency-in-four-key-markets-301512401.html}.
Accessed 22 July 2022

\leavevmode\hypertarget{ref-nickszabo}{}%
Szabo, N. (2008). Bit gold.
\url{https://unenumerated.blogspot.com/2005/12/bit-gold.html}. Accessed
28 July 2022

\leavevmode\hypertarget{ref-tuccella2021}{}%
Tuccella, J.-N., Nadler, P., \& Șerban, O. (2021). Protecting retail
investors from order book spoofing using a gru-based detection model.

\end{document}